---
title: |
  Does reported walking measure accelerometer activity —\
  or walking?
subtitle: "Validating self-reported walking in the OAI against accelerometry — a proposal for joint work"
mainfont: Arial
fontsize: 10pt
# Navy title, subtitle and section headings: the template's heading-color and a heading show rule
brand:
  typography:
    headings:
      color: "#0b2e59"
bibliography: references.bib
bibliographystyle: american-medical-association
format:
  typst:
    papersize: us-letter
    margin:
      x: 0.85in
      y: 0.75in
    include-in-header:
      text: |
        #set text(hyphenate: false)
        #set bibliography(title: [References])
        // The AMA style ends every title with a full stop, which doubles after a title ending in "?"
        #show bibliography: it => { show regex("\\?\\."): "?"; it }
        #set page(footer: context align(center, text(size: 8pt, fill: luma(90))[Brief — page #counter(page).display()]))
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: cairo_pdf
    fig-width: 7
---

```{r setup}
run <- oaireport::load_results()$default
assumption <- function(key) run$assumptions$assumptions[[key]]$value
num <- function(x, digits = 2) formatC(x, format = "f", digits = digits, big.mark = ",")
count <- function(x) formatC(x, format = "d", big.mark = ",")
fmt <- function(x) format(x, trim = TRUE, big.mark = ",")
pct <- function(x) paste0(num(100 * x, 0), "%")
range_text <- function(x, digits = 2) {
  lo <- num(min(x), digits); hi <- num(max(x), digits)
  if (lo == hi) lo else paste0(lo, "–", hi)
}
colours <- oaireport::brief_colours
# The one row matching every condition; stops when there is none or several (a changed schema)
one <- function(df, ...) {
  cond <- list(...)
  keep <- Reduce(`&`, Map(function(col, value) !is.na(df[[col]]) & df[[col]] == value, names(cond), cond))
  row <- df[keep, , drop = FALSE]
  if (nrow(row) != 1) {
    stop("expected one row for ", paste(names(cond), unlist(cond), sep = " = ", collapse = ", "),
         ", found ", nrow(row), call. = FALSE)
  }
  row
}
cls <- run$validity_classification
cp <- run$validity_components_pase
cp$visit <- sprintf("%02d", as.integer(cp$visit))
ci <- run$validity_components_item
pase <- function(v, component, comparator, level, statistic) {
  one(cp, visit = v, component = component, comparator = comparator, level = level, statistic = statistic)
}
item <- function(sample, component, level, statistic) {
  one(ci, sample = sample, component = component, level = level, statistic = statistic)
}
labels <- lapply(assumption("components.answer_labels"), unlist)
# Cut k of an answer means "answer at or above level cuts[k]". The PASE days answer has levels 0
# to 3 (labels$pase_days[i + 1] names level i), the walking item's times-per-month answer levels
# 1 to 3 (labels$times[i]), so the labels follow components.*_cuts and not a fixed position.
pase_cuts <- unlist(assumption("components.pase_frequency_cuts"))
item_cuts <- unlist(assumption("components.item_times_cuts"))
# oaireport::cut_labels() names each cut: "At least <label>", or the label alone at the top level
pase_names <- oaireport::cut_labels(pase_cuts, labels$pase_days, first_level = 0)
item_names <- oaireport::cut_labels(item_cuts, labels$times, first_level = 1)
lower_first <- function(x) paste0(tolower(substr(x, 1, 1)), substring(x, 2))
visits <- c("06", "08")
wave_names <- c(`06` = "48 months", `08` = "72 months")
paired <- assumption("pase.device_pairing") == "adjacent_wave"
device_wave_text <- if (paired) "that visit's own device wave" else "the participant's combined device waves"
pase_any_is_walker <- assumption("pase.walker_threshold") == 0
walker_rule <- assumption("reference.walker_rule")
rule_short <- switch(walker_rule,
  bout_days = sprintf("on average at least %s days a week with a purposeful bout", fmt(assumption("reference.min_bout_days_per_week"))),
  any_bout = "at least one purposeful bout",
  bout_minutes = sprintf("on average at least %s minutes a week in purposeful bouts", fmt(assumption("reference.min_bout_minutes_per_week"))),
  stop("unknown reference.walker_rule: ", walker_rule))

n_validation <- run$metrics[["sample.validation.persons"]]
n_walkers <- run$metrics[["sample.validation.walkers"]]
se <- one(cls, sample = "validation", measure = "se")$estimate
sp <- one(cls, sample = "validation", measure = "sp")$estimate
j_item <- vapply(c("cut1", "cut3"), function(k) item("validation", "times", k, "j")$estimate, numeric(1))
j_pase_any <- vapply(visits, function(v) pase(v, "frequency", "device_walker", "cut1", "j")$estimate, numeric(1))
j_pase_mid <- vapply(visits, function(v) pase(v, "frequency", "device_walker", "cut2", "j")$estimate, numeric(1))
correction <- vapply(c("cut1", "cut3"), function(k) item("lo_subset", "times", k, "correction")$estimate, numeric(1))
strict <- lower_first(item_names[["cut3"]])  # cut 3 of the item: the stricter walker definition
# Lo et al. 2022's published new-pain odds ratio, and the tipping-point grid's new-pain cells
t2 <- run$lo2022_t2
published_or <- oaireport::parse_or(t2$published[t2$metric == "t2.new_pain.or_adj"])$or
grid_np <- run$bias_tipping[run$bias_tipping$outcome == "new_pain" & !is.na(run$bias_tipping$or), ]
grid_holds <- sum(grid_np$excludes_1)
grid_feasible <- nrow(grid_np)
grid_se_floor <- min(grid_np$se)
bout_minutes <- assumption("device.purposeful_bout_minutes")
cutpoint <- assumption("device.mv_cutpoint")
guide <- assumption("reference.min_bout_minutes_per_week")
rho_week <- vapply(visits, function(v) pase(v, "weekly", "purposeful_week", "all", "rho")$estimate, numeric(1))
diff_week <- vapply(visits, function(v) pase(v, "weekly", "purposeful_week", "all", "median_diff")$estimate, numeric(1))
kappa_week <- vapply(visits, function(v) pase(v, "weekly", "purposeful_week", "guideline", "kappa")$estimate, numeric(1))
repro <- run$device_reproduction
# Every column the brief reads; a missing one would otherwise give NULL (and a sum of 0), silently
repro_columns <- c("wave", "release_days", "matched_days", "mv_mismatch_days", "bout_mismatch_days",
                   "wear_mismatch_days")
if (length(absent <- setdiff(repro_columns, names(repro)))) {
  stop("device_reproduction lacks column(s) ", paste(absent, collapse = ", "), call. = FALSE)
}
repro$wave_name <- wave_names[sprintf("%02d", as.integer(repro$wave))]
```

```{=typst}
#set par(justify: false)
```

```{r authors}
#| results: asis
cat("```{=typst}",
    sprintf("#text(size: 9.5pt, fill: luma(70))[%s] #v(4pt)",
            oaireport::typst_text(oaireport::author_block("brief.local.yml"))),
    "```", sep = "\n")
```

```{=typst}
#set text(size: 11pt)
```

The Osteoarthritis Initiative (OAI) asked participants whether they had walked for exercise since age 50, and the answer predicts knee outcomes [@lo2022]. In that paper, walkers had about `r pct(1 - published_or)` lower odds of developing frequent knee pain; we also asked how sensitive that association is to how walkers are classified. We checked that answer, and the walking question of the Physical Activity Scale for the Elderly (PASE), against what the ActiGraph GT1M recorded at 48 and 72 months in `r count(n_validation)` participants. Most say yes (`r pct(n_walkers / n_validation)`), so the question alone separates walkers poorly; *how often* people say they walk carries most of the signal. Our device reference is built from activity counts, which cannot tell walking from other moderate activity, and that is where your step and cadence data come in.

```{=typst}
#set text(size: 10pt)
```

```{r fig1}
#| include: false
cuts <- c("cut1", "cut2", "cut3")
item_pts <- do.call(rbind, lapply(cuts, function(k) data.frame(
  source = "Walking item (times per month)", cut = k,
  se = item("validation", "times", k, "se")$estimate,
  fpr = 1 - item("validation", "times", k, "sp")$estimate,
  current = k == "cut1")))
pase_pts <- do.call(rbind, lapply(visits, function(v) do.call(rbind, lapply(cuts, function(k) data.frame(
  source = sprintf("PASE days per week, %s", wave_names[[v]]), cut = k,
  se = pase(v, "frequency", "device_walker", k, "se")$estimate,
  fpr = 1 - pase(v, "frequency", "device_walker", k, "sp")$estimate,
  current = k == "cut1" && pase_any_is_walker)))))
pts <- rbind(item_pts, pase_pts)
pts$source <- factor(pts$source, levels = unique(pts$source))
# One source per row (a single row is wider than the column); keys show the line, solid or dashed
source_key <- ggplot2::guide_legend(ncol = 1, order = 1, override.aes = list(shape = NA),
                                    theme = ggplot2::theme(legend.key.width = ggplot2::unit(2, "lines")))
fig1 <- ggplot2::ggplot(pts, ggplot2::aes(fpr, se, colour = source, linetype = source)) +
  oaireport::youden_contours() +
  ggplot2::geom_line(linewidth = 0.6) +
  # Filled markers with a white outline, item drawn last: its triangle sits just below PASE's
  ggplot2::geom_point(data = pts[rev(seq_len(nrow(pts))), ], ggplot2::aes(shape = current, fill = source),
                      size = 2.6, colour = "white", stroke = 0.6) +
  ggplot2::scale_colour_manual(values = c(colours[["item"]], colours[["pase"]], colours[["pase"]]), name = NULL) +
  ggplot2::scale_linetype_manual(values = c("solid", "solid", "dashed"), name = NULL) +
  ggplot2::scale_fill_manual(values = c(colours[["item"]], colours[["pase"]], colours[["pase"]]), name = NULL) +
  ggplot2::scale_shape_manual(values = c(`TRUE` = 24, `FALSE` = 21), name = NULL,
                              labels = c(`TRUE` = "Definition in use", `FALSE` = "Stricter, frequency-based")) +
  ggplot2::coord_equal(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
  ggplot2::labs(x = "1 − specificity", y = "Sensitivity") +
  ggplot2::guides(colour = source_key, linetype = source_key, fill = source_key,
                  shape = ggplot2::guide_legend(order = 2, override.aes = list(fill = "black"))) +
  oaireport::theme_oai(base_size = 8) +
  ggplot2::theme(legend.position = "bottom", legend.box = "vertical",
                 legend.spacing.y = ggplot2::unit(2, "pt"), legend.key.spacing.y = ggplot2::unit(0, "pt"),
                 legend.margin = ggplot2::margin(0, 0, 0, 0), legend.box.spacing = ggplot2::unit(4, "pt"),
                 plot.margin = ggplot2::margin(14, 8, 4, 4))
oaireport::save_figure(fig1, "brief_fig1", "1col", height = 3.35)
```

```{r page1}
#| results: asis
# Sidebar values stay short (15 pt in a narrow column): PASE gives the 48-month value and puts
# the 72-month one in its label
results <- data.frame(
  value = c(sprintf("%s / %s", num(se), num(sp)),
            sprintf("%s → %s", num(j_item[[1]]), num(j_item[[2]])),
            sprintf("%s → %s", num(j_pase_any[["06"]]), num(j_pase_mid[["06"]])),
            sprintf("%d of %d", grid_holds, grid_feasible)),
  label = c("sensitivity / specificity of “walked for exercise” against the device",
            sprintf("Youden's J: any walker → walks %s", strict),
            sprintf("Youden's J for PASE at 48 months: any walking → %s a week (72 months: %s → %s)",
                    lower_first(pase_names[["cut2"]]), num(j_pase_any[["08"]]), num(j_pase_mid[["08"]])),
            sprintf("plausible accuracy settings (sensitivity %s or more) in which the walking–knee-pain association in Lo et al. 2022 holds", num(grid_se_floor))))
cat(oaireport::sidebar_figure(
  results,
  ask = "A joint study: rebuild the reference from your step and cadence data (for example, minutes at 100 steps/min or more) and re-run the validation and the classification sensitivity analysis, pre-specified and jointly authored.",
  image = "figures/brief_fig1.png", sidebar = 0.35,
  caption = "Each point is a walker definition, placed by its sensitivity and 1 − specificity against the device-defined walker. Triangles: the definitions in use (any walking); the three nearly coincide at the top right. Stricter, frequency-based definitions (circles) give up some sensitivity but gain more specificity, so they move down and to the left, away from the chance line toward higher J. Grey dashed lines join points of equal Youden's J, with each value above the top edge; J = 0 is chance."),
  sep = "\n")
```

*The appendix summarises every result for clinical readers; the full technical report is available on request.*

```{=typst}
#pagebreak()
```

## What the counts already show

Reported walking frequency tracks the device in a clear gradient, which a yes/no question discards.

```{r fig2}
#| fig-height: 2.6
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
level_rows <- function(get, levels, names, question, fill) {
  do.call(rbind, lapply(seq_along(levels), function(i) {
    r <- get(levels[[i]])
    data.frame(question = question, answer = names[[i]], estimate = r$estimate, lo = r$lo, hi = r$hi,
               fill = if (i == 1) paste0(fill, "_ref") else fill)
  }))
}
# PASE days answer levels 0 to length(labels) - 1: level i is labels$pase_days[i + 1]
pase_levels <- as.character(seq_along(labels$pase_days) - 1)
bars <- rbind(
  level_rows(function(l) item("validation", "times", l, "share"),
             c("none", as.character(seq_along(labels$times))), c("Non-walkers", labels$times),
             "Walking item: times per month", "item"),
  level_rows(function(l) pase("06", "frequency", "device_walker", l, "share"),
             pase_levels, labels$pase_days, "PASE: days walked last week (48 months)", "pase"))
bars$answer <- factor(bars$answer, levels = rev(unique(bars$answer)))
bars$question <- factor(bars$question, levels = unique(bars$question))
x_top <- min(1, max(bars$hi, na.rm = TRUE) + 0.15)
fig2 <- ggplot2::ggplot(bars, ggplot2::aes(estimate, answer, fill = fill)) +
  ggplot2::geom_col(width = 0.65) +
  ggplot2::geom_errorbar(ggplot2::aes(xmin = lo, xmax = hi), width = 0.25, linewidth = 0.3, orientation = "y") +
  ggplot2::geom_text(ggplot2::aes(x = hi, label = pct(estimate)), hjust = -0.25, size = 2.6,
                     family = oaireport::oai_font()) +
  ggplot2::scale_fill_manual(values = c(item_ref = colours[["muted"]], item = colours[["item"]],
                                        pase_ref = colours[["pase_muted"]], pase = colours[["pase"]]),
                             guide = "none") +
  ggplot2::scale_x_continuous(labels = function(x) paste0(100 * x, "%"), limits = c(0, x_top), expand = c(0, 0)) +
  ggplot2::facet_wrap(~question, scales = "free_y") +
  ggplot2::labs(x = "Share who are device-defined walkers (95% CI)", y = NULL) +
  oaireport::theme_oai(base_size = 8)
oaireport::save_figure(fig2, "brief_fig2", "2col", height = 2.6)
print(fig2)
pase72 <- vapply(pase_levels, function(l) pase("08", "frequency", "device_walker", l, "share")$estimate, numeric(1))
cat(oaireport::table_note(sprintf(
  "Share of participants who are device-defined walkers (%s, where a purposeful bout is at least %s minutes at %s counts/min or more), by reported walking frequency, with Wilson 95%% CIs. Walking item: validation sample, non-walkers and the walkers who answered how many times a month they walked. PASE: participants with a 48-month PASE answer, against %s. At 72 months the PASE shares run from %s (%s) to %s (%s).",
  rule_short, fmt(bout_minutes), fmt(cutpoint), device_wave_text,
  pct(pase72[[1]]), tolower(labels$pase_days[1]), pct(pase72[[length(pase72)]]), labels$pase_days[length(labels$pase_days)]),
  label = "Figure 2.", close_group = TRUE), sep = "\n")
```

Combining PASE's two walking answers (days × hours a day) gives an estimate of weekly walking. It ranks people only modestly against the device's purposeful-bout minutes (Spearman ρ `r range_text(rho_week)`), sits a median of about `r range_text(diff_week, 0)` minutes a week above the device, and agrees weakly on who walks at least `r fmt(guide)` minutes a week (Cohen's κ `r range_text(kappa_week)`). That is below the 0.40 to 0.57 reported for walking questions against device step counts in older adults, and in the lower part of the 0.17 to 0.52 reported for overall-activity scores [@hagiwara2008; @heesch2011; @giles2009; @harris2009]. Those studies used steps as the reference, which is part of the case for a step-based reference here.

```{r fig3}
#| fig-height: 2.6
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
cal <- cp[cp$component == "weekly" & cp$comparator == "purposeful_week" & cp$statistic == "median", ]
cal$level <- factor(cal$level, levels = unique(cal$level))  # the step writes the bins in order
cal$visit_name <- wave_names[cal$visit]
fig3 <- ggplot2::ggplot(cal, ggplot2::aes(level, estimate)) +
  ggplot2::geom_pointrange(ggplot2::aes(ymin = lo, ymax = hi), colour = colours[["pase"]], size = 0.3) +
  ggplot2::geom_text(ggplot2::aes(y = hi, label = paste0("n = ", count(n))), vjust = -0.6, size = 2.2,
                     colour = "grey35", family = oaireport::oai_font()) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0.05, 0.15))) +
  ggplot2::facet_wrap(~visit_name) +
  ggplot2::labs(x = "Estimated weekly walking from PASE (hours per week)",
                y = "Device purposeful-bout\nminutes per week") +
  oaireport::theme_oai(base_size = 8)
oaireport::save_figure(fig3, "brief_fig3", "2col", height = 2.6)
print(fig3)
cat(oaireport::table_note(sprintf(
  "Device purposeful-bout minutes per week (point: median; bar: interquartile range) within each band of estimated weekly walking from PASE (days × hours a day at the PASE scoring midpoints), against %s. n: participants in the band. A self-report that tracked the device would show medians rising steadily from left to right.",
  device_wave_text), label = "Figure 3.", close_group = TRUE), sep = "\n")
```

```{=typst}
#pagebreak()
```

## Why counts aren't enough, and what steps would add

```{r weak-link}
#| results: asis
cat(oaireport::callout(sprintf(
  "A purposeful bout lasts at least %s minutes at moderate-to-vigorous intensity: it starts when %s of %s minutes reach %s counts/min and ends when %s minutes in the window fall below it. Other moderate activity, such as brisk housework, dancing or aerobics, can form a bout, while slow walking may never reach the cut-point. The device-defined walker is therefore a proxy: people whose walking is too slow to register count as device non-walkers, which pulls the questions' specificity down, so part of the low specificity may belong to the reference rather than to the questions.",
  fmt(bout_minutes), fmt(assumption("device.bout_need")), fmt(assumption("device.bout_window")), fmt(cutpoint),
  fmt(assumption("device.bout_stop_below"))), kind = "warn", title = "The weak link."), sep = "\n")
```

Timing is a second gap: the walking item was asked at 96 months about walking since age 50, years after the device waves. Timing alone does not explain the weak agreement, though: PASE, asked at the device visits themselves, separates walkers about as weakly (Youden's J `r range_text(j_pase_any)` for any walking). The question also asks about a habit of walking for exercise since age 50, the device one week of current activity, so the two may measure different things.

**How sensitive is the Lo et al. 2022 association to classification?** If some participants were misclassified, a re-estimate that allows for it divides the gap in walking between people with and without the outcome by Youden's J. With the walker definition in use, that multiplies the gap by `r num(correction[[1]], 1)`; with the strictest frequency-based definition, by `r num(correction[[2]], 1)`, so a stricter definition gives a steadier re-estimate. If misclassification is the same in people with and without the outcome, the walking–knee-pain association holds in `r grid_holds` of `r grid_feasible` plausible accuracy settings (appendix).

```{r fig4}
#| fig-height: 2.1
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
cut_names <- c(cut1 = "Any walker (in use)", cut2 = item_names[["cut2"]], cut3 = item_names[["cut3"]])
cf <- do.call(rbind, lapply(names(cut_names), function(k) {
  r <- item("lo_subset", "times", k, "correction")
  data.frame(definition = cut_names[[k]], estimate = r$estimate, lo = r$lo, hi = r$hi)
}))
cf$definition <- factor(cf$definition, levels = rev(cut_names))
top <- max(c(cf$hi, cf$estimate), na.rm = TRUE) * 1.15
cf$hi_draw <- ifelse(is.na(cf$hi), top, cf$hi)
fig4 <- ggplot2::ggplot(cf, ggplot2::aes(estimate, definition)) +
  ggplot2::geom_col(fill = colours[["item"]], width = 0.6) +
  ggplot2::geom_errorbar(data = cf[!is.na(cf$hi), , drop = FALSE], ggplot2::aes(xmin = lo, xmax = hi),
                         width = 0.25, linewidth = 0.3, orientation = "y") +
  ggplot2::geom_segment(data = cf[is.na(cf$hi), , drop = FALSE],
                        ggplot2::aes(x = lo, xend = hi_draw, yend = definition), linewidth = 0.4,
                        arrow = ggplot2::arrow(length = ggplot2::unit(0.08, "in"))) +
  ggplot2::geom_text(ggplot2::aes(x = 0, label = paste0(num(estimate, 1), "×")), hjust = -0.2,
                     colour = "white", size = 2.8, fontface = "bold", family = oaireport::oai_font()) +
  ggplot2::scale_x_continuous(limits = c(0, top), expand = c(0, 0)) +
  ggplot2::labs(x = "Multiplier 1/J (95% CI)", y = NULL) +
  oaireport::theme_oai(base_size = 8)
oaireport::save_figure(fig4, "brief_fig4", "2col", height = 2.1)
print(fig4)
cat(oaireport::table_note(sprintf(
  "Lo et al. 2022 participants who answered the walking item and have device data (%s classified). For each walker definition by the times-per-month answer, 1/J shows how much a re-estimate that allows for misclassification would multiply the difference in walking between people with and without the outcome — how much the re-estimate depends on the walker definition; the interval comes from J's 95%% CI.%s",
  count(item("lo_subset", "times", "cut1", "correction")$n),
  if (any(is.na(cf$hi))) " An arrow marks an interval with no upper limit (J's interval reaches 0)." else ""),
  label = "Figure 4.", close_group = TRUE), sep = "\n")
```

```{r steps-add}
#| results: asis
cat(oaireport::callout(
  "Walking-specific references: steps per day, minutes at 100 steps/min or more (about 3 METs in ostensibly healthy adults aged 61 to 85) and step-based bouts. OAI cadence bands at this threshold have already been used.",
  kind = "note", title = "What steps and cadence add.", cite = c("tudorlocke2021", "fenton2018", "master2021")), sep = "\n")
cat(oaireport::callout(
  "GT1M step counts come from the device's step-detection algorithm, which undercounts slow walking: 64% of steps were counted at 54 m/min, and 19% were missed below 0.8 m/s in adults around 79 years old. How should slow walkers with knee OA be handled?",
  kind = "warn", title = "An open question for your expertise.", cite = c("abel2008", "storti2008")), sep = "\n")
```

```{=typst}
#pagebreak()
```

## The proposed joint study

**Aim.** Revalidate the 96-month walking item and PASE walking, and revisit the associations in Lo et al. 2022, against walking-specific references (steps per day, minutes at 100 steps/min or more, step-based bouts) defined together and pre-specified.

```{r roles}
roles <- data.frame(
  `You bring` = c("Step and cadence data, and how they were derived",
                  "Step-algorithm expertise; handling of slow gait",
                  "Co-design of the reference definitions"),
  `We bring` = c("Coding of the walking item and PASE; an open, reproducible pipeline",
                 "Validation statistics; the Lo et al. 2022 sensitivity analysis",
                 "Drafting; shared code; aggregate-only outputs"),
  check.names = FALSE)
oaireport::compare_table(roles, widths = c(1, 1), multipage = FALSE)
```

**What we would need.** The existing daily step and cadence-band summaries at 48 and 72 months [@lo2015; @fenton2018; @master2021]; per-minute steps, if available; and a short description of how the steps were derived. The public release holds activity counts only: which files hold the steps?

```{r built}
#| results: asis
matched <- sprintf("%s of %s release person-days at %s", count(repro$matched_days), count(repro$release_days),
                   repro$wave_name)
differ_text <- if (all(repro$mv_mismatch_days == 0 & repro$bout_mismatch_days == 0)) {
  "with the same moderate-to-vigorous and bout minutes on every matched day"
} else sprintf("with moderate-to-vigorous minutes differing on %s and bout minutes on %s matched days",
               count(sum(repro$mv_mismatch_days)), count(sum(repro$bout_mismatch_days)))
# wear_mismatch_days counts matched days whose wear time differs by half a minute or more
wear_text <- if (sum(repro$wear_mismatch_days) > 0) {
  sprintf(", and wear time within half a minute on all but %s of them", count(sum(repro$wear_mismatch_days)))
} else ", with the same wear time"
cat(oaireport::callout(sprintf(
  "Our pipeline rebuilds the OAI release's day-level accelerometry files from the minute counts, following the OAI processing standard: it matches %s, %s%s. A step-based reference therefore drops into the same validation and sensitivity analyses.",
  paste(matched, collapse = " and "), differ_text, wear_text), kind = "note", title = "Already built.", cite = "song2010"),
  sep = "\n")
```

**Outputs.** A joint paper (or papers) and shared, open code.

```{=typst}
#pagebreak()
```

**Appendix.** The appendix summarises every result for clinical readers; the full technical report is available on request.
