---
title: "Walking question and accelerometry in the OAI: results"
subtitle: "Appendix to the brief: the main results in plain language. The full technical report is available on request."
mainfont: Arial
fontsize: 10pt
brand:
  typography:
    headings:
      color: "#0b2e59"
bibliography: references.bib
bibliographystyle: american-medical-association
format:
  typst:
    papersize: us-letter
    margin:
      x: 0.85in
      y: 0.75in
    include-in-header:
      text: |
        #set text(hyphenate: false, overhang: false)
        #set bibliography(title: [References])
        #set page(footer: context align(center, text(size: 8pt, fill: luma(90))[Appendix — page #counter(page).display()]))
        // The AMA style ends every title with a full stop, which doubles after a title ending in "?"
        #show bibliography: it => { show regex("\\?\\."): "?"; it }
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: cairo_pdf
    fig-width: 7
---

```{=typst}
#set par(justify: false)
```

```{r setup}
source("report_setup.R")
lab <- oaireport::numberer("A")
colours <- oaireport::brief_colours
reps <- fmt(assumption("validity.bootstrap_reps"))  # every bootstrap in validity.R and components.R
adjusted_ci <- if (assumption("validity.median_regression_se") == "boot") {
  sprintf("CI for the adjusted difference: the estimate ± 1.96 × its standard error, the standard error from refitting the model in %s bootstrap resamples of participants.", reps)
} else {
  "CI for the adjusted difference: the estimate ± 1.96 × its standard error, the standard error from the median-regression sandwich formula (quantreg's \"nid\" method)."
}
# What makes a setting or a draw impossible (oaimodels::reclassify): the grid's blank cells come from a
# sensitivity below the share of walkers by the question in a group; the simulation's draws mostly from a
# specificity so low that a group has no true walkers left
impossible_grid <- "a negative number of people in a group (for example, of true non-walkers, when the sensitivity is below the share of people the question calls walkers among those with or without the outcome) or sensitivity + specificity of 1 or less"
impossible_sim <- "a negative number of people in a group (for example, of true walkers among people with the outcome) or sensitivity + specificity of 1 or less"
notes <- oaireport::tech_notes(c(
  median_difference = "Median difference: the Hodges–Lehmann estimate, the median of the differences across all pairs of people, one from each group; it is robust to skewed data and need not equal the gap between the two group medians. Its 95% CI is the Hodges–Lehmann interval from the Wilcoxon rank-sum test.",
  adjusted_ci = adjusted_ci,
  bootstrap_ci = sprintf("Bootstrap 95%% CI: participants are resampled with replacement %s times, the estimate is recomputed each time, and the middle 95%% of the results forms the interval (a percentile interval).", reps),
  rank_correlation = "Rank correlation: Spearman's correlation between the ranks of two measures, from −1 to 1. Its 95% CI uses Fisher's z transformation with the Bonett–Wright standard error.",
  trend_test = sprintf("Test for trend: the Jonckheere–Terpstra test of whether a measure rises across ordered groups; its p-value comes from %s random permutations of the group labels.", count(jt_b)),
  # Part 2 (validity.R's classification_by_stratum; bias.R's tipping grid and pba draws)
  subgroup_test = "Subgroup differences: a likelihood-ratio test from logistic regression of whether the answer matches the device on the subgroup, among device-defined walkers (sensitivity) and device-defined non-walkers (specificity) separately.",
  tipping_point = sprintf("Tipping-point analysis: the replication's adjusted model is refitted with the walking answers reclassified at random as if the question had a fixed sensitivity (%s to 1) and specificity (%s to 1), in steps of %s, the same in people with and without the outcome. Each setting is refitted %s times and shows the median odds ratio, with a 95%% interval that also includes the model's own uncertainty. A setting is impossible when it implies %s.",
                          num(min(run$bias_tipping$se)), num(min(run$bias_tipping$sp)), num(assumption("bias.tipping_step")),
                          count(assumption("bias.tipping_iterations")), impossible_grid),
  misclassification_simulation = sprintf("Simulation of misclassification (probabilistic bias analysis): in each of %s draws, sensitivity and specificity are drawn from beta distributions fitted to the counts in the Lo et al. 2022 participants with accelerometer data and the outcome observed, the walking answers are reclassified at random to match, and the replication's model is refitted; when misclassification differs between groups, people with and without the outcome get their own draws from their own counts. A draw is discarded when it implies %s, or when the model cannot be refitted; a re-estimate with more than %s of its draws discarded, the limit set in advance, is not reliable.",
                                         count(assumption("bias.iterations")), impossible_sim, pct(assumption("bias.max_discard_share")))
))
# Device measures per day: counts without decimals, minutes to 1 decimal
cnum <- function(x, measure) ifelse(measure == "counts_per_day", num(x, 0), num(x, 1))
cnum_ci <- function(est, lo, hi, measure) {
  ifelse(measure == "counts_per_day", oaireport::fmt_ci(est, lo, hi, 0), oaireport::fmt_ci(est, lo, hi, 1))
}
plain_measures <- c(purposeful_min = sprintf("Minutes in bouts of %s+ minutes", purposeful_minutes),
                    counts_per_day = "Activity counts", light_min = "Light-activity minutes")
bold_rule <- "Bold: the 95% CI excludes no difference, no association or no agreement beyond chance (0; or 1 for odds ratios)."
# A figure another document saved (the brief and report.qmd render before this document)
figure_file <- function(name, width = "100%") {
  path <- file.path("figures", paste0(name, ".png"))
  if (!file.exists(path)) {
    stop("clinical.qmd: ", path, " is missing; the document that saves it must render before clinical.qmd ([report] documents)", call. = FALSE)
  }
  c("", "```{=typst}", sprintf('#image("%s", width: %s)', path, width), "```", "")
}
# "(95% CI)" in a table header, kept on one line
ci_header <- function(x) gsub("(95% CI)", "(95%\u00a0CI)", x, fixed = TRUE)
# Sensitivity, specificity, predictive values and Youden's J in one sample (no bold: no null)
class_table <- function(sample) {
  get <- function(m, col = "estimate") pick(sample, m, col)
  rows <- c(se = "Sensitivity", sp = "Specificity", ppv = "Positive predictive value", npv = "Negative predictive value")
  tab <- data.frame(
    Measure = c(unname(rows), "Youden's J"),
    `Estimate (95% CI)` = c(vapply(names(rows), function(m) oaireport::fmt_ci(get(m), get(m, "lo"), get(m, "hi")), character(1)),
                            num(youden(sample))),
    `Based on` = c(vapply(names(rows), function(m) sprintf("%s of %s", count(get(m, "k")), count(get(m, "n"))), character(1)),
                   "sensitivity + specificity − 1"),
    check.names = FALSE)
  names(tab) <- ci_header(names(tab))
  tab
}
lower_first <- function(x) paste0(tolower(substr(x, 1, 1)), substring(x, 2))
and_list <- function(x) if (length(x) <= 1) x else paste(paste(x[-length(x)], collapse = ", "), "and", x[length(x)])
flow <- stats::setNames(run$flow$persons, run$flow$step)
n_val <- run$metrics[["sample.validation.persons"]]
n_walk <- run$metrics[["sample.validation.walkers"]]
n_lo <- run$metrics[["sample.lo_subset.persons"]]
kg <- run$validity_known_groups[run$validity_known_groups$sample == "validation", ]
kg_bout <- kg[kg$measure == "purposeful_min", ]
dz <- run$validity_dose[run$validity_dose$sample == "validation", ]
j_strict <- ci_get("validation", "times", "cut3", "j")
grid_np <- run$bias_tipping[run$bias_tipping$outcome == "new_pain" & !is.na(run$bias_tipping$or), ]
# A grid axis as "floor–top", the top written "1" at perfect accuracy
grid_span <- function(floor, x) paste0(num(floor), "–", if (isTRUE(all.equal(max(x), 1))) "1" else num(max(x)))
# The walker by the question, as the run codes a yes without an amount (report_setup.R's item_def)
walker_plain <- switch(yes_without,
  `non-walker` = , exclude = "answered yes to the 96-month question “walked for exercise since age 50” (at least 20 minutes a day, at least 10 times) and said how much they walk (years, months a year or times a month).",
  walker = "answered yes to the 96-month question “walked for exercise since age 50” (at least 20 minutes a day, at least 10 times), whether or not they said how much they walk.",
  stop("unknown exposure.yes_without_amount_as: ", yes_without, call. = FALSE))
nonwalker_plain <- switch(yes_without,
  `non-walker` = "said no, or said yes without saying how much they walk",
  walker = "said no",
  exclude = "said no (a yes without saying how much they walk is left out)",
  stop("unknown exposure.yes_without_amount_as: ", yes_without, call. = FALSE))
# The device rule and the valid-day rule in plain words, built from the assumptions
bout_plain <- sprintf("%s+ minutes of moderate-to-vigorous activity", purposeful_minutes)
rule_plain <- switch(walker_rule,
  bout_days = sprintf("on average at least %s days a week with a bout of %s", fmt(assumption("reference.min_bout_days_per_week")), bout_plain),
  any_bout = sprintf("at least one bout of %s", bout_plain),
  bout_minutes = sprintf("on average at least %s minutes a week in bouts of %s", fmt(assumption("reference.min_bout_minutes_per_week")), bout_plain),
  stop("unknown reference.walker_rule: ", walker_rule, call. = FALSE))
valid_day_plain <- sprintf("a valid day has at least %s wear hours and a usable visit at least %s valid days",
                           fmt(assumption("device.valid_day_hours")), fmt(assumption("device.min_valid_days")))
```

## Key points

- `r count(n_val)` participants had a walking answer and accelerometer data (48 or 72 months); `r pct(n_walk / n_val)` were walkers.
- A walker by the question typically recorded `r num(kg_bout$hl, 1)` more minutes a day in bouts of `r purposeful_minutes`+ minutes than a non-walker (the median of the differences over all walker–non-walker pairs; 95% CI `r num(kg_bout$hl_lo, 1)`–`r num(kg_bout$hl_hi, 1)`).
- The walking question separates device-defined walkers poorly (sensitivity `r num(pick("validation", "se"))`, specificity `r num(pick("validation", "sp"))`, Youden's J `r num(youden("validation"))`); asking how often helps (J `r num(j_strict)` for walking `r tolower(cut_names("times")[["cut3"]])`).
- If misclassification is the same with and without the outcome, walking's link to less new frequent knee pain in Lo et al. 2022 [@lo2022] holds in `r sum(grid_np$excludes_1)` of `r nrow(grid_np)` possible accuracy settings examined (sensitivity `r grid_span(min(grid_np$se), grid_np$se)`, specificity `r grid_span(min(grid_np$sp), grid_np$sp)`).
- Activity counts are not specific to walking; a step-based device measure would remove that difference between the question and the device, though not the difference in time frame.

```{r words}
#| results: asis
pase_weight <- assumption("pase.walking_scoring")$weight
cat(oaireport::reading_guide(c(
  `Walker (by the question)` = walker_plain,
  `Non-walker (by the question)` = paste0(nonwalker_plain, "."),
  `Device-defined walker` = paste0(rule_plain, "."),
  `Activity counts` = "the accelerometer's measure of movement intensity, summed over the day.",
  `Light-activity minutes` = sprintf("minutes a day between sitting still and moderate activity (%s–%s counts a minute).", fmt(assumption("device.light_floor")), fmt(cutpoint - 1)),
  `PASE walker` = if (pase_threshold == 0) "any walking outside the home in the past 7 days (PASE, the Physical Activity Scale for the Elderly)." else sprintf("a PASE walking score above %s (PASE, the Physical Activity Scale for the Elderly).", fmt(pase_threshold)),
  `PASE walking score` = sprintf("days walked in the past week ÷ 7 × hours a day (answer midpoints) × %s; 0 for no walking.", fmt(pase_weight)),
  `Sensitivity, specificity` = "of the people the device calls walkers, the share the question also calls walkers; of those it calls non-walkers, the share the question also calls non-walkers.",
  `Predictive values` = "of the walkers by the question, the share the device calls walkers (PPV); of the non-walkers, the share it calls non-walkers (NPV).",
  `Youden's J` = "sensitivity + specificity − 1; 0 means no better than chance, 1 means perfect.",
  `Agreement beyond chance (kappa)` = "0 means no better than chance, 1 means perfect agreement.",
  `95% CI` = "the 95% confidence interval, written as 0.89 (0.86–0.91)."),
  notation = paste(bold_rule, "Letters in the text (a, b, …) link to the technical notes at the end."),
  title = "Words used in this appendix"), sep = "\n")
```

```{r who}
oaireport::compare_table(data.frame(
  Part = c("Part 1. Does the walking question agree with the device?", "Part 2. What the agreement means for Lo et al. 2022"),
  Who = c("Answered the walking question and have accelerometer data at 48 or 72 months",
          sprintf("Of those, members of the Lo et al. 2022 cohort (the replication uses the cohort of %s)", count(flow[["lo2022_cohort"]]))),
  Participants = count(c(n_val, n_lo)), check.names = FALSE), widths = c(2.4, 3, 1), multipage = FALSE)
```

```{=typst}
#pagebreak()
```

# Part 1. Does the walking question agree with the device?

This part uses all `r count(n_val)` participants with a walking answer and accelerometer data.

## Who was studied

Table A1 shows how the sample was formed.

```{=typst}
#block(breakable: false)[
```

```{r a-flow}
oaireport::compare_table(data.frame(
  Group = c("Answered the walking question (96 months)", "…with accelerometer data at 48 or 72 months",
            "Walkers by the question", "Non-walkers by the question", "Walkers by the device"),
  Participants = count(c(flow[["answered_walking_item"]], flow[["with_valid_device_wave"]], n_walk, n_val - n_walk,
                         run$metrics[["sample.validation.device_walkers"]])),
  check.names = FALSE), widths = c(4, 1.2), multipage = FALSE)
```

```{r note-a-flow}
#| results: asis
cat(oaireport::table_note(sprintf("Participants of the Osteoarthritis Initiative who answered the walking question at the 96-month visit, and of those the ones with usable accelerometer data at the 48- or 72-month visit (%s). Non-walkers by the question %s; walkers by the device are defined from the accelerometer alone (see Words used in this appendix).", valid_day_plain, nonwalker_plain), label = lab("Table"), close_group = TRUE), sep = "\n")
```

## Do people who say they walk move more?

Walkers by the question recorded more activity than non-walkers on `r sum(!is.na(kg$hl_lo) & kg$hl_lo > 0)` of `r nrow(kg)` device measures (the 95% CI of the difference above 0). Table A2 gives the median difference`r notes$ref("median_difference")` with its 95% CI, and the difference adjusted for age, sex and body-mass index`r notes$ref("median_regression")` with its 95% CI`r notes$ref("adjusted_ci")`; Figure A1 shows the spread.

```{=typst}
#block(breakable: false)[
```

```{r a-known-groups}
kg_table <- data.frame(
  `Device measure, per day` = named(kg$measure, plain_measures),
  walk = cnum(kg$median_walkers, kg$measure),
  dont = cnum(kg$median_nonwalkers, kg$measure),
  diff = cnum_ci(kg$hl, kg$hl_lo, kg$hl_hi, kg$measure),
  `Adjusted difference (95% CI)` = cnum_ci(kg$adj_diff, kg$adj_lo, kg$adj_hi, kg$measure),
  check.names = FALSE)
names(kg_table)[2:4] <- c(sprintf("Walkers, median (n = %s)", count(kg$n_walkers[1])),
                          sprintf("Non-walkers, median (n = %s)", count(kg$n_nonwalkers[1])),
                          "Difference (95% CI)")
kg_bold <- cbind(FALSE, FALSE, FALSE, oaireport::ci_excludes(kg$hl_lo, kg$hl_hi), oaireport::ci_excludes(kg$adj_lo, kg$adj_hi))
names(kg_table) <- ci_header(names(kg_table))
oaireport::compare_table(kg_table, widths = c(2.2, 1.5, 1.5, 1.7, 1.7), multipage = FALSE, bold = kg_bold)
```

```{r note-a-known-groups}
#| results: asis
# A median of 0.0 bout minutes reads as "no one had a bout": say how many in each group had none
bout_groups <- c(nonwalkers = "non-walkers", walkers = "walkers")
bout_none <- unlist(Map(function(g, label) {
  if (kg_bout[[paste0("median_", g)]] == 0) sprintf("at least half of %s", label)
  else if (kg_bout[[paste0("q25_", g)]] == 0) sprintf("at least a quarter of %s", label)
}, names(bout_groups), bout_groups))
bout_zero <- names(bout_groups)[vapply(names(bout_groups), function(g) kg_bout[[paste0("median_", g)]] == 0, logical(1))]
bout_some <- length(bout_zero) > 0 && all(vapply(bout_zero, function(g) kg_bout[[paste0("q75_", g)]] > 0, logical(1)))
bout_text <- if (length(bout_none)) {
  paste0(toupper(substr(paste(bout_none, collapse = " and "), 1, 1)), substring(paste(bout_none, collapse = " and "), 2),
         " recorded no minutes in bouts", if (bout_some) ", so a median of 0.0 does not mean that no one in the group had a bout", ".")
} else ""
cat(oaireport::table_note(paste(sprintf("Validation sample, by the walking question; non-walkers %s.", nonwalker_plain), "Medians are per day of accelerometer wear; the difference is the typical (median) difference between a walker and a non-walker, so it need not equal the gap between the two medians, and the adjusted difference is the difference in medians after allowing for age, sex and body-mass index.", bout_text, bold_rule), label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{r a-known-groups-figure}
#| fig-height: 2.1
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
kgf <- data.frame(
  measure = rep(kg$measure, 2),
  group = rep(c("Walkers", "Non-walkers"), each = nrow(kg)),
  median = c(kg$median_walkers, kg$median_nonwalkers),
  q25 = c(kg$q25_walkers, kg$q25_nonwalkers), q75 = c(kg$q75_walkers, kg$q75_nonwalkers))
kgf$measure <- factor(named(kgf$measure, plain_measures), levels = unname(plain_measures))
fig_kg <- ggplot2::ggplot(kgf, ggplot2::aes(median, group, colour = group)) +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = q25, xmax = q75), size = 0.3) +
  ggplot2::facet_wrap(~measure, scales = "free_x", ncol = 3) +
  ggplot2::scale_x_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)) +
  ggplot2::scale_colour_manual(values = c(Walkers = colours[["item"]], `Non-walkers` = colours[["muted"]]), guide = "none") +
  ggplot2::labs(x = "Median (interquartile range), per day", y = NULL) + oaireport::theme_oai()
oaireport::save_figure(fig_kg, "clinical_known_groups", "2col", height = 2.1)
print(fig_kg)
cat(oaireport::table_note("Validation sample; walkers and non-walkers by the walking question. Point: median; bar: interquartile range (the middle half of participants).", label = lab("Figure"), close_group = TRUE), sep = "\n")
```

## Does more reported walking go with more recorded activity?

```{r dose-sentence}
rises <- !is.na(dz$jt_p) & dz$jt_p < 0.05
dose_names <- lower_first(named(dz$measure, plain_measures))
groups_text <- "from non-walkers to walkers with fewer and then more lifetime walking sessions"
dose_sentence <- if (all(rises)) {
  sprintf("Recorded activity rose %s on every device measure.", groups_text)
} else if (!any(rises)) {
  sprintf("Recorded activity did not clearly rise %s on any device measure.", groups_text)
} else {
  sprintf("Recorded activity rose %s for %s, but not clearly for %s.", groups_text, and_list(dose_names[rises]),
          and_list(dose_names[!rises]))
}
```

`r dose_sentence` Table A3 gives the medians and a test for trend`r notes$ref("trend_test")` across the three groups.

```{=typst}
#block(breakable: false)[
```

```{r a-dose}
dose_table <- data.frame(
  `Device measure, per day` = named(dz$measure, plain_measures),
  none = cnum(dz$median_none, dz$measure), lower = cnum(dz$median_lower, dz$measure),
  upper = cnum(dz$median_upper, dz$measure), `Test for trend, p` = comp_p(dz$jt_p), check.names = FALSE)
names(dose_table)[2:4] <- c(sprintf("Non-walkers (n = %s)", count(dz$n_none[1])),
                            sprintf("Walkers, fewer sessions (n = %s)", count(dz$n_lower[1])),
                            sprintf("Walkers, more sessions (n = %s)", count(dz$n_upper[1])))
dose_bold <- cbind(matrix(FALSE, nrow(dz), 4), rises)
oaireport::compare_table(dose_table, widths = c(2.2, 1.4, 1.4, 1.4, 1.2), multipage = FALSE, bold = dose_bold)
```

```{r note-a-dose}
#| results: asis
cat(oaireport::table_note(sprintf("Validation sample; medians per day; non-walkers by the walking question %s. Walkers are split at the %s of lifetime walking sessions (years × months a year × times a month, from their answers); the %s walkers without a session count (an amount answer missing) are in neither walker group. Bold p: the measure rises across the three groups (p < 0.05).", nonwalker_plain, assumption("exposure.amount_split"), count(dz$n_walkers_no_level[1])), label = lab("Table"), close_group = TRUE), sep = "\n")
```

## How well does “yes” pick out walkers by the device?

Table A4 gives the sensitivity and specificity`r notes$ref("sensitivity_specificity")` of the answer against the device, with 95% CIs`r notes$ref("wilson_ci")`, the predictive values`r notes$ref("predictive_values")`, and Youden's J`r notes$ref("youden_j")`.

```{=typst}
#block(breakable: false)[
```

```{r a-class}
oaireport::compare_table(class_table("validation"), widths = c(2.6, 2, 2.2), multipage = FALSE)
```

```{r note-a-class}
#| results: asis
cat(oaireport::table_note("Validation sample, against the device-defined walker. “Based on”: the count behind each share. Sensitivity, specificity and the predictive values are never bold; Youden's J has no CI here (Table A5 gives J with CIs; its walking-question rows leave out walkers who did not say how many times a month they walk, so its first row differs slightly from this table).", label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{=typst}
#block(breakable: false)[
```

## Does asking how often help?

Defining walkers by how often they walk, rather than by any walking, raises Youden's J for both the walking question and the PASE walking item (Table A5, Figure A2). Table A5 gives each definition's J with its 95% CI`r notes$ref("bootstrap_ci")`.

```{r a-cuts}
freq_rows <- function(get, names_k) do.call(rbind, lapply(names(names_k), function(k) data.frame(
  Definition = names_k[[k]],
  Sensitivity = oaireport::fmt_ci(get(k, "se"), get(k, "se", "lo"), get(k, "se", "hi")),
  Specificity = oaireport::fmt_ci(get(k, "sp"), get(k, "sp", "lo"), get(k, "sp", "hi")),
  `Youden's J (95% CI)` = oaireport::fmt_ci(get(k, "j"), get(k, "j", "lo"), get(k, "j", "hi")),
  bold_j = oaireport::ci_excludes(get(k, "j", "lo"), get(k, "j", "hi")), check.names = FALSE)))
# Participants classified in a block: the device walkers (se's n) and non-walkers (sp's n) at its first cut
classified <- function(get, names_k) get(names(names_k)[1], "se", "n") + get(names(names_k)[1], "sp", "n")
item_names <- cut_names("times", prefix = "Walks at least", top_prefix = "Walks")
pase_names <- cut_names("pase", prefix = "Walked at least", top_prefix = "Walked")
item_get <- function(k, m, col = "estimate") ci_get("validation", "times", k, m, col)
pase_get <- function(v) function(k, m, col = "estimate") cp_get(v, "frequency", "device_walker", k, m, col)
blocks <- c(list(freq_rows(item_get, item_names)), lapply(comp_visits, function(v) freq_rows(pase_get(v), pase_names)))
block_n <- c(classified(item_get, item_names), vapply(comp_visits, function(v) classified(pase_get(v), pase_names), numeric(1)))
titles <- sprintf("%s (n = %s)", c("Walking question: times per month", sprintf("PASE: days walked in the past week, %s", wave(comp_visits))),
                  count_na(block_n))
freq <- do.call(rbind, blocks)
starts <- cumsum(c(1, vapply(blocks, nrow, integer(1))))[seq_along(blocks)]
freq_bold <- cbind(matrix(FALSE, nrow(freq), 3), freq$bold_j)
freq$bold_j <- NULL
names(freq) <- ci_header(names(freq))
oaireport::compare_table(freq, widths = c(2.4, 1.6, 1.6, 1.8), multipage = FALSE, bold = freq_bold,
                         groups = as.list(stats::setNames(starts, titles)))
```

```{r note-a-cuts}
#| results: asis
cat(oaireport::table_note(paste("Each row defines walkers by one answer cut-off and compares them with the device-defined walker: the walking question in the validation sample, PASE at each accelerometer visit (48 or 72 months); n in each block's title is the participants classified.", bold_rule), label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{r a-cuts-figure}
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
cat(figure_file("brief_fig1", width = "58%"), sep = "\n")
cat(oaireport::table_note("The definitions of Table A5, each placed by its sensitivity and 1 − specificity. Further up and to the left is better agreement; grey dashed lines join points of equal Youden's J (value above the top edge; J = 0 is chance). Triangles: the definitions in use (any walking).", label = lab("Figure"), close_group = TRUE), sep = "\n")
```

PASE also asks how many hours a day people walked. Days × hours gives an estimate of weekly walking, which Table A6 compares with the device's minutes in bouts: as a rank correlation`r notes$ref("rank_correlation")`, as the median of the differences between the two, and as agreement beyond chance`r notes$ref("kappa")` on walking `r fmt(assumption("reference.min_bout_minutes_per_week"))`+ minutes a week; the 95% CIs of the last two come from resampling participants`r notes$ref("bootstrap_ci")`. Figure A3 shows the device's minutes within each band of reported weekly walking.

```{=typst}
#block(breakable: false)[
```

```{r a-weekly}
guide <- assumption("reference.min_bout_minutes_per_week")
wk <- do.call(rbind, lapply(comp_visits, function(v) {
  r <- function(level, stat, col = "estimate") cp_get(v, "weekly", "purposeful_week", level, stat, col)
  data.frame(
    Visit = sprintf("%s (n = %s)", wave(v), count_na(r("all", "median_diff", "n"))),
    rho = oaireport::fmt_ci(r("all", "rho"), r("all", "rho", "lo"), r("all", "rho", "hi")),
    diff = oaireport::fmt_ci(r("all", "median_diff"), r("all", "median_diff", "lo"), r("all", "median_diff", "hi"), 0),
    kappa = oaireport::fmt_ci(r("guideline", "kappa"), r("guideline", "kappa", "lo"), r("guideline", "kappa", "hi")),
    b_rho = oaireport::ci_excludes(r("all", "rho", "lo"), r("all", "rho", "hi")),
    b_diff = oaireport::ci_excludes(r("all", "median_diff", "lo"), r("all", "median_diff", "hi")),
    b_kappa = oaireport::ci_excludes(r("guideline", "kappa", "lo"), r("guideline", "kappa", "hi")),
    check.names = FALSE)
}))
wk_bold <- cbind(FALSE, wk$b_rho, wk$b_diff, wk$b_kappa)
wk <- wk[c("Visit", "rho", "diff", "kappa")]
names(wk)[2:4] <- c("Rank correlation (95% CI)", "Reported − device, median minutes a week (95% CI)", sprintf("Agreement on %s+ minutes a week, kappa (95%% CI)", fmt(guide)))
names(wk) <- ci_header(names(wk))
oaireport::compare_table(wk, widths = c(1.5, 1.8, 2.1, 2.1), multipage = FALSE, bold = wk_bold)
```

```{r note-a-weekly}
#| results: asis
cat(oaireport::table_note(paste("Participants with a PASE walking answer and accelerometer data at the same visit (n beside each visit). Reported weekly walking: days × hours a day from PASE; device: minutes a week in bouts of", purposeful_minutes, "or more minutes; reported weekly minutes take few distinct values, so the median difference moves in steps and can sit at a limit of its own 95% CI.", bold_rule), label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{r a-calibration}
#| fig-height: 2.6
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
cal <- cp[cp$component == "weekly" & cp$comparator == "purposeful_week" & cp$statistic == "median", ]
cal$level <- factor(cal$level, levels = unique(cal$level))
cal$visit_name <- wave(cal$visit)
fig_cal <- ggplot2::ggplot(cal, ggplot2::aes(level, estimate)) +
  ggplot2::geom_pointrange(ggplot2::aes(ymin = lo, ymax = hi), colour = colours[["pase"]], size = 0.3) +
  ggplot2::facet_wrap(~visit_name) +
  ggplot2::labs(x = "Weekly walking reported on PASE (hours a week)",
                y = sprintf("Device minutes a week\nin bouts of %s+ minutes", purposeful_minutes)) +
  oaireport::theme_oai()
oaireport::save_figure(fig_cal, "clinical_calibration", "2col", height = 2.6)
print(fig_cal)
cal_n <- tapply(cal$n, cal$visit, sum)
cat(oaireport::table_note(sprintf("Participants with a PASE walking answer and accelerometer data at the same visit (%s). Point: median; bar: interquartile range of the device's minutes a week, within each band of reported weekly walking. A report that tracked the device would show medians rising from left to right.", and_list(sprintf("n = %s at %s", count(cal_n), wave(names(cal_n))))), label = lab("Figure"), close_group = TRUE), sep = "\n")
```

## How does PASE compare?

Table A7 compares the PASE walker with the device-defined walker at each PASE visit.

```{=typst}
#block(breakable: false)[
```

```{r a-pase}
pz <- run$validity_pase
pz$visit <- sprintf("%02d", as.integer(pz$visit))
pv <- function(v, measure, stat, col = "estimate") {
  x <- pz[pz$visit == v & pz$measure == measure & pz$statistic == stat, col]
  if (length(x) == 1) x else NA
}
pase_visits <- unique(pz$visit)
pt <- do.call(rbind, lapply(pase_visits, function(v) data.frame(
  Visit = sprintf("%s (n = %s)", wave(v), count_na(pv(v, "device_walker", "se", "n") + pv(v, "device_walker", "sp", "n"))),
  Sensitivity = oaireport::fmt_ci(pv(v, "device_walker", "se"), pv(v, "device_walker", "se", "lo"), pv(v, "device_walker", "se", "hi")),
  Specificity = oaireport::fmt_ci(pv(v, "device_walker", "sp"), pv(v, "device_walker", "sp", "lo"), pv(v, "device_walker", "sp", "hi")),
  `Youden's J` = num(pv(v, "device_walker", "se") + pv(v, "device_walker", "sp") - 1),
  rho = oaireport::fmt_ci(pv(v, "purposeful_min", "rho"), pv(v, "purposeful_min", "rho", "lo"), pv(v, "purposeful_min", "rho", "hi")),
  b_rho = oaireport::ci_excludes(pv(v, "purposeful_min", "rho", "lo"), pv(v, "purposeful_min", "rho", "hi")),
  check.names = FALSE)))
pt_bold <- cbind(matrix(FALSE, nrow(pt), 4), pt$b_rho)
pt$b_rho <- NULL
names(pt)[5] <- "Rank correlation with device bout minutes (95% CI)"
rho_n <- vapply(pase_visits, function(v) pv(v, "purposeful_min", "rho", "n"), numeric(1))
names(pt) <- ci_header(names(pt))
oaireport::compare_table(pt, widths = c(1.4, 1.6, 1.6, 1.05, 2), multipage = FALSE, bold = pt_bold)
```

```{r note-a-pase}
#| results: asis
cat(oaireport::table_note(paste(sprintf("The PASE walker against the device-defined walker at each PASE visit (n: participants with a PASE answer and accelerometer data); the rank correlation, of the PASE walking score with the device's minutes in bouts, is among PASE walkers only (%s).", and_list(sprintf("n = %s at %s", count_na(rho_n), wave(pase_visits)))), "The 96-month visit has no accelerometer data of its own and is compared with the participant's accelerometer visits.", bold_rule), label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{=typst}
#pagebreak()
```

# Part 2. What the agreement means for Lo et al. 2022

```{r b-numbers}
n_lo_cohort <- flow[["lo2022_cohort"]]
j_val <- youden("validation"); j_lo <- youden("lo_subset")
j_compare <- if (abs(j_lo - j_val) < 0.05) "about as weak here as" else if (j_lo < j_val) "weaker here than" else "stronger here than"
st <- run$validity_strata[run$validity_strata$sample == "lo_subset" & run$validity_strata$measure %in% c("se", "sp"), ]
strata_differ <- any(!is.na(st$p_differs) & st$p_differs < 0.05)
strata_text <- if (strata_differ) "differ across some subgroups" else "do not clearly differ across subgroups"
# Outcome names inside a sentence: "new frequent knee pain", but "KL grade worsening" kept
lower_name <- function(x) ifelse(grepl("^[A-Z][a-z]", x), lower_first(x), x)
lo_t2 <- run$lo2022_table2
cc_knees <- lo_t2$n[match(paste(names(outcome_names), "or_adj"), paste(lo_t2$outcome, lo_t2$model))]  # complete-case knees
cc_text <- sprintf("each outcome's model uses the knees with that outcome and the model's covariates observed (complete cases: %s knees)",
                   and_list(count(cc_knees)))
```

The agreement results here use the `r count(n_lo)` participants of Lo et al. 2022 [@lo2022] who answered the walking question and have accelerometer data, a subset of the Part 1 sample. The replication and the sensitivity analyses use the replicated cohort of `r count(n_lo_cohort)` participants, in which each outcome's model uses the knees with that outcome and the model's covariates observed. The worked example and the simulation of misclassification take the question's accuracy from the participants with accelerometer data; the tipping-point grid uses assumed accuracy values.

## Agreement in the Lo et al. 2022 participants

Table A8 gives the sensitivity and specificity`r notes$ref("sensitivity_specificity")` of the answer against the device in these participants, with 95% CIs`r notes$ref("wilson_ci")`, the predictive values`r notes$ref("predictive_values")` and Youden's J`r notes$ref("youden_j")`. Agreement is `r j_compare` in Part 1 (J `r num(j_lo)`, against `r num(j_val)`), and sensitivity and specificity `r strata_text``r notes$ref("subgroup_test")` (Figure A4).

```{=typst}
#block(breakable: false)[
```

```{r b-class}
oaireport::compare_table(class_table("lo_subset"), widths = c(2.6, 2, 2.2), multipage = FALSE)
```

```{r note-b-class}
#| results: asis
cat(oaireport::table_note("Lo et al. 2022 participants with accelerometer data, against the device-defined walker. “Based on”: the count behind each share. Sensitivity, specificity and the predictive values are never bold; Youden's J has no CI here.", label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{r b-strata}
#| fig-height: 4
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
var_names <- c(kl = "KL grade, worse knee", pain = "Frequent knee pain", outcome_names)
vars <- unique(st$variable)
group_n <- vapply(vars, function(v) sum(st$n[st$variable == v]), numeric(1))  # se n + sp n over its strata
group_label <- stats::setNames(sprintf("%s (n\u00a0=\u00a0%s)", named(vars, var_names), count(group_n)), vars)
st$group <- factor(group_label[st$variable], levels = group_label)
st$stratum <- gsub("-", "–", st$stratum)
st$measure_name <- factor(ifelse(st$measure == "se", "Sensitivity", "Specificity"), levels = c("Sensitivity", "Specificity"))
fig_st <- ggplot2::ggplot(st, ggplot2::aes(estimate, stratum, colour = measure_name)) +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = lo, xmax = hi), position = ggplot2::position_dodge(width = 0.5), size = 0.25) +
  ggplot2::facet_grid(group ~ ., scales = "free_y", space = "free_y", labeller = ggplot2::label_wrap_gen(18)) +
  ggplot2::scale_colour_manual(values = c(Sensitivity = colours[["item"]], Specificity = colours[["navy"]]), name = NULL) +
  ggplot2::scale_x_continuous(limits = c(0, 1)) +
  ggplot2::labs(x = "Estimate (95% CI)", y = NULL) + oaireport::theme_oai() +
  ggplot2::theme(legend.position = "bottom", strip.text.y = ggplot2::element_text(angle = 0, hjust = 0))
oaireport::save_figure(fig_st, "clinical_strata", "2col", height = 4)
print(fig_st)
cat(oaireport::table_note(sprintf("Lo et al. 2022 participants with accelerometer data (n = %s; n in each panel: those with the subgroup observed): sensitivity and specificity of the walking answer against the device-defined walker, within subgroups. Subgroups: the Kellgren–Lawrence (KL) grade of the worse knee and frequent knee pain in either knee at baseline; for each outcome, those of the %s with the outcome observed, by whether it occurred in either knee during follow-up (medial JSN: medial joint-space narrowing).", count(n_lo), count(n_lo)), label = lab("Figure"), close_group = TRUE), sep = "\n")
```

## Two ways to read low agreement

Low agreement between the question and the accelerometer has two readings, and the data here cannot tell them apart.

- **Reporting.** Some people may misremember or misreport their walking, so the question misclassifies them.
- **Different constructs.** The question asks about a long-standing habit of walking for exercise since age 50, at the 96-month visit. The accelerometer recorded one week of current activity of any kind, years earlier. And activity counts do not single out walking. The two may simply measure different things.

A walking-specific accelerometer measure, such as steps and step rate, would narrow these readings: it removes one difference (activity of any kind against walking), though not the difference in time frame.

## The results of Lo et al. 2022, replicated

Table A9 sets the adjusted odds ratios of this analysis's replication beside the published ones.

```{=typst}
#block(breakable: false)[
```

```{r b-lo}
t2 <- run$lo2022_t2
rows <- t2[match(sprintf("t2.%s.or_adj", names(outcome_names)), t2$metric), ]
pub <- oaireport::parse_or(rows$published)
ours <- oaireport::parse_or(rows$ours)
lo_table <- data.frame(
  Outcome = unname(outcome_names),
  `Published odds ratio (95% CI)` = sub(" \\*$", "", gsub("-", "–", rows$published)),
  ours = oaireport::fmt_ci(ours$or, ours$lo, ours$hi),
  check.names = FALSE)
names(lo_table)[3] <- sprintf("This analysis, n\u00a0=\u00a0%s (95%%\u00a0CI)", count(n_lo_cohort))
# The published column is bold where the paper reports significance (its "*"): a limit rounded to 1.0
# can hide an interval that excludes 1
pub_sig <- !is.na(pub$sig) & pub$sig
lo_bold <- cbind(FALSE, pub_sig, oaireport::ci_excludes(ours$lo, ours$hi, 1))
rounds_to_one <- pub_sig & !oaireport::ci_excludes(pub$lo, pub$hi, 1)
rounds_text <- if (any(rounds_to_one)) {
  sprintf("; its %s %s to 1.0", and_list(sub(" worsening$", "", lower_name(outcome_names[rounds_to_one]))),
          if (sum(rounds_to_one) == 1) "interval rounds" else "intervals round")
} else ""
names(lo_table) <- ci_header(names(lo_table))
oaireport::compare_table(lo_table, widths = c(2.6, 2, 2), multipage = FALSE, bold = lo_bold)
```

```{r note-b-lo}
#| results: asis
cat(oaireport::table_note(sprintf("Adjusted odds ratios for walkers against non-walkers, by knee, from Lo et al. 2022 and from this analysis's replication with the published methods, in which %s. Published column: bold where the paper reports the result as significant%s. This analysis: bold where the 95%% CI excludes 1.", cc_text, rounds_text), label = lab("Table"), close_group = TRUE), sep = "\n")
```

## How sensitive is the association to classification?

```{r b-example}
t2v <- function(metric) as.numeric(t2$published[t2$metric == metric])
w_e <- t2v("t2.new_pain.walkers.events"); w_n <- t2v("t2.new_pain.walkers.n")
nw_e <- t2v("t2.new_pain.nonwalkers.events"); nw_n <- t2v("t2.new_pain.nonwalkers.n")
case_share <- w_e / (w_e + nw_e)
non_share <- (w_n - w_e) / ((w_n - w_e) + (nw_n - nw_e))
gap <- abs(case_share - non_share)
# The example: the feasible grid cell nearest sensitivity 0.9 and specificity 0.6, so it is a cell of Figure A5
ex <- grid_np[which.min(abs(grid_np$se - 0.9) + abs(grid_np$sp - 0.6)), ]
ex_se <- ex$se; ex_sp <- ex$sp; ex_j <- ex_se + ex_sp - 1
# The accuracy of the definition in use, as Table A11's first row (and the brief) give it
dev <- function(stat) ci_get("lo_subset", "times", "cut1", stat)
dev_se <- dev("se"); dev_sp <- dev("sp"); dev_j <- dev("j")
case_dev <- (case_share - (1 - dev_sp)) / dev_j
# Table A10's columns, named in the text below as in the table
sim_cols <- c(nondiff = "Misclassification the same in both groups", diff = "Misclassification differs between groups")
```

If the walking question misclassifies some people, how would the association in Lo et al. 2022 change? In the paper's published table, `r pct(case_share, 1)` of the knees that developed new frequent knee pain belonged to walkers, against `r pct(non_share, 1)` of the knees that did not: a gap of `r num(100 * gap, 1)` percentage points. If misclassification is the same in people with and without the outcome`r notes$ref("nondifferential")`, allowing for it divides that gap by Youden's J. If the question's sensitivity were `r num(ex_se)` and its specificity `r num(ex_sp)` (J `r num(ex_j)`), the gap would grow to `r num(100 * gap / ex_j, 1)` points, so the association looks stronger, not weaker. Against the accelerometer, sensitivity is `r num(dev_se)` and specificity `r num(dev_sp)` (J `r num(dev_j)`): the gap would be multiplied by `r num(1 / dev_j, 1)`, and the share of walkers among knees with new pain would become `r pct(case_dev)`, below zero, which is impossible.

The tipping-point analysis`r notes$ref("tipping_point")` repeats this over a grid of accuracy settings, refitting the adjusted model of Lo et al. 2022 in each (Figure A5). The grid assumes misclassification is the same in people with and without the outcome, and allowing for that kind can only move the odds ratio away from 1; so the grid cannot show whether an association survives misclassification that differs by group. The simulation below also tries that case (Table A10, column “`r sim_cols[["diff"]]`”).

```{r b-tipping}
#| fig-height: 4.6
#| results: asis
tp <- run$bias_tipping
tip <- do.call(rbind, lapply(names(outcome_names), function(o) {
  g <- tp[tp$outcome == o & !is.na(tp$or), ]
  perfect <- g$excludes_1[abs(g$se - 1) < 1e-9 & abs(g$sp - 1) < 1e-9]  # the observed association
  data.frame(outcome = outcome_names[[o]], holds = sum(g$excludes_1, na.rm = TRUE), feasible = nrow(g),
             perfect = length(perfect) == 1 && isTRUE(perfect))
}))
# "Holds" only where the interval excludes 1 at perfect accuracy too
tip_names <- lower_name(tip$outcome)
tip_counts <- sprintf("%d of %d", tip$holds, tip$feasible)
partial <- !tip$perfect & tip$holds > 0
never <- !tip$perfect & tip$holds == 0
tip_text <- paste0(cap1(paste(c(
  if (any(tip$perfect)) sprintf("for %s the association holds in %s possible settings", and_list(tip_names[tip$perfect]), and_list(tip_counts[tip$perfect])),
  if (any(partial)) sprintf("for %s the interval excludes 1 only under assumed misclassification (%s possible settings), and at perfect accuracy it includes 1",
                            and_list(tip_names[partial]), and_list(tip_counts[partial])),
  if (any(never)) sprintf("for %s the interval includes 1 in every possible setting", and_list(tip_names[never]))),
  collapse = "; ")), ".")
# A float: at the foot of this page if it fits, else at the top of the next, with the text after it filling this page
cat("", "```{=typst}", "#place(auto, float: true, clearance: 1.5em)[", "#set align(start)", "```", "", sep = "\n")
cat(oaireport::figure_group(), sep = "\n")
# report.qmd's tipping figure, drawn again with plain words in its key
tp$outcome <- factor(named(tp$outcome, outcome_names), levels = outcome_names)
tp$label <- ifelse(is.na(tp$or), "–", num(tp$or))
tp$state <- factor(ifelse(is.na(tp$or) | is.na(tp$excludes_1), "impossible", ifelse(tp$excludes_1, "excludes", "includes")),
                   levels = c("excludes", "includes", "impossible"))
fig_tip <- ggplot2::ggplot(tp, ggplot2::aes(x = factor(num(sp)), y = factor(num(se)), fill = state)) +
  ggplot2::geom_tile(colour = "white", linewidth = 0.4) +
  ggplot2::geom_text(ggplot2::aes(label = label), size = 1.9, family = oaireport::oai_font(), colour = "grey10") +
  ggplot2::scale_fill_manual(values = c(excludes = "#cde2fb", includes = "grey92", impossible = "white"),
                             labels = c(excludes = "95% interval excludes 1", includes = "95% interval includes 1",
                                        impossible = "Impossible combination"),
                             drop = FALSE, name = NULL) +
  ggplot2::guides(fill = ggplot2::guide_legend(override.aes = list(colour = "grey60"))) +
  ggplot2::facet_wrap(~outcome, ncol = 2) +
  ggplot2::labs(x = "Specificity", y = "Sensitivity") + oaireport::theme_oai() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5))
oaireport::save_figure(fig_tip, "clinical_tipping", "2col", height = 4.6)
print(fig_tip)
cat(oaireport::table_note(sprintf("Replicated Lo et al. 2022 cohort (n = %s; for each outcome, the complete cases of Table A9). Each cell: the adjusted odds ratio if the walking question had that sensitivity (rows) and specificity (columns), the same in people with and without the outcome; shaded cells keep a 95%% interval that excludes 1, and “–” marks an impossible combination. %s",
  count(n_lo_cohort), tip_text), label = lab("Figure"), close_group = TRUE), sep = "\n")
cat("", "```{=typst}", "]", "```", "", sep = "\n")
```

```{r b-simulation-text}
pba <- run$bias_pba
adj <- pba[pba$model == "or_adj", ]
limit <- pct(assumption("bias.max_discard_share"))
sim_text <- if (all(adj$flagged)) {
  sprintf("cannot give a reliable single re-estimate: for every outcome, %s of its simulations imply impossible values, more than the limit of %s set in advance (Table A10). That reflects how weakly the question and activity counts agree; activity counts are not specific to walking.",
          pct_range(adj$discarded), limit)
} else if (!any(adj$flagged)) {
  sprintf("gives a usable re-estimate: at most %s of its simulations imply impossible values, within the limit of %s set in advance (Table A10).",
          pct(max(adj$discarded)), limit)
} else {
  sprintf("gives a reliable re-estimate in only %d of %d combinations of outcome and assumption: in the others, more than %s of its simulations, the limit set in advance, imply impossible values (Table A10).",
          sum(!adj$flagged), nrow(adj), limit)
}
```

A simulation of misclassification`r notes$ref("misclassification_simulation")` that takes the question's accuracy from the accelerometer `r sim_text`

```{=typst}
#block(breakable: false)[
```

```{r b-simulation}
share <- function(o, scenario) pct(adj$discarded[adj$outcome == o & adj$scenario == scenario])
sim <- data.frame(
  Outcome = unname(outcome_names),
  nondiff = vapply(names(outcome_names), share, character(1), scenario = "non-differential"),
  diff = vapply(names(outcome_names), share, character(1), scenario = "differential"))
names(sim) <- c(sprintf("Outcome (n\u00a0=\u00a0%s)", count(n_lo_cohort)), sim_cols[["nondiff"]], sim_cols[["diff"]])
priors <- run$bias_priors[run$bias_priors$scenario == "non-differential", ]
prior_n <- (priors$se_n + priors$sp_n)[match(names(outcome_names), priors$outcome)]  # accuracy from these
oaireport::compare_table(sim, widths = c(2.6, 2, 2), multipage = FALSE)
```

```{r note-b-simulation}
#| results: asis
cat(oaireport::table_note(sprintf("Share of the %s simulations of the adjusted model of Lo et al. 2022, in the replicated cohort (n = %s; for each outcome, the complete cases of Table A9), that imply an impossible value, for example a negative number of true walkers among people with the outcome. The walking question's accuracy is taken from the accelerometer in those of the %s participants with accelerometer data who have the outcome observed (%s, by outcome); groups: people with and without the outcome, who share one accuracy or each have their own. The full results are in the technical report.",
  count(adj$n[1]), count(n_lo_cohort), count(n_lo), and_list(count(prior_n))), label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{r b-stricter}
item_names <- cut_names("times", prefix = "Walks at least", top_prefix = "Walks")
defs <- c(cut1 = "Any walker (in use)", cut2 = item_names[["cut2"]], cut3 = item_names[["cut3"]])
st_rows <- do.call(rbind, lapply(names(defs), function(k) {
  g <- function(stat, col = "estimate") ci_get("lo_subset", "times", k, stat, col)
  data.frame(
    Definition = defs[[k]],
    Sensitivity = oaireport::fmt_ci(g("se"), g("se", "lo"), g("se", "hi")),
    Specificity = oaireport::fmt_ci(g("sp"), g("sp", "lo"), g("sp", "hi")),
    `Youden's J (95% CI)` = oaireport::fmt_ci(g("j"), g("j", "lo"), g("j", "hi")),
    `Multiplier 1/J (95% CI)` = ifelse(is.na(g("correction")), "–",
      sprintf("%s (%s–%s)", num(g("correction"), 1), num_na(g("correction", "lo"), 1),
              ifelse(is.na(g("correction", "hi")), "∞", num(g("correction", "hi"), 1)))),
    b_j = oaireport::ci_excludes(g("j", "lo"), g("j", "hi")),
    j = g("j"), n = g("j", "n"),
    check.names = FALSE)
}))
steadier <- all(diff(st_rows$j) > 0)
st_n <- st_rows$n[1]
st_rows$Definition <- gsub("/", "\u2060/\u2060", st_rows$Definition)
names(st_rows)[1] <- sprintf("Definition (n\u00a0=\u00a0%s)", count_na(st_n))
names(st_rows) <- ci_header(names(st_rows))
st_bold <- cbind(matrix(FALSE, nrow(st_rows), 3), st_rows$b_j, FALSE)
st_rows$b_j <- NULL; st_rows$j <- NULL; st_rows$n <- NULL
no_times <- ci_get("lo_subset", "times", "no_band", "count")  # walkers who did not say how many times a month they walk
```

`r if (steadier) "A stricter definition makes a re-estimate steadier, because it raises J: the multiplier 1/J falls (Table A11)." else "Stricter definitions do not raise J steadily here, so the multiplier 1/J does not fall steadily (Table A11)."` Table A11 gives J with its 95% CI`r notes$ref("bootstrap_ci")` for each definition.

```{=typst}
#block(breakable: false)[
```

```{r b-stricter-table}
oaireport::compare_table(st_rows, widths = c(2.2, 1.6, 1.6, 1.6, 1.6), multipage = FALSE, bold = st_bold)
```

```{r note-b-stricter}
#| results: asis
cat(oaireport::table_note(paste(sprintf("Lo et al. 2022 participants with accelerometer data, walkers defined by the times-per-month answer (the %s walkers who did not say how many times a month they walk are left out). The multiplier shows how much a re-estimate that allows for misclassification would multiply the gap in walking between people with and without the outcome; its interval comes from J's.", count(no_times)), bold_rule), label = lab("Table"), close_group = TRUE), sep = "\n")
```

## An accelerometry-based alternative

The accelerometry group's step and step-rate data make a walking-specific exposure possible:

- steps per day;
- minutes a day at 100 or more steps per minute (about moderate intensity in adults aged 61 to 85 [@tudorlocke2021]);
- bouts of continuous stepping.

Defined together and pre-specified, such an exposure could be put into the models of Lo et al. 2022 in place of the walking question, and the results compared with the question's. That would show whether, and how much, the association depends on how walking is measured.

```{=typst}
#pagebreak()
```

# Methods in brief

```{r methods}
# The device processing in plain words, from the assumptions (report_setup.R's device_detail and rule_short use technical terms)
across_visits <- if (combination == "mean") "averaged over the person's usable visits" else
  sprintf("at the %s visit", sub(" months", "-month", wave(combination)))
week_plain <- switch(walker_rule,
  bout_days = sprintf(" Days a week with a bout: the share of a person's first %s valid days with one, × 7, %s.", fmt(assumption("device.max_valid_days")), across_visits),
  bout_minutes = sprintf(" Minutes a week in bouts: minutes in bouts per valid day × 7, %s.", across_visits),
  any_bout = "",
  stop("unknown reference.walker_rule: ", walker_rule, call. = FALSE))
device_plain <- sprintf("A bout starts when %s of %s consecutive minutes reach %s counts a minute (moderate-to-vigorous activity) and ends when %s minutes in the %s-minute window fall below that; bouts of %s or more minutes count.%s",
                        fmt(assumption("device.bout_need")), fmt(assumption("device.bout_window")), fmt(cutpoint),
                        fmt(assumption("device.bout_stop_below")), fmt(assumption("device.bout_window")), purposeful_minutes, week_plain)
```

- **Participants.** Part 1: Osteoarthritis Initiative participants who answered the walking question at the 96-month visit and had usable accelerometer data at the `r sub(" or ", "- or ", sub(" months", "-month", waves_at))` visit (`r valid_day_plain`). Part 2: the `r count(n_lo)` of them in the Lo et al. 2022 cohort; the replication uses the replicated cohort of `r count(n_lo_cohort)` participants, each outcome's model the knees with that outcome and the model's covariates observed.
- **Accelerometer.** An ActiGraph GT1M worn at the hip for a week [@song2010]. `r device_plain` A device-defined walker shows `r rule_plain`.
- **Statistics.** Medians and median differences with 95% CIs, and differences adjusted for age, sex and body-mass index. Shares (sensitivity, specificity, predictive values) carry 95% CIs. Rank correlations, agreement beyond chance (kappa), and tests for trend across ordered groups. For Lo et al. 2022: the published models replicated, then refitted under assumed accuracy values (tipping-point grid) and under simulated misclassification.
- **Source.** Every number comes from the same analysis run as the brief. The full technical report gives every analysis, setting and assumption, and is available on request.

```{r notes-list}
#| results: asis
cat(notes$list(), sep = "\n")
```
