---
title: "Walking question and accelerometry in the OAI: results"
subtitle: "Appendix to the brief: every result in plain language. The full technical report is available on request."
mainfont: Arial
fontsize: 10pt
brand:
  typography:
    headings:
      color: "#0b2e59"
bibliography: references.bib
bibliographystyle: american-medical-association
format:
  typst:
    papersize: us-letter
    margin:
      x: 0.85in
      y: 0.75in
    include-in-header:
      text: |
        #set text(hyphenate: false, overhang: false)
        #set bibliography(title: [References])
        #set page(footer: context align(center, text(size: 8pt, fill: luma(90))[Appendix — page #counter(page).display()]))
        // The AMA style ends every title with a full stop, which doubles after a title ending in "?"
        #show bibliography: it => { show regex("\\?\\."): "?"; it }
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: cairo_pdf
    fig-width: 7
---

```{=typst}
#set par(justify: false)
```

```{r setup}
source("report_setup.R")
lab <- oaireport::numberer("A")
colours <- oaireport::brief_colours
reps <- fmt(assumption("validity.bootstrap_reps"))  # every bootstrap in validity.R and components.R
adjusted_ci <- if (assumption("validity.median_regression_se") == "boot") {
  sprintf("CI for the adjusted difference: the estimate ± 1.96 × its standard error, the standard error from refitting the model in %s bootstrap resamples of participants.", reps)
} else {
  "CI for the adjusted difference: the estimate ± 1.96 × its standard error, the standard error from the median-regression sandwich formula (quantreg's \"nid\" method)."
}
notes <- oaireport::tech_notes(c(
  median_difference = "Median difference: the Hodges–Lehmann estimate, the median of the differences across all pairs of people, one from each group; it is robust to skewed data and need not equal the gap between the two group medians. Its 95% CI is the Hodges–Lehmann interval from the Wilcoxon rank-sum test.",
  adjusted_ci = adjusted_ci,
  bootstrap_ci = sprintf("Bootstrap 95%% CI: participants are resampled with replacement %s times, the estimate is recomputed each time, and the middle 95%% of the results forms the interval (a percentile interval).", reps),
  trend_test = sprintf("Test for trend: the Jonckheere–Terpstra test of whether a measure rises across ordered groups; its p-value comes from %s random permutations of the group labels.", count(jt_b))
))
# Device measures per day: counts without decimals, minutes to 1 decimal
cnum <- function(x, measure) ifelse(measure == "counts_per_day", num(x, 0), num(x, 1))
cnum_ci <- function(est, lo, hi, measure) {
  ifelse(measure == "counts_per_day", oaireport::fmt_ci(est, lo, hi, 0), oaireport::fmt_ci(est, lo, hi, 1))
}
plain_measures <- c(purposeful_min = sprintf("Minutes in bouts of %s+ minutes", purposeful_minutes),
                    counts_per_day = "Activity counts", light_min = "Light-activity minutes")
bold_rule <- "Bold: the 95% CI excludes no difference, no association or no agreement beyond chance (0; or 1 for odds ratios)."
# A figure another document saved (the brief and report.qmd render before this document)
figure_file <- function(name, width = "100%") {
  path <- file.path("figures", paste0(name, ".png"))
  if (!file.exists(path)) {
    stop("clinical.qmd: ", path, " is missing; the document that saves it must render before clinical.qmd ([report] documents)", call. = FALSE)
  }
  c("", "```{=typst}", sprintf('#image("%s", width: %s)', path, width), "```", "")
}
# Sensitivity, specificity, predictive values and Youden's J in one sample (no bold: no null)
class_table <- function(sample) {
  get <- function(m, col = "estimate") pick(sample, m, col)
  rows <- c(se = "Sensitivity", sp = "Specificity", ppv = "Positive predictive value", npv = "Negative predictive value")
  data.frame(
    Measure = c(unname(rows), "Youden's J"),
    `Estimate (95% CI)` = c(vapply(names(rows), function(m) oaireport::fmt_ci(get(m), get(m, "lo"), get(m, "hi")), character(1)),
                            num(youden(sample))),
    `Based on` = c(vapply(names(rows), function(m) sprintf("%s of %s", count(get(m, "k")), count(get(m, "n"))), character(1)),
                   "sensitivity + specificity − 1"),
    check.names = FALSE)
}
lower_first <- function(x) paste0(tolower(substr(x, 1, 1)), substring(x, 2))
and_list <- function(x) if (length(x) <= 1) x else paste(paste(x[-length(x)], collapse = ", "), "and", x[length(x)])
flow <- stats::setNames(run$flow$persons, run$flow$step)
n_val <- run$metrics[["sample.validation.persons"]]
n_walk <- run$metrics[["sample.validation.walkers"]]
n_lo <- run$metrics[["sample.lo_subset.persons"]]
kg <- run$validity_known_groups[run$validity_known_groups$sample == "validation", ]
kg_bout <- kg[kg$measure == "purposeful_min", ]
dz <- run$validity_dose[run$validity_dose$sample == "validation", ]
j_strict <- ci_get("validation", "times", "cut3", "j")
grid_np <- run$bias_tipping[run$bias_tipping$outcome == "new_pain" & !is.na(run$bias_tipping$or), ]
# The walker by the question, as the run codes a yes without an amount (report_setup.R's item_def)
walker_plain <- switch(yes_without,
  `non-walker` = , exclude = "answered yes to the 96-month question “walked for exercise since age 50” (at least 20 minutes a day, at least 10 times) and said how much they walk (years, months a year or times a month).",
  walker = "answered yes to the 96-month question “walked for exercise since age 50” (at least 20 minutes a day, at least 10 times), whether or not they said how much they walk.",
  stop("unknown exposure.yes_without_amount_as: ", yes_without, call. = FALSE))
nonwalker_plain <- switch(yes_without,
  `non-walker` = "said no, or said yes without saying how much they walk",
  walker = "said no",
  exclude = "said no (a yes without saying how much they walk is left out)",
  stop("unknown exposure.yes_without_amount_as: ", yes_without, call. = FALSE))
# The device rule and the valid-day rule in plain words, built from the assumptions
bout_plain <- sprintf("%s+ minutes of moderate-to-vigorous activity", purposeful_minutes)
rule_plain <- switch(walker_rule,
  bout_days = sprintf("on average at least %s days a week with a bout of %s", fmt(assumption("reference.min_bout_days_per_week")), bout_plain),
  any_bout = sprintf("at least one bout of %s", bout_plain),
  bout_minutes = sprintf("on average at least %s minutes a week in bouts of %s", fmt(assumption("reference.min_bout_minutes_per_week")), bout_plain),
  stop("unknown reference.walker_rule: ", walker_rule, call. = FALSE))
valid_day_plain <- sprintf("a valid day has at least %s wear hours and a usable visit at least %s valid days",
                           fmt(assumption("device.valid_day_hours")), fmt(assumption("device.min_valid_days")))
```

## Key points

- `r count(n_val)` participants had a walking answer and accelerometer data (48 or 72 months); `r pct(n_walk / n_val)` were walkers.
- A walker by the question typically recorded `r num(kg_bout$hl, 1)` more minutes a day in bouts of `r purposeful_minutes`+ minutes than a non-walker (the median of the differences over all walker–non-walker pairs; 95% CI `r num(kg_bout$hl_lo, 1)`–`r num(kg_bout$hl_hi, 1)`).
- The walking question separates device-defined walkers poorly (sensitivity `r num(pick("validation", "se"))`, specificity `r num(pick("validation", "sp"))`, Youden's J `r num(youden("validation"))`); asking how often helps (J `r num(j_strict)` for walking `r tolower(cut_names("times")[["cut3"]])`).
- In Lo et al. 2022's participants [@lo2022], if misclassification is the same with and without the outcome, walking's link to less new frequent knee pain holds in `r sum(grid_np$excludes_1)` of `r nrow(grid_np)` possible accuracy settings (sensitivity `r num(min(grid_np$se))` or more).
- Activity counts are not specific to walking; a step-based device measure would show whether the question and the device measure the same thing.

```{r words}
#| results: asis
pase_weight <- assumption("pase.walking_scoring")$weight
cat(oaireport::reading_guide(c(
  `Walker (by the question)` = walker_plain,
  `Non-walker (by the question)` = paste0(nonwalker_plain, "."),
  `Device-defined walker` = paste0(rule_plain, "."),
  `Activity counts` = "the accelerometer's measure of movement intensity, summed over the day.",
  `Light-activity minutes` = sprintf("minutes a day between sitting still and moderate activity (%s–%s counts a minute).", fmt(assumption("device.light_floor")), fmt(cutpoint - 1)),
  `PASE walker` = if (pase_threshold == 0) "any walking outside the home in the past 7 days (PASE, the Physical Activity Scale for the Elderly)." else sprintf("a PASE walking score above %s (PASE, the Physical Activity Scale for the Elderly).", fmt(pase_threshold)),
  `PASE walking score` = sprintf("days walked in the past week ÷ 7 × hours a day (answer midpoints) × %s; 0 for no walking.", fmt(pase_weight)),
  `Sensitivity, specificity` = "of the people the device calls walkers, the share the question also calls walkers; of those it calls non-walkers, the share the question also calls non-walkers.",
  `Predictive values` = "of the walkers by the question, the share the device calls walkers (PPV); of the non-walkers, the share it calls non-walkers (NPV).",
  `Youden's J` = "sensitivity + specificity − 1; 0 means no better than chance, 1 means perfect.",
  `Agreement beyond chance (kappa)` = "0 means no better than chance, 1 means perfect agreement.",
  `95% CI` = "the 95% confidence interval, written as 0.89 (0.86–0.91)."),
  notation = paste(bold_rule, "Letters in the text (a, b, …) link to the technical notes at the end."),
  title = "Words used in this appendix"), sep = "\n")
```

```{r who}
oaireport::compare_table(data.frame(
  Part = c("Part 1. Does the walking question agree with the device?", "Part 2. What the agreement means for Lo et al. 2022"),
  Who = c("Answered the walking question and have accelerometer data at 48 or 72 months",
          "Of those, members of the Lo et al. 2022 cohort"),
  Participants = count(c(n_val, n_lo)), check.names = FALSE), widths = c(2.4, 3, 1), multipage = FALSE)
```

```{=typst}
#pagebreak()
```

# Part 1. Does the walking question agree with the device?

This part uses all `r count(n_val)` participants with a walking answer and accelerometer data.

## Who was studied

Table A1 shows how the sample was formed.

```{=typst}
#block(breakable: false)[
```

```{r a-flow}
oaireport::compare_table(data.frame(
  Group = c("Answered the walking question (96 months)", "…with accelerometer data at 48 or 72 months",
            "Walkers by the question", "Non-walkers by the question", "Walkers by the device"),
  Participants = count(c(flow[["answered_walking_item"]], flow[["with_valid_device_wave"]], n_walk, n_val - n_walk,
                         run$metrics[["sample.validation.device_walkers"]])),
  check.names = FALSE), widths = c(4, 1.2), multipage = FALSE)
```

```{r note-a-flow}
#| results: asis
cat(oaireport::table_note(sprintf("Participants of the Osteoarthritis Initiative who answered the walking question at the 96-month visit, and of those the ones with usable accelerometer data at the 48- or 72-month visit (%s). Non-walkers by the question %s; walkers by the device are defined from the accelerometer alone (see Words used in this appendix).", valid_day_plain, nonwalker_plain), label = lab("Table"), close_group = TRUE), sep = "\n")
```

## Do people who say they walk move more?

Walkers by the question recorded more activity than non-walkers on `r sum(!is.na(kg$hl_lo) & kg$hl_lo > 0)` of `r nrow(kg)` device measures (the 95% CI of the difference above 0). Table A2 gives the median difference`r notes$ref("median_difference")` with its 95% CI, and the difference adjusted for age, sex and body-mass index`r notes$ref("median_regression")` with its 95% CI`r notes$ref("adjusted_ci")`; Figure A1 shows the spread.

```{=typst}
#block(breakable: false)[
```

```{r a-known-groups}
kg_table <- data.frame(
  `Device measure, per day` = named(kg$measure, plain_measures),
  walk = cnum(kg$median_walkers, kg$measure),
  dont = cnum(kg$median_nonwalkers, kg$measure),
  diff = cnum_ci(kg$hl, kg$hl_lo, kg$hl_hi, kg$measure),
  `Adjusted difference (95% CI)` = cnum_ci(kg$adj_diff, kg$adj_lo, kg$adj_hi, kg$measure),
  check.names = FALSE)
names(kg_table)[2:4] <- c(sprintf("Walkers, median (n = %s)", count(kg$n_walkers[1])),
                          sprintf("Non-walkers, median (n = %s)", count(kg$n_nonwalkers[1])),
                          "Difference (95% CI)")
kg_bold <- cbind(FALSE, FALSE, FALSE, oaireport::ci_excludes(kg$hl_lo, kg$hl_hi), oaireport::ci_excludes(kg$adj_lo, kg$adj_hi))
oaireport::compare_table(kg_table, widths = c(2.2, 1.5, 1.5, 1.7, 1.7), multipage = FALSE, bold = kg_bold)
```

```{r note-a-known-groups}
#| results: asis
cat(oaireport::table_note(paste(sprintf("Validation sample, by the walking question; non-walkers %s.", nonwalker_plain), "Medians are per day of accelerometer wear; the difference is the typical (median) difference between a walker and a non-walker, so it need not equal the gap between the two medians, and the adjusted difference is the difference in medians after allowing for age, sex and body-mass index.", bold_rule), label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{r a-known-groups-figure}
#| fig-height: 2.1
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
kgf <- data.frame(
  measure = rep(kg$measure, 2),
  group = rep(c("Walkers", "Non-walkers"), each = nrow(kg)),
  median = c(kg$median_walkers, kg$median_nonwalkers),
  q25 = c(kg$q25_walkers, kg$q25_nonwalkers), q75 = c(kg$q75_walkers, kg$q75_nonwalkers))
kgf$measure <- factor(named(kgf$measure, plain_measures), levels = unname(plain_measures))
fig_kg <- ggplot2::ggplot(kgf, ggplot2::aes(median, group, colour = group)) +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = q25, xmax = q75), size = 0.3) +
  ggplot2::facet_wrap(~measure, scales = "free_x", ncol = 3) +
  ggplot2::scale_x_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)) +
  ggplot2::scale_colour_manual(values = c(Walkers = colours[["item"]], `Non-walkers` = colours[["muted"]]), guide = "none") +
  ggplot2::labs(x = "Median (interquartile range), per day", y = NULL) + oaireport::theme_oai()
oaireport::save_figure(fig_kg, "clinical_known_groups", "2col", height = 2.1)
print(fig_kg)
cat(oaireport::table_note("Validation sample; walkers and non-walkers by the walking question. Point: median; bar: interquartile range (the middle half of participants).", label = lab("Figure"), close_group = TRUE), sep = "\n")
```

## Does more reported walking go with more recorded activity?

```{r dose-sentence}
rises <- !is.na(dz$jt_p) & dz$jt_p < 0.05
dose_names <- lower_first(named(dz$measure, plain_measures))
groups_text <- "from non-walkers to walkers with fewer and then more lifetime walking sessions"
dose_sentence <- if (all(rises)) {
  sprintf("Recorded activity rose %s on every device measure.", groups_text)
} else if (!any(rises)) {
  sprintf("Recorded activity did not clearly rise %s on any device measure.", groups_text)
} else {
  sprintf("Recorded activity rose %s for %s, but not clearly for %s.", groups_text, and_list(dose_names[rises]),
          and_list(dose_names[!rises]))
}
```

`r dose_sentence` Table A3 gives the medians and a test for trend`r notes$ref("trend_test")` across the three groups.

```{=typst}
#block(breakable: false)[
```

```{r a-dose}
dose_table <- data.frame(
  `Device measure, per day` = named(dz$measure, plain_measures),
  none = cnum(dz$median_none, dz$measure), lower = cnum(dz$median_lower, dz$measure),
  upper = cnum(dz$median_upper, dz$measure), `Test for trend, p` = comp_p(dz$jt_p), check.names = FALSE)
names(dose_table)[2:4] <- c(sprintf("Non-walkers (n = %s)", count(dz$n_none[1])),
                            sprintf("Walkers, fewer sessions (n = %s)", count(dz$n_lower[1])),
                            sprintf("Walkers, more sessions (n = %s)", count(dz$n_upper[1])))
dose_bold <- cbind(matrix(FALSE, nrow(dz), 4), rises)
oaireport::compare_table(dose_table, widths = c(2.2, 1.4, 1.4, 1.4, 1.2), multipage = FALSE, bold = dose_bold)
```

```{r note-a-dose}
#| results: asis
cat(oaireport::table_note(sprintf("Validation sample; medians per day; non-walkers by the walking question %s. Walkers are split at the %s of lifetime walking sessions (years × months a year × times a month, from their answers); the %s walkers without a session count (an amount answer missing) are in neither walker group. Bold p: the measure rises across the three groups (p < 0.05).", nonwalker_plain, assumption("exposure.amount_split"), count(dz$n_walkers_no_level[1])), label = lab("Table"), close_group = TRUE), sep = "\n")
```

## How well does “yes” pick out walkers by the device?

Table A4 gives the sensitivity and specificity`r notes$ref("sensitivity_specificity")` of the answer against the device, with 95% CIs`r notes$ref("wilson_ci")`, the predictive values`r notes$ref("predictive_values")`, and Youden's J`r notes$ref("youden_j")`.

```{=typst}
#block(breakable: false)[
```

```{r a-class}
oaireport::compare_table(class_table("validation"), widths = c(2.6, 2, 2.2), multipage = FALSE)
```

```{r note-a-class}
#| results: asis
cat(oaireport::table_note("Validation sample, against the device-defined walker. “Based on”: the count behind each share. Sensitivity, specificity and the predictive values are never bold; Youden's J has no CI here (Table A5 gives J with CIs; its walking-question rows leave out walkers who did not say how many times a month they walk, so its first row differs slightly from this table).", label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{=typst}
#block(breakable: false)[
```

## Does asking how often help?

Defining walkers by how often they walk, rather than by any walking, raises Youden's J for both the walking question and the PASE walking item (Table A5, Figure A2). Table A5 gives each definition's J with its 95% CI`r notes$ref("bootstrap_ci")`.

```{r a-cuts}
freq_rows <- function(get, names_k) do.call(rbind, lapply(names(names_k), function(k) data.frame(
  Definition = names_k[[k]],
  Sensitivity = oaireport::fmt_ci(get(k, "se"), get(k, "se", "lo"), get(k, "se", "hi")),
  Specificity = oaireport::fmt_ci(get(k, "sp"), get(k, "sp", "lo"), get(k, "sp", "hi")),
  `Youden's J (95% CI)` = oaireport::fmt_ci(get(k, "j"), get(k, "j", "lo"), get(k, "j", "hi")),
  bold_j = oaireport::ci_excludes(get(k, "j", "lo"), get(k, "j", "hi")), check.names = FALSE)))
# Participants classified in a block: the device walkers (se's n) and non-walkers (sp's n) at its first cut
classified <- function(get, names_k) get(names(names_k)[1], "se", "n") + get(names(names_k)[1], "sp", "n")
item_names <- cut_names("times", prefix = "Walks at least", top_prefix = "Walks")
pase_names <- cut_names("pase", prefix = "Walked at least", top_prefix = "Walked")
item_get <- function(k, m, col = "estimate") ci_get("validation", "times", k, m, col)
pase_get <- function(v) function(k, m, col = "estimate") cp_get(v, "frequency", "device_walker", k, m, col)
blocks <- c(list(freq_rows(item_get, item_names)), lapply(comp_visits, function(v) freq_rows(pase_get(v), pase_names)))
block_n <- c(classified(item_get, item_names), vapply(comp_visits, function(v) classified(pase_get(v), pase_names), numeric(1)))
titles <- sprintf("%s (n = %s)", c("Walking question: times per month", sprintf("PASE: days walked in the past week, %s", wave(comp_visits))),
                  count_na(block_n))
freq <- do.call(rbind, blocks)
starts <- cumsum(c(1, vapply(blocks, nrow, integer(1))))[seq_along(blocks)]
freq_bold <- cbind(matrix(FALSE, nrow(freq), 3), freq$bold_j)
freq$bold_j <- NULL
oaireport::compare_table(freq, widths = c(2.4, 1.6, 1.6, 1.8), multipage = FALSE, bold = freq_bold,
                         groups = as.list(stats::setNames(starts, titles)))
```

```{r note-a-cuts}
#| results: asis
cat(oaireport::table_note(paste("Each row defines walkers by one answer cut-off and compares them with the device-defined walker: the walking question in the validation sample, PASE at each accelerometer visit (48 or 72 months); n in each block's title is the participants classified.", bold_rule), label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{r a-cuts-figure}
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
cat(figure_file("brief_fig1", width = "58%"), sep = "\n")
cat(oaireport::table_note("The definitions of Table A5, each placed by its sensitivity and 1 − specificity. Further up and to the left is better agreement; grey dashed lines join points of equal Youden's J (value above the top edge; J = 0 is chance). Triangles: the definitions in use (any walking).", label = lab("Figure"), close_group = TRUE), sep = "\n")
```

PASE also asks how many hours a day people walked. Days × hours gives an estimate of weekly walking, which Table A6 compares with the device's minutes in bouts: as a rank correlation`r notes$ref("rank_correlation")`, as the median of the differences between the two, and as agreement beyond chance`r notes$ref("kappa")` on walking `r fmt(assumption("reference.min_bout_minutes_per_week"))`+ minutes a week; the 95% CIs of the last two come from resampling participants`r notes$ref("bootstrap_ci")`. Figure A3 shows the device's minutes within each band of reported weekly walking.

```{=typst}
#block(breakable: false)[
```

```{r a-weekly}
guide <- assumption("reference.min_bout_minutes_per_week")
wk <- do.call(rbind, lapply(comp_visits, function(v) {
  r <- function(level, stat, col = "estimate") cp_get(v, "weekly", "purposeful_week", level, stat, col)
  data.frame(
    Visit = sprintf("%s (n = %s)", wave(v), count_na(r("all", "median_diff", "n"))),
    rho = oaireport::fmt_ci(r("all", "rho"), r("all", "rho", "lo"), r("all", "rho", "hi")),
    diff = oaireport::fmt_ci(r("all", "median_diff"), r("all", "median_diff", "lo"), r("all", "median_diff", "hi"), 0),
    kappa = oaireport::fmt_ci(r("guideline", "kappa"), r("guideline", "kappa", "lo"), r("guideline", "kappa", "hi")),
    b_rho = oaireport::ci_excludes(r("all", "rho", "lo"), r("all", "rho", "hi")),
    b_diff = oaireport::ci_excludes(r("all", "median_diff", "lo"), r("all", "median_diff", "hi")),
    b_kappa = oaireport::ci_excludes(r("guideline", "kappa", "lo"), r("guideline", "kappa", "hi")),
    check.names = FALSE)
}))
wk_bold <- cbind(FALSE, wk$b_rho, wk$b_diff, wk$b_kappa)
wk <- wk[c("Visit", "rho", "diff", "kappa")]
names(wk)[2:4] <- c("Rank correlation (95% CI)", "Reported − device, median minutes a week (95% CI)", sprintf("Agreement on %s+ minutes a week, kappa (95%% CI)", fmt(guide)))
oaireport::compare_table(wk, widths = c(1.5, 1.8, 2.1, 2.1), multipage = FALSE, bold = wk_bold)
```

```{r note-a-weekly}
#| results: asis
cat(oaireport::table_note(paste("Participants with a PASE walking answer and accelerometer data at the same visit (n beside each visit). Reported weekly walking: days × hours a day from PASE; device: minutes a week in bouts of", purposeful_minutes, "or more minutes; reported weekly minutes take few distinct values, so the median difference moves in steps and can sit at a limit of its own 95% CI.", bold_rule), label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{r a-calibration}
#| fig-height: 2.6
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
cal <- cp[cp$component == "weekly" & cp$comparator == "purposeful_week" & cp$statistic == "median", ]
cal$level <- factor(cal$level, levels = unique(cal$level))
cal$visit_name <- wave(cal$visit)
fig_cal <- ggplot2::ggplot(cal, ggplot2::aes(level, estimate)) +
  ggplot2::geom_pointrange(ggplot2::aes(ymin = lo, ymax = hi), colour = colours[["pase"]], size = 0.3) +
  ggplot2::facet_wrap(~visit_name) +
  ggplot2::labs(x = "Weekly walking reported on PASE (hours a week)",
                y = sprintf("Device minutes a week\nin bouts of %s+ minutes", purposeful_minutes)) +
  oaireport::theme_oai()
oaireport::save_figure(fig_cal, "clinical_calibration", "2col", height = 2.6)
print(fig_cal)
cal_n <- tapply(cal$n, cal$visit, sum)
cat(oaireport::table_note(sprintf("Participants with a PASE walking answer and accelerometer data at the same visit (%s). Point: median; bar: interquartile range of the device's minutes a week, within each band of reported weekly walking. A report that tracked the device would show medians rising from left to right.", and_list(sprintf("n = %s at %s", count(cal_n), wave(names(cal_n))))), label = lab("Figure"), close_group = TRUE), sep = "\n")
```

## How does PASE compare?

Table A7 compares the PASE walker with the device-defined walker at each PASE visit.

```{=typst}
#block(breakable: false)[
```

```{r a-pase}
pz <- run$validity_pase
pz$visit <- sprintf("%02d", as.integer(pz$visit))
pv <- function(v, measure, stat, col = "estimate") {
  x <- pz[pz$visit == v & pz$measure == measure & pz$statistic == stat, col]
  if (length(x) == 1) x else NA
}
pase_visits <- unique(pz$visit)
pt <- do.call(rbind, lapply(pase_visits, function(v) data.frame(
  Visit = sprintf("%s (n = %s)", wave(v), count_na(pv(v, "device_walker", "se", "n") + pv(v, "device_walker", "sp", "n"))),
  Sensitivity = oaireport::fmt_ci(pv(v, "device_walker", "se"), pv(v, "device_walker", "se", "lo"), pv(v, "device_walker", "se", "hi")),
  Specificity = oaireport::fmt_ci(pv(v, "device_walker", "sp"), pv(v, "device_walker", "sp", "lo"), pv(v, "device_walker", "sp", "hi")),
  `Youden's J` = num(pv(v, "device_walker", "se") + pv(v, "device_walker", "sp") - 1),
  rho = oaireport::fmt_ci(pv(v, "purposeful_min", "rho"), pv(v, "purposeful_min", "rho", "lo"), pv(v, "purposeful_min", "rho", "hi")),
  b_rho = oaireport::ci_excludes(pv(v, "purposeful_min", "rho", "lo"), pv(v, "purposeful_min", "rho", "hi")),
  check.names = FALSE)))
pt_bold <- cbind(matrix(FALSE, nrow(pt), 4), pt$b_rho)
pt$b_rho <- NULL
names(pt)[5] <- "Rank correlation with device bout minutes (95% CI)"
rho_n <- vapply(pase_visits, function(v) pv(v, "purposeful_min", "rho", "n"), numeric(1))
oaireport::compare_table(pt, widths = c(1.4, 1.6, 1.6, 1.05, 2), multipage = FALSE, bold = pt_bold)
```

```{r note-a-pase}
#| results: asis
cat(oaireport::table_note(paste(sprintf("The PASE walker against the device-defined walker at each PASE visit (n: participants with a PASE answer and accelerometer data); the rank correlation, of the PASE walking score with the device's minutes in bouts, is among PASE walkers only (%s).", and_list(sprintf("n = %s at %s", count_na(rho_n), wave(pase_visits)))), "The 96-month visit has no accelerometer data of its own and is compared with the participant's accelerometer visits.", bold_rule), label = lab("Table"), close_group = TRUE), sep = "\n")
```

```{r notes-list}
#| results: asis
cat(notes$list(), sep = "\n")
```
