---
title: "Walking for exercise: validity of the OAI item against accelerometry"
subtitle: "The Lo et al. 2022 exposure, with a probabilistic bias analysis"
date: today
mainfont: Arial
fontsize: 10pt
# Navy title, subtitle and section headings, as in the brief
brand:
  typography:
    headings:
      color: "#0b2e59"
format:
  typst:
    papersize: us-letter
    margin:
      x: 0.9in
      y: 0.9in
    include-in-header:
      text: |
        #set text(hyphenate: false, overhang: false)
        #set page(footer: context align(center, text(size: 8pt, fill: luma(90))[Technical report — page #counter(page).display()]))
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: cairo_pdf
    fig-width: 7
---

```{r setup}
run <- oaireport::load_results()$default
num <- function(x, digits = 2) formatC(x, format = "f", digits = digits, big.mark = ",")
count <- function(x) formatC(x, format = "d", big.mark = ",")
pct <- function(x, digits = 0) paste0(num(100 * x, digits), "%")
pct_range <- function(x) {
  lo <- num(100 * min(x), 0)
  hi <- num(100 * max(x), 0)
  if (lo == hi) sprintf("%s%%", lo) else sprintf("%s–%s%%", lo, hi)
}
named <- function(key, map) ifelse(key %in% names(map), map[key], key)
# Odds ratios: "–" when every draw was discarded, "<0.01" for limits that collapse toward 0
or_text <- function(x) ifelse(is.na(x), "–", ifelse(x < 0.01, "<0.01", num(x)))
or_interval <- function(or, lo, hi) {
  ifelse(is.na(or), "–", sprintf("%s (%s–%s)", or_text(or), or_text(lo), or_text(hi)))
}
flag_mark <- function(flagged) ifelse(flagged, "✓", "")
# Device measures: counts/day without decimals, minutes/day to 2 decimals
dnum <- function(x, measure) ifelse(measure == "counts_per_day", num(x, 0), num(x))
# An estimate with its 95% interval, in the device measure's digits (counts/day without decimals)
dnum_ci <- function(est, lo, hi, measure) {
  ifelse(measure == "counts_per_day", oaireport::fmt_ci(est, lo, hi, 0), oaireport::fmt_ci(est, lo, hi, 2))
}
# Tables and figures are numbered A1, A2, ... (the report is the brief's appendix)
lab <- oaireport::numberer("A")
class_names <- c(se = "Se", sp = "Sp", ppv = "PPV", npv = "NPV", auc = "AUC")  # as in the legends
measure_names <- c(purposeful_min = "Purposeful-bout min/day", counts_per_day = "Counts/day",
                   light_min = "Light min/day")
outcome_names <- c(new_pain = "New frequent knee pain", kl_worse = "KL grade worsening",
                   jsn_worse = "Medial JSN worsening", improved_pain = "Resolution of frequent knee pain")
model_names <- c(or_unadj = "Unadjusted", or_adj = "Adjusted")
sample_names <- c(validation = "Validation", lo_subset = "Lo 2022 subset")
wave_names <- c(`06` = "48 months", `08` = "72 months", `10` = "96 months")
wave <- function(x) named(sprintf("%02d", as.integer(x)), wave_names)
assumption <- function(key) run$assumptions$assumptions[[key]]$value
threshold <- assumption("bias.max_discard_share")
cls <- run$validity_classification
pick <- function(sample, measure, col = "estimate") cls[[col]][cls$sample == sample & cls$measure == measure]
youden <- function(sample) pick(sample, "se") + pick(sample, "sp") - 1
# Share answering yes: device walkers answering yes (se_k) plus device non-walkers answering yes (sp_n - sp_k)
yes_share <- (pick("lo_subset", "se", "k") + pick("lo_subset", "sp", "n") - pick("lo_subset", "sp", "k")) /
  (pick("lo_subset", "se", "n") + pick("lo_subset", "sp", "n"))
# The same share in each outcome group (case / noncase) of the Lo subset with device data, from the
# differential prior counts; a tipping cell is infeasible when Se is below the higher group's share
group_priors <- run$bias_priors[run$bias_priors$scenario == "differential", ]
group_share <- (group_priors$se_k + group_priors$sp_n - group_priors$sp_k) / (group_priors$se_n + group_priors$sp_n)
higher_share <- tapply(group_share, group_priors$outcome, max)
pba <- run$bias_pba
nondiff <- pba$scenario == "non-differential"
flag_text <- if (all(pba$flagged)) "every estimate" else sprintf("%d of %d estimates", sum(pba$flagged), nrow(pba))
which_ors <- if (all(pba$flagged)) "its" else "those"
# The instability story is told only when the run has flagged estimates
any_flagged <- any(pba$flagged)
flagged_points_text <- if (!any_flagged) {
  "No bias-adjusted estimate is flagged."
} else if (all(pba$flagged)) {
  "Every bias-adjusted estimate is flagged, so these points are shown for completeness, not interpretation."
} else {
  sprintf("%d of %d bias-adjusted estimates are flagged; the flagged ones are shown for completeness, not interpretation.",
          sum(pba$flagged), nrow(pba))
}
discarded_text <- sprintf("%s of draws in the non-differential scenario and %s in the differential one",
                          pct_range(pba$discarded[nondiff]), pct_range(pba$discarded[!nondiff]))
instability_text <- if (any_flagged) {
  sprintf("The probabilistic bias analysis was unstable, discarding %s; %s exceeded the pre-specified %s discard threshold, so %s corrected odds ratios are not interpretable and the tipping-point grid is the primary sensitivity analysis.",
          discarded_text, flag_text, pct(threshold), which_ors)
} else {
  sprintf("The probabilistic bias analysis discarded %s; no estimate exceeded the pre-specified %s discard threshold.",
          discarded_text, pct(threshold))
}
tp <- run$bias_tipping
grid_np_all <- tp[tp$outcome == "new_pain", ]
grid_np <- grid_np_all[!is.na(grid_np_all$or), ]
grid_np_excludes <- grid_np[grid_np$excludes_1, ]
all_exclude <- nrow(grid_np_excludes) == nrow(grid_np)
lowest_sum <- min(grid_np$se + grid_np$sp)  # lowest feasible Se + Sp
# The feasible cell(s) at that corner: the Se and Sp to name, not the extremes of each axis
corner <- grid_np[abs(grid_np$se + grid_np$sp - lowest_sum) < 1e-9, ]
corner_cells <- paste(sprintf("Se %s and Sp %s", num(corner$se), num(corner$sp)), collapse = "; ")
corner_text <- if (nrow(corner) > 1) {
  sprintf("the feasible cells with the lowest Se + Sp are %s (each a sum of %s)", corner_cells, num(lowest_sum))
} else {
  sprintf("the feasible cell with the lowest Se + Sp has %s (a sum of %s)", corner_cells, num(lowest_sum))
}
headline_diff <- pba[pba$outcome == "new_pain" & pba$model == "or_adj" & pba$scenario == "differential", ]
# Differential rows of both bias analyses (record-level PBA and summary-level correction)
sl_all <- run$bias_summary_level
sl_all_flagged <- if ("flagged" %in% names(sl_all)) sl_all$flagged else sl_all$discarded > threshold
pba_diff_flagged <- pba$flagged[!nondiff]
sl_diff_flagged <- sl_all_flagged[sl_all$scenario == "differential"]
differential_text <- if (all(pba_diff_flagged) && all(sl_diff_flagged)) {
  "The differential analyses below, the record-level probabilistic bias analysis and the summary-level correction, are both flagged"
} else {
  sprintf("The differential analyses below are the record-level probabilistic bias analysis (%d of %d rows flagged) and the summary-level correction (%d of %d rows flagged)",
          sum(pba_diff_flagged), length(pba_diff_flagged), sum(sl_diff_flagged), length(sl_diff_flagged))
}
# The grid is non-differential, so it cannot speak to differential error (Summary clause)
diff_flagged <- c(pba_diff_flagged, sl_diff_flagged)
differential_clause <- if (all(diff_flagged)) {
  "both differential bias analyses (record-level and summary-level) are flagged as unstable"
} else if (!any(diff_flagged)) {
  "neither differential bias analysis (record-level or summary-level) is flagged as unstable"
} else {
  sprintf("the differential bias analyses are flagged as unstable in %d of %d rows", sum(diff_flagged), length(diff_flagged))
}
grid_sentence <- sprintf(
  "Under non-differential misclassification, the tipping-point grid for new frequent knee pain (adjusted) keeps an interval that excludes 1 in %s feasible cells (the lowest feasible Se + Sp is %s), but non-differential error biases toward the null, so the grid cannot show whether the association survives differential error, and %s.",
  if (all_exclude) sprintf("all %d", nrow(grid_np)) else sprintf("%d of %d", nrow(grid_np_excludes), nrow(grid_np)),
  num(lowest_sum), differential_clause)
info <- run$run_info
code_version <- paste0("oai-analytics ", info$oai_version,
                       if (is.null(info$git_commit)) "" else paste0(", commit ", substr(info$git_commit, 1, 7)),
                       if (isTRUE(info$git_dirty)) " (with uncommitted changes)" else "")
versions <- unlist(info$data_versions)
data_release <- if (length(versions)) paste(sprintf("%s %s", names(versions), versions), collapse = ", ") else "not recorded"
```

```{r legends}
# Legends under the tables and figures, in plain words. Every number comes from the run's
# assumptions or results, so a variant run is described as it ran.
fmt <- function(x) format(x, trim = TRUE, big.mark = ",")
cap1 <- function(x) paste0(toupper(substr(x, 1, 1)), substring(x, 2))
combination <- assumption("device.wave_combination")
waves_at <- if (combination == "mean") "48 or 72 months" else wave(combination)
averaged <- if (combination == "mean") "averaged over a person's valid waves" else
  sprintf("at the %s wave", sub(" months", "-month", wave(combination)))
cutpoint <- assumption("device.mv_cutpoint")
purposeful_minutes <- fmt(assumption("device.purposeful_bout_minutes"))
valid_day <- sprintf("a valid day has at least %s wear hours and a valid wave at least %s valid days",
                     fmt(assumption("device.valid_day_hours")), fmt(assumption("device.min_valid_days")))

# The walking item, defined in full once, in the first legend that uses it
yes_without <- assumption("exposure.yes_without_amount_as")
item_def <- switch(yes_without,
  `non-walker` = "Walker: answered yes to the 96-month question on walking for exercise (at least 20 minutes a day, at least 10 times, at age 50 or older; V10WLKAR4) and gave at least one amount (years, months per year or times per month); non-walker: answered no, or yes with no amount.",
  walker = "Walker: answered yes to the 96-month question on walking for exercise (at least 20 minutes a day, at least 10 times, at age 50 or older; V10WLKAR4), with or without an amount; non-walker: answered no.",
  exclude = "Walker: answered yes to the 96-month question on walking for exercise (at least 20 minutes a day, at least 10 times, at age 50 or older; V10WLKAR4) and gave at least one amount (years, months per year or times per month); non-walker: answered no (a yes with no amount is left out).",
  stop("unknown exposure.yes_without_amount_as: ", yes_without))
# The walkers with no lifetime-sessions score (an amount item is missing); a "yes" with no
# amount at all is one of them only under walker coding
no_amount_group <- if (yes_without == "walker") {
  "walkers who gave no amount or only some of the amount items"
} else "walkers who answered some but not all of the amount items"

# The device reference standard, defined in full once (Classification)
walker_rule <- assumption("reference.walker_rule")
rule_short <- switch(walker_rule,
  bout_days = sprintf("on average at least %s days a week with a purposeful bout", fmt(assumption("reference.min_bout_days_per_week"))),
  any_bout = "at least one purposeful bout",
  bout_minutes = sprintf("on average at least %s minutes a week in purposeful bouts", fmt(assumption("reference.min_bout_minutes_per_week"))),
  stop("unknown reference.walker_rule: ", walker_rule))
device_def <- sprintf("Device walker (the reference standard): defined from the accelerometer alone (ActiGraph GT1M, %s), as a person who shows %s; a purposeful bout is at least %s minutes of moderate-to-vigorous activity.",
                      waves_at, rule_short, purposeful_minutes)
per_week <- switch(walker_rule,
  bout_days = sprintf("Days a week = the share of a person's first %s valid days with a bout × 7, %s. ", fmt(assumption("device.max_valid_days")), averaged),
  bout_minutes = sprintf("Minutes a week = purposeful-bout minutes per valid day × 7, %s. ", averaged),
  "")
device_detail <- sprintf("A bout starts when %s of %s minutes reach %s counts/min and ends when %s minutes in the window fall below it. %s%s.",
                         fmt(assumption("device.bout_need")), fmt(assumption("device.bout_window")), fmt(cutpoint),
                         fmt(assumption("device.bout_stop_below")), per_week, cap1(valid_day))
measures_def <- sprintf("Device measures are averages per valid day: purposeful-bout min/day (minutes inside purposeful bouts), counts/day (total activity counts) and light min/day (minutes at %s–%s counts/min).",
                        fmt(assumption("device.light_floor")), fmt(cutpoint - 1))

# PASE
pase_threshold <- assumption("pase.walker_threshold")
pase_def <- if (pase_threshold == 0) {
  "PASE (Physical Activity Scale for the Elderly) walker: a PASE walking subscore above 0, i.e. any walking outside the home in the past 7 days (PASE item 2), at that visit."
} else {
  sprintf("PASE (Physical Activity Scale for the Elderly) walker: a PASE walking subscore above %s (PASE item 2, walking outside the home in the past 7 days) at that visit.", fmt(pase_threshold))
}

# Models and tests
reps_text <- if (assumption("validity.median_regression_se") == "boot") {
  sprintf("bootstrap 95%% CI, %s resamples", fmt(assumption("validity.bootstrap_reps")))
} else "95% CI from sandwich standard errors"
adjusted_def <- sprintf("median regression adjusting for age, sex and body-mass index (BMI) (%s)", reps_text)

legend <- list()
legend$flow <- sprintf("Validation sample (first two rows): participants who answered the walking item, and of those the ones with a valid device wave at %s (%s). Lo 2022 cohort (last three rows): the cohort of the replication, then its members who answered the item, then those who also have a valid device wave, the Lo 2022 subset.",
                       waves_at, valid_day)
legend$reproduction <- sprintf("Release days: person-days in the release's by-day accelerometry file; matched: those also among our first %s valid days. Wear, MV and bout differs: matched days on which our wear minutes (by half a minute or more), moderate-to-vigorous (MV) minutes or MV-bout minutes differ from the release's. Valid persons: participants with at least %s valid days, ours / the release's own count.",
                               fmt(assumption("device.max_valid_days")), fmt(assumption("device.min_valid_days")))
legend$known_groups_figure <- sprintf("Point: median; bar: interquartile range, by walking-item status, in the validation sample. %s %s", item_def, measures_def)
legend$known_groups <- sprintf("Validation sample and Lo 2022 subset; walkers and device measures as defined in the figure legend above. Hodges–Lehmann: median of all walker − non-walker differences (95%% confidence interval, CI); rank-biserial: effect size from −1 to 1, positive when walkers are higher; adjusted median difference: walkers minus non-walkers from %s.",
                               adjusted_def)
legend$convergent <- "Walkers with a known amount (n), in the validation sample and the Lo 2022 subset; walkers, lifetime sessions and device measures as defined under Known groups and dose–response. Spearman ρ: rank correlation (95% CI) between lifetime sessions and the device measure; between-wave ρ: Spearman correlation of that measure between 48 and 72 months in participants valid at both waves, estimated in the validation sample and used for both samples; reliability: that of each walker's device mean over their one or two valid waves, from the between-wave ρ by Spearman–Brown, given the share of these walkers with two valid waves; deattenuated ρ = ρ\u00a0/\u00a0√reliability (capped at 1), which corrects for the device measure's unreliability only."
legend$classification <- sprintf("Validation sample and Lo 2022 subset; walkers by the item as defined under Known groups. Se (sensitivity): share of device walkers who are walkers by the item; Sp (specificity): share of device non-walkers who are non-walkers by the item; PPV and NPV (positive and negative predictive value): share of item walkers and of item non-walkers who are device walkers and non-walkers; AUC (area under the receiver operating characteristic curve): chance that a device walker has more lifetime sessions than a device non-walker (item non-walkers score 0, and %s are left out); k / n: count over its denominator. %s %s",
                                 no_amount_group, device_def, device_detail)
legend$strata <- "Se and Sp (95% CI) against the device walker (as defined under Classification), within subgroups of the validation sample and the Lo 2022 subset: KL 0–1, KL 2 and KL 3–4 are the baseline Kellgren–Lawrence (KL) grade of the worse knee; frequent pain and no frequent pain refer to either knee at baseline; event and no event (Lo 2022 subset only) mean the outcome occurred in either knee during follow-up. The four outcomes are new frequent knee pain, KL grade worsening, medial joint-space narrowing (JSN) worsening and resolution of frequent knee pain."
legend$tipping <- sprintf("Knees of the Lo 2022 cohort with the outcome and the model's covariates observed (the replication's complete cases), one panel per outcome (outcomes as named in the subgroup figure legend). Rows: the walking item's assumed sensitivity (Se); columns: its assumed specificity (Sp), taken to be the same in participants with and without the outcome (non-differential). Each cell is the median adjusted odds ratio over %s draws; shaded cells have a 95%% simulation interval that excludes 1; “–” marks a combination that is infeasible (every draw discarded).",
                          fmt(assumption("bias.tipping_iterations")))
legend$summary_level <- sprintf("Published Table 2 counts of knees in the Lo 2022 cohort (events and total knees among walkers and non-walkers; clustering ignored), by outcome. Crude OR (odds ratio): from those counts; Corrected (95%% SI, simulation interval): median corrected odds ratio over the kept draws with its 95%% simulation interval, “–” when every draw was discarded. Scenario is as in the probabilistic bias analysis table. Discarded: share of the %s draws with Se + Sp ≤ 1 in either outcome group or a corrected 2×2 cell count of zero or less; ✓ (flagged) = more than %s discarded.",
                                 count(run$bias_summary_level$n[1]), pct(threshold))
```

```{=typst}
#set text(hyphenate: auto)
#show figure: set block(sticky: true)
#show table: set text(size: 8.5pt, hyphenate: false)
#show table: set par(justify: false)
```

```{=typst}
#outline(title: [Contents], depth: 1, indent: auto)
#v(8pt)
```

```{r reading-guide}
#| results: asis
cat(oaireport::reading_guide(
  c(`Item walker` = sub("^[^:]+: ", "", item_def),
    `Device walker (the reference standard)` = paste(sub("^[^:]+: ", "", device_def), device_detail),
    `PASE walker` = sub("^[^:]+: ", "", pase_def)),
  notation = "Estimates are shown with their 95% confidence interval as 0.89 (0.86–0.91); “to” joins the limits when one is negative; “–” marks a value that cannot be computed. Tables and figures are numbered A1, A2, … (the assumptions ledger at the end is not); each legend defines its own labels and population."),
  sep = "\n")
```

```{=typst}
#pagebreak()
```

# Summary

The exposure in Lo et al. 2022 is the 96-month survey item "walked for exercise since age 50". We compared it with OAI accelerometry from the 48- and 72-month waves in `r count(run$metrics[["sample.validation.persons"]])` participants, `r count(run$metrics[["sample.lo_subset.persons"]])` of them from the paper's cohort. The device is an imperfect reference: it measures current activity, while the item asks about walking since age 50. Sensitivity and specificity are therefore convergent-validity estimates. Data: the OAI public release; input table versions `r data_release`. Code: `r code_version`.

**Headline.** Against the device-defined habitual walker, the item's sensitivity is `r num(pick("validation", "se"))` (95% CI `r num(pick("validation", "se", "lo"))`–`r num(pick("validation", "se", "hi"))`) and its specificity is `r num(pick("validation", "sp"))` (`r num(pick("validation", "sp", "lo"))`–`r num(pick("validation", "sp", "hi"))`). Se + Sp − 1 (Youden's J) is `r num(youden("validation"))`: the item barely separates device-defined walkers from non-walkers. `r instability_text` `r grid_sentence`

# Samples and device processing

```{r samples}
flow_labels <- c(answered_walking_item = "Answered the walking item (96 months)",
                 with_valid_device_wave = "With a valid device wave (48 or 72 months)",
                 lo2022_cohort = "In the Lo 2022 cohort",
                 lo2022_answered_item = "In the Lo 2022 cohort and answered the walking item",
                 lo2022_answered_with_device = "In the Lo 2022 cohort, answered the item, with device data")
oaireport::compare_table(data.frame(Step = named(run$flow$step, flow_labels),
                                    Participants = count(run$flow$persons)),
                         widths = c(4, 1.5), multipage = FALSE)
```

```{r note-flow}
#| results: asis
cat(oaireport::table_note(legend$flow, label = lab("Table")), sep = "\n")
```

```{r flow-note}
flow <- stats::setNames(run$flow$persons, run$flow$step)
lo_unanswered <- unname(flow["lo2022_cohort"] - flow["lo2022_answered_item"])
lo_unanswered_device <- unname(run$metrics["flow.lo2022_unanswered_with_device"])
flow_note <- if (anyNA(c(lo_unanswered, lo_unanswered_device))) "" else
  sprintf("%s participants in the Lo 2022 cohort did not answer the walking item; %s of them have a valid device wave.",
          count(lo_unanswered), count(lo_unanswered_device))
```

`r flow_note`

```{=typst}
#block(sticky: true)[Our processing of the minute data reproduces the release’s own day-level files:]
```

```{r reproduction}
r <- run$device_reproduction
oaireport::compare_table(data.frame(
  Wave = wave(r$wave), `Release days` = count(r$release_days),
  Matched = count(r$matched_days), `Wear differs` = count(r$wear_mismatch_days),
  `MV differs` = count(r$mv_mismatch_days), `Bout differs` = count(r$bout_mismatch_days),
  `Valid persons (ours / release)` = sprintf("%s / %s", count(r$valid_persons), count(r$release_valid_persons)),
  check.names = FALSE), widths = c(1.4, 1.2, 1, 1.1, 1, 1, 2), multipage = FALSE)
```

```{r note-reproduction}
#| results: asis
cat(oaireport::table_note(legend$reproduction, label = lab("Table")), sep = "\n")
```

# Known groups and dose–response

```{r known-groups-figure}
#| results: asis
#| fig-height: 2.4
cat(oaireport::figure_group(), sep = "\n")
kg <- run$validity_known_groups
kg <- kg[kg$sample == "validation", ]
long <- rbind(
  data.frame(measure = kg$measure, group = "Walkers", median = kg$median_walkers, q25 = kg$q25_walkers, q75 = kg$q75_walkers),
  data.frame(measure = kg$measure, group = "Non-walkers", median = kg$median_nonwalkers, q25 = kg$q25_nonwalkers, q75 = kg$q75_nonwalkers)
)
long$measure <- factor(named(long$measure, measure_names), levels = measure_names)
kg_plot <- ggplot2::ggplot(long, ggplot2::aes(x = median, y = group, colour = group)) +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = q25, xmax = q75), size = 0.3, linewidth = 0.6) +
  ggplot2::facet_wrap(~measure, scales = "free_x") +
  ggplot2::scale_x_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)) +
  ggplot2::scale_colour_manual(values = c(Walkers = oaireport::brief_colours[["item"]],
                                          `Non-walkers` = oaireport::brief_colours[["muted"]]), guide = "none") +
  ggplot2::labs(x = "Median (interquartile range)", y = NULL) + oaireport::theme_oai()
oaireport::save_figure(kg_plot, "known_groups", "2col", height = 2.4)
kg_plot
cat(oaireport::table_note(legend$known_groups_figure, close_group = TRUE, label = lab("Figure")), sep = "\n")
```

The figure shows the validation sample. The table below gives both samples, with differences as walkers minus non-walkers.

```{r known-groups-table}
all_kg <- run$validity_known_groups
m <- all_kg$measure
oaireport::compare_table(data.frame(
  Sample = named(all_kg$sample, sample_names), Measure = named(m, measure_names),
  `Walkers, median` = dnum(all_kg$median_walkers, m), `Non-walkers, median` = dnum(all_kg$median_nonwalkers, m),
  `Hodges–Lehmann (95% CI)` = dnum_ci(all_kg$hl, all_kg$hl_lo, all_kg$hl_hi, m),
  `Rank-biserial` = num(all_kg$rank_biserial),
  `Adjusted median difference` = dnum_ci(all_kg$adj_diff, all_kg$adj_lo, all_kg$adj_hi, m),
  check.names = FALSE), widths = c(1.3, 2.6, 1, 1.1, 1.9, 0.9, 1.9), multipage = FALSE)
```

```{r note-known-groups}
#| results: asis
cat(oaireport::table_note(legend$known_groups, label = lab("Table")), sep = "\n")
```

```{r dose-table}
dz <- run$validity_dose
m <- dz$measure
jt_b <- assumption("validity.jt_permutations")
jt_floor <- 1 / (jt_b + 1)  # smallest attainable permutation p-value
round_up <- function(x, digits = 2) {  # round up to `digits` significant digits, so "<=" stays true
  if (x <= 0) return(0)
  step <- 10^(floor(log10(x)) - digits + 1)
  ceiling(x / step - 1e-9) * step
}
jt_text <- ifelse(dz$jt_p <= jt_floor * (1 + 1e-9),
                  paste0("≤ ", formatC(round_up(jt_floor), format = "g", digits = 2)),
                  formatC(dz$jt_p, format = "g", digits = 2))
level_cell <- function(median, n) sprintf("%s (%s)", dnum(median, m), count(n))
oaireport::compare_table(data.frame(
  Sample = named(dz$sample, sample_names), Measure = named(m, measure_names),
  `None, median (n)` = level_cell(dz$median_none, dz$n_none),
  `Lower, median (n)` = level_cell(dz$median_lower, dz$n_lower),
  `Upper, median (n)` = level_cell(dz$median_upper, dz$n_upper),
  `Trend p (JT)` = jt_text,
  `Upper vs none, adjusted` = dnum_ci(dz$adj_upper, dz$adj_upper_lo, dz$adj_upper_hi, m),
  check.names = FALSE), widths = c(1.08, 2.24, 1.28, 1.28, 1.28, 0.94, 1.54), multipage = FALSE)
no_level <- tapply(dz$n_walkers_no_level, dz$sample, max)  # the same in every measure
no_level <- no_level[intersect(names(sample_names), names(no_level))]
sample_phrase <- c(validation = "the validation sample", lo_subset = "the Lo 2022 subset")
no_level_who <- paste0(cap1(no_amount_group), " have no level")
no_level_text <- if (all(no_level == 0)) {
  "Every walker has an amount level, so none is left out of the lower and upper groups."
} else {
  sprintf("%s and are left out of the lower and upper groups: %s.", no_level_who,
          paste(sprintf("%s in %s", count(no_level), named(names(no_level), sample_phrase)), collapse = " and "))
}
```

```{r note-dose}
#| results: asis
dose_note <- sprintf("Validation sample and Lo 2022 subset; walkers and device measures as defined under Known groups. Median (n) of the device measure: none = non-walkers; lower and upper = walkers at or below and above the validation sample's %s of lifetime sessions (years × months per year × times per month, at the midpoints of the answer categories) among its walkers (the same cut in both samples); n = participants with the measure. %s Trend p: Jonckheere–Terpstra permutation test (%s permutations, so the smallest p-value is 1/%s, shown as “≤ %s”); upper vs none, adjusted: %s.",
                     assumption("exposure.amount_split"), no_level_text, count(jt_b), count(jt_b + 1),
                     formatC(round_up(jt_floor), format = "g", digits = 2), adjusted_def)
cat(oaireport::table_note(dose_note, label = lab("Table")), sep = "\n")
```

# Convergent ranking

```{r convergent}
cv <- run$validity_convergent
oaireport::compare_table(data.frame(
  Sample = named(cv$sample, sample_names), Measure = named(cv$measure, measure_names), n = count(cv$n),
  `Spearman ρ (95% CI)` = oaireport::fmt_ci(cv$rho, cv$lo, cv$hi),
  `Between-wave ρ` = num(cv$between_wave_rho), Reliability = num(cv$reliability),
  `Deattenuated ρ` = num(cv$rho_deattenuated),
  check.names = FALSE), widths = c(1.55, 2.3, 0.75, 2, 1, 1.1, 1.35), multipage = FALSE)
```

```{r note-convergent}
#| results: asis
cat(oaireport::table_note(legend$convergent, label = lab("Table")), sep = "\n")
```

# Classification and differential misclassification

```{r classification}
oaireport::compare_table(data.frame(
  Sample = named(cls$sample, sample_names), Measure = named(cls$measure, class_names), `Estimate (95% CI)` = oaireport::fmt_ci(cls$estimate, cls$lo, cls$hi),
  `k / n` = ifelse(is.na(cls$k), count(cls$n), sprintf("%s / %s", count(cls$k), count(cls$n))),
  check.names = FALSE), widths = c(1.4, 1, 2, 1.4), multipage = FALSE)
```

```{r note-classification}
#| results: asis
cat(oaireport::table_note(legend$classification, label = lab("Table")), sep = "\n")
```

```{r strata-figure}
#| results: asis
#| fig-height: 3.8
cat(oaireport::figure_group(), sep = "\n")
st <- run$validity_strata
capitalise <- function(x) paste0(toupper(substr(x, 1, 1)), substring(x, 2))
st$label <- ifelse(st$variable %in% names(outcome_names),
                   sprintf("%s: %s", named(st$variable, outcome_names), st$stratum),
                   ifelse(st$variable == "pain", sprintf("%s at baseline", capitalise(st$stratum)),
                          gsub("-", "–", st$stratum)))
variable_order <- c("kl", "pain", names(outcome_names))
st <- st[order(match(st$variable, variable_order)), ]
st$label <- factor(st$label, levels = rev(unique(st$label)))
st$sample <- factor(named(st$sample, sample_names), levels = sample_names)
st$measure <- ifelse(st$measure == "se", "Sensitivity", "Specificity")
st_plot <- ggplot2::ggplot(st, ggplot2::aes(x = estimate, y = label, colour = sample, shape = sample)) +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = lo, xmax = hi), size = 0.25, linewidth = 0.5,
                           position = ggplot2::position_dodge(width = 0.5, reverse = TRUE)) +
  ggplot2::facet_wrap(~measure) + ggplot2::scale_x_continuous(limits = c(0, 1)) +
  ggplot2::scale_colour_manual(values = stats::setNames(unname(oaireport::brief_colours[c("item", "navy")]), sample_names)) +
  ggplot2::labs(x = "Estimate (95% CI)", y = NULL) + oaireport::theme_oai()
oaireport::save_figure(st_plot, "strata", "2col", height = 3.8)
st_plot
cat(oaireport::table_note(legend$strata, close_group = TRUE, label = lab("Figure")), sep = "\n")
```

```{r kl-strata}
kl <- run$validity_strata[run$validity_strata$variable == "kl", ]
kl_first <- function(sample) gsub("-", "–", sort(unique(kl$stratum[kl$sample == sample]))[1])
kl_lo <- kl_first("lo_subset")
kl_val <- kl_first("validation")
# Only the strata of this run are described; the radiographic-OA reason is given when they show it
kl_text <- if (is.na(kl_lo) || is.na(kl_val)) {
  ""
} else if (kl_lo == kl_val) {
  sprintf("Both samples' KL strata start at %s.", kl_val)
} else {
  sprintf("The Lo 2022 subset's KL strata start at %s%s; the validation sample's start at %s.",
          kl_lo, if (kl_lo == "KL 2") " because the cohort requires radiographic OA" else "", kl_val)
}
```

`r kl_text`

# PASE walking subscore (benchmark)

```{r pase-pairing}
combination <- assumption("device.wave_combination")
combined_text <- if (combination == "mean") "the mean over each participant's valid device waves" else
  sprintf("the %s device wave", sub(" months", "-month", wave(combination)))
pairing <- assumption("pase.device_pairing")
# A single-wave combination limits the sample to people valid at that wave, so the adjacent-wave
# comparison at the other visit needs a valid device wave at both waves
both_waves_text <- if (pairing == "adjacent_wave" && combination != "mean") {
  other_visit <- sub(" months", "-month", wave(setdiff(c("06", "08"), combination)))
  sprintf(" The sample requires a valid %s device wave, so the %s comparison covers only participants valid at both device waves.",
          sub(" months", "-month", wave(combination)), other_visit)
} else ""
pairing_text <- if (pairing == "adjacent_wave") {
  sprintf("PASE at 48 and 72 months is compared with the device wave of the same visit (that wave's measures and device-defined walker, in participants valid at that wave, deattenuated by the single-wave reliability); PASE at 96 months, the walking item's visit, has no device wave, so it is compared with %s, as the walking item is.%s", combined_text, both_waves_text)
} else {
  sprintf("PASE at every visit is compared with %s, as the walking item is.", combined_text)
}
```

The same statistics for the PASE walking item (walking outside the home in the past 7 days) at each visit, as a benchmark for the walking item. `r pairing_text`

```{=typst}
// Tighter rows for this one long table, so it and its legend fit on the page with the section's opening
#set table(inset: (x: 5pt, y: 3pt))
```

```{r pase}
pz <- run$validity_pase
statistic_names <- c(hl = "Hodges–Lehmann", rank_biserial = "Rank-biserial", adj_diff = "Adjusted difference",
                     rho = "Spearman ρ", rho_deattenuated = "Deattenuated ρ")
statistic <- ifelse(pz$statistic %in% names(statistic_names), statistic_names[pz$statistic], named(pz$statistic, class_names))
# One row per measure and statistic, one estimate (95% CI) and n per visit
pz$cell <- ifelse(pz$statistic %in% c("hl", "adj_diff"), dnum_ci(pz$estimate, pz$lo, pz$hi, pz$measure),
                  oaireport::fmt_ci(pz$estimate, pz$lo, pz$hi))
pz$statistic_name <- statistic
pz$key <- paste(pz$measure, pz$statistic)
pase_rows <- unique(pz[c("measure", "statistic_name", "key")])
pase_wide <- data.frame(Statistic = pase_rows$statistic_name)
pase_labels <- "Statistic"
for (v in unique(pz$visit)) {
  at_v <- pz[pz$visit == v, ][match(pase_rows$key, pz$key[pz$visit == v]), ]
  pase_wide[[paste0("estimate_", v)]] <- at_v$cell
  pase_wide[[paste0("n_", v)]] <- count(at_v$n)
  pase_labels <- c(pase_labels, wave(v), "n")
}
# Each measure (and the device walker) heads its block of rows
block_start <- which(!duplicated(pase_rows$measure))
block_names <- named(pase_rows$measure[block_start], c(measure_names, device_walker = "Against the device walker"))
tinytable::group_tt(
  oaireport::compare_table(pase_wide, labels = pase_labels, widths = c(2.15, 2.2, 0.68, 2.2, 0.68, 2.2, 0.68), multipage = FALSE),
  i = as.list(stats::setNames(block_start, block_names)))
```

```{r note-pase}
#| results: asis
pairing_short <- if (pairing == "adjacent_wave") {
  sprintf("48 and 72 months are paired with the device wave of the same visit and 96 months with %s", combined_text)
} else sprintf("every visit is paired with %s", combined_text)
pase_note <- sprintf("Validation sample only. %s Each block is a device measure (as defined under Known groups) or the device walker (as defined under Classification); each visit shows the estimate (95%% CI) and n, which is each statistic's own denominator: for Hodges–Lehmann, rank-biserial, adjusted difference and AUC, the participants with a PASE walking subscore and valid device data at that visit; for Spearman ρ and deattenuated ρ, the PASE walkers among them; for Se, the device walkers; for Sp, the device non-walkers; for PPV, the PASE walkers; for NPV, the PASE non-walkers. The statistics are otherwise defined as in the known-groups legend (Hodges–Lehmann, rank-biserial, adjusted difference), the convergent legend (Spearman ρ, deattenuated ρ) and the classification legend (Se, Sp, PPV, NPV, AUC), with PASE walkers in place of item walkers and the subscore (0 for no walking) in place of lifetime sessions. Visits: %s.",
                     pase_def, pairing_short)
cat(oaireport::table_note(pase_note, label = lab("Table")), sep = "\n")
```

```{=typst}
#set table(inset: 5pt)
```

# Answer-level sub-analyses

```{r components-setup}
cp <- run$validity_components_pase
ci <- run$validity_components_item
hx <- run$validity_components_hex
cp$visit <- sprintf("%02d", as.integer(cp$visit))
if (nrow(hx)) hx$visit <- sprintf("%02d", as.integer(hx$visit))
answer_labels <- lapply(assumption("components.answer_labels"), unlist)
# Missing values show "–", never "NA"
num_na <- function(x, digits = 2) ifelse(is.na(x), "–", num(x, digits))
count_na <- function(x) ifelse(is.na(x), "–", count(x))
# Permutation p-values, floored as in the dose–response table (jt_floor, round_up)
comp_p <- function(p) {
  ifelse(is.na(p), "–", ifelse(p <= jt_floor * (1 + 1e-9), paste0("≤ ", formatC(round_up(jt_floor), format = "g", digits = 2)),
                              formatC(p, format = "g", digits = 2)))
}
cp_get <- function(v, component, comparator, level, statistic, col = "estimate") {
  x <- cp[cp$visit == v & cp$component == component & cp$comparator == comparator &
          cp$level == level & cp$statistic == statistic, col]
  if (length(x) == 1) x else NA
}
ci_get <- function(sample, component, level, statistic, col = "estimate") {
  x <- ci[ci$sample == sample & ci$component == component & ci$level == level & ci$statistic == statistic, col]
  if (length(x) == 1) x else NA
}
est_ci <- function(est, lo, hi, digits = 2) oaireport::fmt_ci(est, lo, hi, digits)
pct_ci <- function(est, lo, hi) {
  ifelse(is.na(est), "–", sprintf("%s (%s–%s)", num(100 * est, 0), num(100 * lo, 0), num(100 * hi, 0)))
}
med_iqr <- function(m, lo, hi, digits = 1) ifelse(is.na(m), "–", sprintf("%s (%s–%s)", num(m, digits), num(lo, digits), num(hi, digits)))
comp_visits <- c("06", "08")
boot_text <- sprintf("person-level bootstrap 95%% CI, %s resamples", fmt(assumption("validity.bootstrap_reps")))
# The device reference and the PASE walker follow pase.device_pairing and pase.walker_threshold,
# as in the PASE benchmark section (pairing, combined_text, pase_threshold)
comp_paired <- pairing == "adjacent_wave"
comp_device <- if (comp_paired) "the device wave of the same visit" else combined_text
comp_population <- if (comp_paired) "at each visit with its own device wave" else sprintf("at each visit, against %s", combined_text)
comp_wave <- if (comp_paired) "that wave's " else ""
pase_walker_rule <- if (pase_threshold == 0) "at least one day of walking outside the home" else
  sprintf("a PASE walking subscore above %s", fmt(pase_threshold))
```

PASE item 2 asks about walking outside the home in the past 7 days in two parts: on how many days, and for how many hours a day. Each part is compared here with `r comp_device`, first separately and then combined into an estimate of weekly walking time. The walking item's own amount answers are then examined the same way. These analyses are secondary. They were pre-specified on 2026-10-02 after an exploratory look at the PASE frequency answer, and they are reported for every pre-listed cut, without choosing an optimal one.

```{r components-frequency}
freq_table <- do.call(rbind, lapply(comp_visits, function(v) {
  data.frame(
    Answer = answer_labels$pase_days,
    n = count_na(sapply(0:3, function(l) cp_get(v, "frequency", "device_walker", as.character(l), "share", "n"))),
    `Device walkers, % (95% CI)` = sapply(0:3, function(l) pct_ci(
      cp_get(v, "frequency", "device_walker", as.character(l), "share"),
      cp_get(v, "frequency", "device_walker", as.character(l), "share", "lo"),
      cp_get(v, "frequency", "device_walker", as.character(l), "share", "hi"))),
    `Bout-days/week, median (IQR)` = sapply(0:3, function(l) med_iqr(
      cp_get(v, "frequency", "bout_days_per_week", as.character(l), "median"),
      cp_get(v, "frequency", "bout_days_per_week", as.character(l), "median", "lo"),
      cp_get(v, "frequency", "bout_days_per_week", as.character(l), "median", "hi"))),
    check.names = FALSE)
}))
tinytable::group_tt(oaireport::compare_table(freq_table, widths = c(1.6, 0.7, 1.8, 1.8), multipage = FALSE),
                    i = stats::setNames(list(1, 5), wave(comp_visits)))
```

```{r note-components-frequency}
#| results: asis
trend <- vapply(comp_visits, function(v) sprintf("%s, Jonckheere–Terpstra p %s and Spearman ρ %s",
  wave(v), comp_p(cp_get(v, "frequency", "bout_days_per_week", "all", "jt_p")),
  est_ci(cp_get(v, "frequency", "bout_days_per_week", "all", "rho"),
         cp_get(v, "frequency", "bout_days_per_week", "all", "rho", "lo"),
         cp_get(v, "frequency", "bout_days_per_week", "all", "rho", "hi"))), character(1))
freq_population <- if (comp_paired) {
  "Validation-sample participants with a PASE walking answer and a valid device wave at the same visit (48 or 72 months)."
} else sprintf("Validation-sample participants with a PASE walking answer at 48 or 72 months. Every device value in these tables is %s, as for the walking item.", combined_text)
freq_device <- if (comp_paired) "by that visit's device wave alone (the rule defined under Classification, applied to one wave)" else
  "(as defined under Classification)"
cat(oaireport::table_note(sprintf(
  "%s Answer: on how many of the past 7 days the participant walked outside the home (PASE item 2). n: participants giving that answer. Device walkers: the share who are device walkers %s, with Wilson 95%% CI. Bout-days/week: %sdays per week with a purposeful bout, median (interquartile range, IQR). Trend in bout-days/week across the four answers (permutation test as in the dose–response table) and its rank correlation with the answer (95%% CI): %s.",
  freq_population, freq_device, comp_wave, paste(trend, collapse = "; ")), label = lab("Table")), sep = "\n")
```

```{r components-duration}
duration_table <- do.call(rbind, lapply(comp_visits, function(v) {
  data.frame(
    Answer = answer_labels$pase_hours,
    n = count_na(sapply(1:4, function(l) cp_get(v, "duration", "purposeful_min", as.character(l), "median", "n"))),
    `Purposeful-bout min/day, median (IQR)` = sapply(1:4, function(l) med_iqr(
      cp_get(v, "duration", "purposeful_min", as.character(l), "median"),
      cp_get(v, "duration", "purposeful_min", as.character(l), "median", "lo"),
      cp_get(v, "duration", "purposeful_min", as.character(l), "median", "hi"))),
    check.names = FALSE)
}))
tinytable::group_tt(oaireport::compare_table(duration_table, widths = c(2, 0.7, 2.6), multipage = FALSE),
                    i = stats::setNames(list(1, 5), wave(comp_visits)))
```

```{r note-components-duration}
#| results: asis
duration_trend <- vapply(comp_visits, function(v) sprintf("%s, Jonckheere–Terpstra p %s and Spearman ρ %s",
  wave(v), comp_p(cp_get(v, "duration", "purposeful_min", "all", "jt_p")),
  est_ci(cp_get(v, "duration", "purposeful_min", "all", "rho"),
         cp_get(v, "duration", "purposeful_min", "all", "rho", "lo"),
         cp_get(v, "duration", "purposeful_min", "all", "rho", "hi"))), character(1))
cat(oaireport::table_note(sprintf(
  "PASE walkers (%s) in the population of the table above who also gave an hours answer. Answer: for how many hours a day the participant walked outside the home on the days they walked (PASE item 2). n: participants giving that answer. Purposeful-bout min/day: %spurposeful-bout minutes per valid day (as defined under Known groups), median (IQR). Trend across the four answers and rank correlation with the answer (95%% CI), as in the table above: %s.",
  pase_walker_rule, comp_wave, paste(duration_trend, collapse = "; ")), label = lab("Table")), sep = "\n")
```

```{=typst}
#show figure: set block(sticky: false)
```

```{r components-cuts}
pase_cut_names <- c(cut1 = sprintf("At least %s", answer_labels$pase_days[2]),
                    cut2 = sprintf("At least %s", answer_labels$pase_days[3]),
                    cut3 = answer_labels$pase_days[4])
band_cut_names <- function(labels) c(cut1 = sprintf("Walker, at least %s", labels[1]),
                                     cut2 = sprintf("Walker, at least %s", labels[2]),
                                     cut3 = sprintf("Walker, %s", labels[3]))
item_cut_names <- band_cut_names(answer_labels$times)
cut_row <- function(get, definition) {
  cell <- function(m) est_ci(get(m), get(m, "lo"), get(m, "hi"))
  data.frame(Definition = definition, Se = cell("se"), Sp = cell("sp"), PPV = cell("ppv"),
             NPV = cell("npv"), J = cell("j"), n = count_na(get("se", "n") + get("sp", "n")), check.names = FALSE)
}
cut_blocks <- list(); cut_points <- list()
for (v in comp_visits) for (k in names(pase_cut_names)) {
  get <- function(m, col = "estimate") cp_get(v, "frequency", "device_walker", k, m, col)
  cut_blocks[[paste("PASE", v)]] <- rbind(cut_blocks[[paste("PASE", v)]], cut_row(get, pase_cut_names[[k]]))
  cut_points[[length(cut_points) + 1]] <- data.frame(source = sprintf("PASE days, %s", wave(v)), cut = k,
    se = get("se"), fpr = 1 - get("sp"), current = k == "cut1" && pase_threshold == 0)
}
for (component in c("times", "months")) for (s in c("validation", "lo_subset")) {
  names_k <- band_cut_names(answer_labels[[component]])
  for (k in names(names_k)) {
    get <- function(m, col = "estimate") ci_get(s, component, k, m, col)
    block <- paste(component, s)
    cut_blocks[[block]] <- rbind(cut_blocks[[block]], cut_row(get, names_k[[k]]))
    if (component == "times" && s == "validation") cut_points[[length(cut_points) + 1]] <- data.frame(
      source = "Walking item, times per month", cut = k, se = get("se"), fpr = 1 - get("sp"), current = k == "cut1")
  }
}
cut_table <- do.call(rbind, cut_blocks)
rownames(cut_table) <- NULL
item_sample_names <- c("validation sample", "Lo 2022 subset")
block_titles <- c(sprintf("PASE walking days, %s", wave(comp_visits)),
                  sprintf("Walking item, times per month: %s", item_sample_names),
                  sprintf("Walking item, months per year: %s", item_sample_names))
starts <- cumsum(c(1, vapply(cut_blocks, nrow, integer(1))))[seq_along(cut_blocks)]
tinytable::group_tt(oaireport::compare_table(cut_table, widths = c(2.4, 1.6, 1.6, 1.6, 1.6, 1.6, 0.7)),
                    i = as.list(stats::setNames(starts, block_titles)))
```

```{r note-components-cuts}
#| results: asis
cut_pase_device <- if (comp_paired) "by the same visit's device wave alone" else sprintf("by %s", combined_text)
cut_pase_first <- if (pase_threshold == 0) "The first PASE row (any walking) is the PASE walker used elsewhere in this report;" else
  sprintf("The first PASE row (any walking) is not the PASE walker used elsewhere in this report, a walking subscore above %s, which also depends on the hours answer;", fmt(pase_threshold))
cat(oaireport::table_note(sprintf(paste(
  "Each row classifies participants as walkers by one cut of an answer and compares that with the device walker (as defined under Classification): for PASE, the device walker %s, in the population of the PASE frequency table; for the walking item, the device walker over the participant's valid waves, in the validation sample or the Lo 2022 subset (as in the sample flow table).",
  "A walking-item cut is a frequency-restricted walker: a walker whose amount answer is in that band or a higher one; non-walkers fall below every cut, and walkers without an answer to that amount are left out of its rows.",
  "%s the first row of each walking-item block is the item walker, among walkers who answered that amount.",
  "Se, Sp, PPV, NPV: as defined under Classification, with Wilson 95%% CIs. J: Youden's J = Se + Sp − 1, with a %s; 0 means no better than chance and 1 is perfect. n: participants classified."),
  cut_pase_device, cut_pase_first, boot_text), label = lab("Table")), sep = "\n")
```

```{=typst}
#show figure: set block(sticky: true)
```

```{r components-roc}
#| fig-height: 3.2
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
pts <- do.call(rbind, cut_points)
pts$source <- factor(pts$source, levels = unique(pts$source))
iso <- do.call(rbind, lapply(c(0, 0.1, 0.2, 0.3, 0.4), function(j) data.frame(j = j, fpr = c(0, 1 - j), se = c(j, 1))))
roc <- ggplot2::ggplot(pts, ggplot2::aes(fpr, se, colour = source, shape = current)) +
  ggplot2::geom_line(data = iso, ggplot2::aes(fpr, se, group = j), inherit.aes = FALSE,
                     linewidth = 0.25, colour = "grey75", linetype = "dashed") +
  ggplot2::geom_text(data = iso[iso$fpr == 0, ], ggplot2::aes(x = 0.02, y = se, label = paste0("J = ", j)),
                     inherit.aes = FALSE, size = 2.2, colour = "grey45", hjust = 0, vjust = -0.3,
                     family = oaireport::oai_font()) +
  ggplot2::geom_line(ggplot2::aes(group = source), linewidth = 0.4) +
  ggplot2::geom_point(size = 2.2) +
  ggplot2::scale_colour_manual(values = oaireport::oai_palette[1:3]) +
  ggplot2::scale_shape_manual(values = c(`TRUE` = 17, `FALSE` = 16),
                              labels = c(`TRUE` = "Definition used elsewhere", `FALSE` = if (pase_threshold == 0) "Stricter cut" else "Other cut")) +
  ggplot2::coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  ggplot2::labs(x = "1 − specificity", y = "Sensitivity") + oaireport::theme_oai() +
  ggplot2::theme(legend.position = "right", legend.box = "vertical")
oaireport::save_figure(roc, "components_roc", "2col", height = 3.2)
print(roc)
roc_marks <- if (pase_threshold == 0) "Triangles mark the definitions used elsewhere in this report (any PASE walking; the walking item); circles, the stricter cuts." else
  sprintf("Triangles mark the walking item, the definition used elsewhere in this report (the PASE walker used elsewhere, a walking subscore above %s, is not a cut of the days answer); circles, the other cuts.", fmt(pase_threshold))
cat(oaireport::table_note(sprintf("Each point is one cut of the PASE days answer (both visits) or of the walking item's times-per-month answer (validation sample) from the table above, placed by its sensitivity and 1 − specificity against the device walker; the points of one answer are joined from the least strict cut (top right) to the strictest. %s Dashed lines join points of equal Youden's J; further up and to the left is better.", roc_marks), close_group = TRUE, label = lab("Figure")), sep = "\n")
```

```{r components-weekly}
comparator_names <- c(purposeful_week = "Purposeful-bout minutes", mv_week = "All moderate-to-vigorous minutes",
                      light_week = "Light minutes")
# One block of rows per visit (a column block per visit does not fit the page width)
weekly_table <- do.call(rbind, lapply(comp_visits, function(v) {
  do.call(rbind, lapply(names(comparator_names), function(cmp) {
    r <- function(level, stat, col = "estimate") cp_get(v, "weekly", cmp, level, stat, col)
    data.frame(
      `Device measure (per week)` = comparator_names[[cmp]],
      `ρ, all (95% CI)` = est_ci(r("all", "rho"), r("all", "rho", "lo"), r("all", "rho", "hi")),
      `Deattenuated ρ` = num_na(r("all", "rho_deattenuated")),
      `ρ, PASE walkers (95% CI)` = est_ci(r("pase_walkers", "rho"), r("pase_walkers", "rho", "lo"), r("pase_walkers", "rho", "hi")),
      `Self-report − device, median (95% CI)` = est_ci(r("all", "median_diff"), r("all", "median_diff", "lo"),
                                                       r("all", "median_diff", "hi"), digits = 0),
      `Limits of agreement` = ifelse(is.na(r("all", "loa", "lo")), "–",
                                     sprintf("%s to %s", num(r("all", "loa", "lo"), 0), num(r("all", "loa", "hi"), 0))),
      check.names = FALSE)
  }))
}))
weekly_starts <- 1 + length(comparator_names) * (seq_along(comp_visits) - 1)
tinytable::group_tt(oaireport::compare_table(weekly_table, widths = c(2.1, 1.7, 1.3, 1.7, 1.8, 1.5), multipage = FALSE),
                    i = as.list(stats::setNames(weekly_starts, wave(comp_visits))))
```

```{r note-components-weekly}
#| results: asis
scoring <- assumption("pase.walking_scoring")
reliability_n <- unique(stats::na.omit(cp$n[cp$component == "weekly" & cp$statistic == "reliability"]))
reliability_n_text <- if (length(reliability_n) == 1) count(reliability_n) else
  sprintf("%s to %s", count(min(reliability_n)), count(max(reliability_n)))
weekly_n <- vapply(comp_visits, function(v) sprintf("%s at %s", count_na(cp_get(v, "weekly", "purposeful_week", "all", "median_diff", "n")),
                                                    wave(v)), character(1))
weekly_population <- sprintf("Participants of the PASE frequency table whose weekly walking can be estimated (a walker needs an hours answer): %s.",
                             paste(weekly_n, collapse = " and "))
weekly_device <- if (comp_paired) "that visit's wave" else combined_text
reliability_text <- if (comp_paired) {
  sprintf("the single-wave device reliability, the between-wave Spearman ρ of that measure among the %s people valid at both device waves", reliability_n_text)
} else sprintf("the reliability of a participant's mean over one or two device waves, from the between-wave Spearman ρ of that measure among the %s people valid at both device waves and the share of the participants correlated who have two valid waves", reliability_n_text)
cat(oaireport::table_note(sprintf(paste(
  "%s Estimated weekly walking: days walked in the past 7 days × hours a day, both at the PASE scoring midpoints (days %s; hours %s), in minutes per week; 0 for participants who did not walk.",
  "Each device measure is %s, per valid day × 7. Purposeful-bout minutes: moderate-to-vigorous minutes inside purposeful bouts (as defined under Classification); all moderate-to-vigorous minutes: every minute at %s counts/min or more, bouted or not; light minutes (as defined under Known groups), because walking outside the home includes slow walking.",
  "ρ: Spearman rank correlation (95%% CI), in everyone and among PASE walkers. Deattenuated ρ: ρ in everyone divided by the square root of %s (capped at 1).",
  "Self-report − device: the median of the per-person differences in minutes per week, with a %s; the interval is asymmetric and can end at the median itself, because the self-reports take only a few distinct values, so the median moves in steps. Limits of agreement: the 2.5th and 97.5th percentiles of the differences.",
  "PASE counts walking of any intensity, the device counts activity above set intensities, so part of any difference is a difference in what is measured, not only error."),
  weekly_population, paste(num(unlist(scoring$days), 1), collapse = "/"), paste(num(unlist(scoring$hours), 1), collapse = "/"),
  weekly_device, fmt(cutpoint), reliability_text, boot_text), label = lab("Table")), sep = "\n")
```

```{r components-guideline}
guide_minutes <- assumption("reference.min_bout_minutes_per_week")
guide_table <- do.call(rbind, lapply(comp_visits, function(v) {
  r <- function(stat, col = "estimate") cp_get(v, "weekly", "purposeful_week", "guideline", stat, col)
  data.frame(Visit = wave(v), Se = est_ci(r("se"), r("se", "lo"), r("se", "hi")),
             Sp = est_ci(r("sp"), r("sp", "lo"), r("sp", "hi")), J = est_ci(r("j"), r("j", "lo"), r("j", "hi")),
             `Cohen's κ` = est_ci(r("kappa"), r("kappa", "lo"), r("kappa", "hi")),
             n = count_na(r("se", "n") + r("sp", "n")), check.names = FALSE)
}))
oaireport::compare_table(guide_table, widths = c(1.2, 1.8, 1.8, 1.8, 1.8, 0.8), multipage = FALSE)
```

```{r note-components-guideline}
#| results: asis
cat(oaireport::table_note(sprintf("Population of the weekly table, %s. Agreement at the guideline volume: estimated weekly walking (as in the table above) of at least %s minutes against device purposeful-bout minutes of at least %s a week. Se: the share meeting the volume by the device who also meet it by self-report; Sp: the share not meeting it by the device who do not meet it by self-report (Wilson 95%% CIs); J: as in the cut table. Cohen's κ: agreement beyond chance (0 none, 1 perfect), with a %s. n: participants classified.",
  comp_population, num(guide_minutes, 0), num(guide_minutes, 0), boot_text), label = lab("Table")), sep = "\n")
```

```{r components-calibration}
#| fig-height: 3
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
# The step writes the bins in order, so their order of appearance is the bin order
cal <- cp[cp$component == "weekly" & cp$comparator == "purposeful_week" & cp$statistic == "median", ]
bin_levels <- unique(cal$level)
cal$level <- factor(cal$level, levels = bin_levels)
cal$visit_name <- wave(cal$visit)
calib <- ggplot2::ggplot(cal, ggplot2::aes(level, estimate)) +
  ggplot2::geom_pointrange(ggplot2::aes(ymin = lo, ymax = hi), colour = oaireport::brief_colours[["pase"]], size = 0.3) +
  ggplot2::geom_text(ggplot2::aes(y = hi, label = paste0("n = ", count(n))), vjust = -0.6, size = 2.2, colour = "grey35",
                     family = oaireport::oai_font()) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0.05, 0.12))) +
  ggplot2::facet_wrap(~visit_name) +
  ggplot2::labs(x = "Estimated weekly walking from PASE (hours per week)", y = "Device purposeful-bout minutes per week") +
  oaireport::theme_oai()
oaireport::save_figure(calib, "components_calibration", "2col", height = 3)
print(calib)
edges <- unlist(assumption("components.pase_weekly_hours_bins"))
n_edges <- length(edges)
bands <- c("0 (did not walk)", sprintf("under %s", fmt(edges[2])),
           if (n_edges > 2) sprintf("%s to under %s", fmt(edges[2:(n_edges - 1)]), fmt(edges[3:n_edges])),
           sprintf("%s or more", fmt(edges[n_edges])))
cat(oaireport::table_note(sprintf("Population of the weekly table, %s. Point: median; bar: IQR of device purposeful-bout minutes per week, within each band of estimated weekly walking from PASE (as in the weekly table; hours per week: %s). n: participants in the band. A self-report that tracked the device would show medians rising steadily from left to right.",
  comp_population, paste(bands, collapse = ", ")), close_group = TRUE, label = lab("Figure")), sep = "\n")
```

```{r components-agreement}
#| fig-height: 3.2
#| results: asis
cat(oaireport::figure_group(), sep = "\n")
thousands <- function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)
if (nrow(hx)) {
  offsets <- data.frame(vertex = 1:6, ox = c(1, 0, -1, -1, 0, 1), oy = c(1, 2, 1, -1, -2, -1))
  hx$cell <- seq_len(nrow(hx))
  # Each hexagon is drawn with its own row's dx and dy (the cell size differs between visits)
  poly <- merge(hx, offsets, by = NULL)
  poly$px <- poly$x + poly$ox * poly$dx
  poly$py <- poly$y + poly$oy * poly$dy
  poly <- poly[order(poly$cell, poly$vertex), ]
  poly$visit_name <- wave(poly$visit)
  ref_lines <- do.call(rbind, lapply(comp_visits, function(v) data.frame(visit_name = wave(v),
    y = c(cp_get(v, "weekly", "purposeful_week", "all", "median_diff"),
          cp_get(v, "weekly", "purposeful_week", "all", "loa", "lo"),
          cp_get(v, "weekly", "purposeful_week", "all", "loa", "hi")),
    kind = c("Median difference", "Limits of agreement", "Limits of agreement"))))
  ref_lines <- ref_lines[!is.na(ref_lines$y), ]
  agree <- ggplot2::ggplot(poly, ggplot2::aes(px, py, group = cell, fill = count)) +
    ggplot2::geom_polygon(colour = NA) +
    ggplot2::geom_hline(data = ref_lines, ggplot2::aes(yintercept = y, linetype = kind), linewidth = 0.3) +
    ggplot2::scale_fill_gradient(low = oaireport::brief_colours[["muted"]], high = oaireport::brief_colours[["navy"]], name = "Participants") +
    ggplot2::scale_linetype_manual(values = c(`Median difference` = "solid", `Limits of agreement` = "dashed"), name = NULL) +
    ggplot2::scale_x_continuous(labels = thousands) + ggplot2::scale_y_continuous(labels = thousands) +
    ggplot2::facet_wrap(~visit_name) +
    ggplot2::labs(x = "Mean of self-report and device (minutes per week)", y = "Self-report − device (minutes per week)") +
    oaireport::theme_oai() +
    ggplot2::theme(legend.title = ggplot2::element_text(size = 7))
  oaireport::save_figure(agree, "components_agreement", "2col", height = 3.2)
  print(agree)
}
shown <- vapply(comp_visits, function(v) sprintf("%s of the %s participants at %s", count(sum(hx$count[hx$visit == v])),
  count_na(cp_get(v, "weekly", "purposeful_week", "all", "median_diff", "n")), wave(v)), character(1))
cat(oaireport::table_note(sprintf("Population of the weekly table, %s. Each hexagon counts the participants whose mean of estimated weekly walking (PASE) and device purposeful-bout minutes per week (x) and difference between them (y, self-report − device) fall in it; its shade gives that count. Hexagons with fewer than %s participants are not shown, so the figure shows no individual; the hexagons shown hold %s. Solid line: the median difference; dashed lines: the limits of agreement (2.5th and 97.5th percentiles of the differences).",
  comp_population, fmt(assumption("components.min_cell_count")), paste(shown, collapse = " and ")), close_group = TRUE, label = lab("Figure")), sep = "\n")
```

```{r components-item}
item_table <- do.call(rbind, lapply(c("times", "months", "years"), function(component) {
  lv <- c("none", as.character(seq_along(answer_labels[[component]])))
  nm <- c("Non-walkers", answer_labels[[component]])
  data.frame(
    Answer = nm,
    n = count_na(sapply(lv, function(l) ci_get("validation", component, l, "share", "n"))),
    `Device walkers, % (95% CI)` = sapply(lv, function(l) pct_ci(ci_get("validation", component, l, "share"),
      ci_get("validation", component, l, "share", "lo"), ci_get("validation", component, l, "share", "hi"))),
    `Bout-days/week, median (IQR)` = sapply(lv, function(l) med_iqr(ci_get("validation", component, l, "median"),
      ci_get("validation", component, l, "median", "lo"), ci_get("validation", component, l, "median", "hi"))),
    check.names = FALSE)
}))
rownames(item_table) <- NULL
component_titles <- c("Times per month (walkers since age 50)", "Months per year", "Years walked")
sizes <- vapply(c("times", "months", "years"), function(cmp) length(answer_labels[[cmp]]) + 1L, integer(1))
tinytable::group_tt(oaireport::compare_table(item_table, widths = c(1.8, 0.7, 1.8, 1.8), multipage = FALSE),
                    i = as.list(stats::setNames(cumsum(c(1, sizes))[1:3], component_titles)))
```

```{r note-components-item}
#| results: asis
no_band <- vapply(c("times", "months", "years"), function(cmp) ci_get("validation", cmp, "no_band", "count"), numeric(1))
cat(oaireport::table_note(sprintf(
  "Validation sample. Answer: the walking item's amount answers (how many times per month, months per year and years the participant walked for exercise since age 50), with non-walkers as their own row. Device walkers and bout-days/week: as in the PASE frequency table, but over the participant's valid device waves. Walkers who did not give a usable answer (left blank or don't know) are not in that block's rows: %s for times per month, %s for months per year and %s for years. Trend in bout-days/week from non-walkers through the bands (Jonckheere–Terpstra, as in the dose–response table): times per month p %s; months per year p %s.",
  count_na(no_band[["times"]]), count_na(no_band[["months"]]), count_na(no_band[["years"]]),
  comp_p(ci_get("validation", "times", "all", "jt_p")), comp_p(ci_get("validation", "months", "all", "jt_p"))), label = lab("Table")), sep = "\n")
```

```{r components-correction}
correction_table <- do.call(rbind, lapply(names(item_cut_names), function(k) {
  g <- function(stat, col = "estimate") ci_get("lo_subset", "times", k, stat, col)
  data.frame(Definition = item_cut_names[[k]], Se = est_ci(g("se"), g("se", "lo"), g("se", "hi")),
             Sp = est_ci(g("sp"), g("sp", "lo"), g("sp", "hi")), J = est_ci(g("j"), g("j", "lo"), g("j", "hi")),
             `Correction factor 1/J` = ifelse(is.na(g("correction")), "–",
               sprintf("%s (%s–%s)", num(g("correction"), 1), num_na(g("correction", "lo"), 1),
                       ifelse(is.na(g("correction", "hi")), "∞", num(g("correction", "hi"), 1)))),
             check.names = FALSE)
}))
oaireport::compare_table(correction_table, widths = c(2.4, 1.6, 1.6, 1.6, 1.9), multipage = FALSE)
# The meaning of an infinite upper limit is given only when the table shows one
infinite_text <- if (any(grepl("∞", correction_table[["Correction factor 1/J"]], fixed = TRUE)))
  " An upper limit of ∞ means the interval for J reaches 0." else ""
```

```{r note-components-correction}
#| results: asis
cat(oaireport::table_note(sprintf("Lo 2022 subset: the participants of Lo et al. 2022 who answered the walking item and have device data (%s classified here). Each row is a frequency-restricted walker definition by the times-per-month answer, with its Se, Sp and Youden's J against the device walker (the Lo 2022 subset rows of the cut table), and 1/J: the factor by which a record-level misclassification correction scales the difference in exposure prevalence between outcome groups, and its sampling error; a larger J (smaller 1/J) gives a more stable correction. The interval for 1/J comes from J's 95%% CI. Descriptive only: the bias analysis of Lo et al. 2022 keeps the published exposure definition.%s",
  count_na(ci_get("lo_subset", "times", "cut1", "correction", "n")), infinite_text), label = lab("Table")), sep = "\n")
```

# Bias analysis of Lo et al. 2022

```{r bias-numbers}
j_lo <- youden("lo_subset")
# The truncation story (surviving draws pile up at a boundary, so lower limits collapse toward 0)
# is told only when the run shows it: some non-differential estimates are flagged and the lower
# limits of the flagged ones are all below `collapse_limit`. The text quotes the largest of them,
# so a reader can judge "effectively 0".
collapse_limit <- 0.05
nondiff_flagged <- nondiff & pba$flagged
nondiff_or <- pba$or[nondiff_flagged]
nondiff_lo <- pba$lo_total[nondiff_flagged & !is.na(pba$lo_total)]
collapse <- any(nondiff_flagged) && length(nondiff_lo) > 0 && max(nondiff_lo) < collapse_limit
which_nondiff <- if (all(pba$flagged[nondiff])) "the non-differential estimates" else "the flagged non-differential estimates"
# Upper bound on those lower limits, rounded up so "at most" stays true; below 0.05, so "g" gives
# plain decimals (0.015) or scientific notation (1.9e-19), never a long string of zeros
nondiff_lo_max <- formatC(round_up(max(nondiff_lo, 0)), format = "g", digits = 2)
discard_vs_threshold <- if (any_flagged) {
  sprintf("above the pre-specified %s threshold for %s", pct(threshold), flag_text)
} else {
  sprintf("never above the pre-specified %s threshold", pct(threshold))
}
collapse_text <- if (collapse) {
  sprintf("The draws that survive are not a random sample of the priors: they sit just past the boundary where one outcome group's implied number of true walkers reaches zero, and near that boundary the corrected odds ratio collapses toward 0. This truncation produces %s of %s–%s, with lower limits that are effectively 0 (all at most %s): the draws behind them leave an outcome group with almost no true walkers, where the model separates, so these limits are numerical artefacts, not quantities.",
          which_nondiff, num(min(nondiff_or, na.rm = TRUE)), num(max(nondiff_or, na.rm = TRUE)), nondiff_lo_max)
} else ""
```

The record-level correction divides by Se + Sp − 1, which is only `r num(j_lo)` in the Lo 2022 subset, so small changes in the drawn sensitivity and specificity move the corrected odds ratio a long way. The device is an imperfect reference: it measures current activity, while the item asks about walking since age 50, so part of the low specificity is device error rather than item error. A draw is discarded when it implies a true walker prevalence outside 0–1 in either outcome group, when Se + Sp ≤ 1, when the reclassified exposure has a single level, or when the refit gives no finite estimate. Depending on outcome and model, that happened to `r pct_range(pba$discarded[nondiff])` of non-differential draws and `r pct_range(pba$discarded[!nondiff])` of differential draws, `r discard_vs_threshold`. `r collapse_text`

```{r tipping-heading}
#| results: asis
cat("\n", if (any_flagged) "## Tipping-point grid (primary)" else "## Tipping-point grid", "\n\n", sep = "")
```

```{r tipping-numbers}
se_ticks <- sort(unique(tp$se))
sp_ticks <- sort(unique(tp$sp))
device_sp <- pick("lo_subset", "sp")
device_sp_text <- sprintf("The device-based specificity of about %s %s.", num(device_sp),
                          if (device_sp < min(sp_ticks)) "lies left of the grid" else "lies within the grid's range")
# The explanation of infeasible cells is given only when this run has some
infeasible_text <- if (any(is.na(tp$or))) {
  sprintf("Cells shown \"–\" are infeasible because every draw was discarded. The grid runs on each outcome's sample from the Lo 2022 cohort and reclassifies within its two outcome groups; a sensitivity below the share answering yes in either group implies a negative number of true non-walkers. In the Lo 2022 participants with device data, the higher group's share is %s depending on the outcome (%s overall); the grid's samples also include participants without device data, so these shares only approximate each outcome's exact edge.",
          pct_range(higher_share), pct(yes_share))
} else ""
new_pain_grid_text <- if (all_exclude) {
  sprintf("For new frequent knee pain the interval excludes 1 in every feasible cell, so the grid has no tipping point there: %s.",
          corner_text)
} else {
  sprintf("For new frequent knee pain the interval excludes 1 in %d of %d feasible cells.", nrow(grid_np_excludes), nrow(grid_np))
}
```

The grid fixes the item's true sensitivity and specificity, non-differentially and without reference to the device, at Se `r num(min(se_ticks))`–`r num(max(se_ticks))` × Sp `r num(min(sp_ticks))`–`r num(max(sp_ticks))`. Each cell shows the median adjusted odds ratio; shaded cells keep a 95% simulation interval that excludes 1. Non-differential error biases toward the null, so the grid shows how much stronger the association could be, not whether it survives differential error. `r differential_text`; the record-level one keeps the published conclusion for new frequent knee pain (adjusted) in `r pct(headline_diff$conclusion_share)` of its kept draws. `r device_sp_text`

`r infeasible_text` `r new_pain_grid_text`

```{r tipping}
#| results: asis
#| fig-height: 4.6
cat(oaireport::figure_group(), sep = "\n")
tp$outcome <- factor(named(tp$outcome, outcome_names), levels = outcome_names)
tp$label <- ifelse(is.na(tp$or), "–", num(tp$or))
tp$state <- factor(ifelse(is.na(tp$or) | is.na(tp$excludes_1), "infeasible",
                          ifelse(tp$excludes_1, "excludes", "includes")),
                   levels = c("excludes", "includes", "infeasible"))
tip <- ggplot2::ggplot(tp, ggplot2::aes(x = factor(num(sp)), y = factor(num(se)), fill = state)) +
  ggplot2::geom_tile(colour = "white", linewidth = 0.4) +
  ggplot2::geom_text(ggplot2::aes(label = label), size = 1.9, family = oaireport::oai_font(), colour = "grey10") +
  ggplot2::scale_fill_manual(values = c(excludes = "#cde2fb", includes = "grey92", infeasible = "white"),
                             labels = c(excludes = "Interval excludes 1", includes = "Interval includes 1",
                                        infeasible = "Infeasible (all draws discarded)"),
                             drop = FALSE) +
  ggplot2::guides(fill = ggplot2::guide_legend(override.aes = list(colour = "grey60"))) +
  ggplot2::facet_wrap(~outcome, ncol = 2) +
  ggplot2::labs(x = "Specificity", y = "Sensitivity") + oaireport::theme_oai() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5))
oaireport::save_figure(tip, "tipping", "2col", height = 4.6)
tip
cat(oaireport::table_note(legend$tipping, close_group = TRUE, label = lab("Figure")), sep = "\n")
```

## Probabilistic bias analysis

```{r bias-table}
oaireport::compare_table(data.frame(
  Outcome = named(pba$outcome, outcome_names), Model = named(pba$model, model_names), Scenario = pba$scenario,
  Discarded = pct(pba$discarded, 1), Flagged = flag_mark(pba$flagged),
  Observed = num(pba$observed_or), `Bias-adjusted (95% SI)` = or_interval(pba$or, pba$lo_total, pba$hi_total),
  `Conclusion holds` = ifelse(is.na(pba$conclusion_share), "–", pct(pba$conclusion_share)),
  check.names = FALSE), widths = c(2.15, 1.25, 1.9, 1.05, 1, 1.05, 2.2, 1.45))
```

```{r note-pba}
#| results: asis
pba_note <- sprintf("Knees of the Lo 2022 cohort, by outcome; each draw refits the replication's models. Model: the paper's unadjusted and adjusted models; scenario: non-differential = one sensitivity and specificity for everyone, differential = separate values for participants with and without the outcome; observed: odds ratio of our replication. Bias-adjusted (95%% SI): median over the kept draws with its 95%% simulation interval (systematic plus random error); conclusion holds: share of kept draws with the published direction and, where the published result was significant, an interval excluding 1. Discarded: share of the %s draws implying an impossible prevalence (outside 0–1 in either outcome group), Se + Sp ≤ 1, a single exposure level, or a refit with no finite estimate; ✓ (flagged) = more than %s discarded.",
                    count(pba$n[1]), pct(threshold))
cat(oaireport::table_note(pba_note, label = lab("Table")), sep = "\n")
```

```{r forest-bias}
#| results: asis
#| fig-height: 4.4
cat(oaireport::figure_group(), sep = "\n")
lo_or <- run$lo2022_t2[run$lo2022_t2$kind == "or", ]
parts <- do.call(rbind, strsplit(sub("^t2\\.", "", lo_or$metric), ".", fixed = TRUE))
pub <- oaireport::parse_or(lo_or$published)
ours <- oaireport::parse_or(lo_or$ours)
base <- data.frame(outcome = unname(outcome_names[parts[, 1]]), model = unname(model_names[parts[, 2]]))
est <- rbind(
  cbind(base, source = "Published", or = pub$or, lo = pub$lo, hi = pub$hi),
  cbind(base, source = "Replication", or = ours$or, lo = ours$lo, hi = ours$hi),
  data.frame(outcome = unname(outcome_names[pba$outcome]), model = unname(model_names[pba$model]),
             source = ifelse(pba$scenario == "differential", "Bias-adjusted (differential)",
                             "Bias-adjusted (non-differential)"),
             or = pba$or, lo = pba$lo_total, hi = pba$hi_total)
)
est <- est[order(match(est$model, model_names), match(est$outcome, outcome_names)), ]
sources <- c("Published", "Replication", "Bias-adjusted (non-differential)", "Bias-adjusted (differential)")
axis_limits <- c(0.01, 100)
axis_breaks <- c(0.01, 0.1, 1, 10, 100)
forest <- oaireport::forest_plot(est, facet = "model", sources = sources) +
  ggplot2::scale_x_log10(breaks = axis_breaks, labels = format(axis_breaks, scientific = FALSE, drop0trailing = TRUE, trim = TRUE)) +
  ggplot2::coord_cartesian(xlim = axis_limits) +
  # The item's colour for our replication and navy for the published result; orange is PASE's
  ggplot2::scale_colour_manual(values = c(Published = oaireport::brief_colours[["navy"]],
                                          Replication = oaireport::brief_colours[["item"]],
                                          `Bias-adjusted (non-differential)` = oaireport::oai_palette[3],
                                          `Bias-adjusted (differential)` = oaireport::oai_palette[4]), breaks = sources)
oaireport::save_figure(forest, "forest_bias", "2col", height = 4.4)
forest
forest_note <- sprintf("Knees of the Lo 2022 cohort, by outcome. Panels: the unadjusted and adjusted models; rows: outcomes. Published: odds ratio and 95%% CI from Lo et al. 2022 Table 2; replication: ours; bias-adjusted: median and 95%% simulation interval over the kept draws, under non-differential and differential misclassification. Intervals are cut at the axis limits (%s and %s). %s",
                       num(axis_limits[1]), num(axis_limits[2], 0), flagged_points_text)
cat(oaireport::table_note(forest_note, close_group = TRUE, label = lab("Figure")), sep = "\n")
```

## Summary-level correction

```{r summary-level}
sl <- run$bias_summary_level
sl_flagged <- if ("flagged" %in% names(sl)) sl$flagged else sl$discarded > threshold
sl_flag_text <- if (all(sl_flagged)) {
  if (any_flagged) "It is flagged in every row as well." else "It is flagged in every row."
} else if (!any(sl_flagged)) {
  "It is flagged in no row."
} else {
  sprintf("It is flagged in %d of %d rows.", sum(sl_flagged), nrow(sl))
}
oaireport::compare_table(data.frame(
  Outcome = named(sl$outcome, outcome_names), Scenario = sl$scenario,
  Discarded = pct(sl$discarded, 1), Flagged = flag_mark(sl_flagged), `Crude OR` = num(sl$crude_or),
  `Corrected (95% SI)` = or_interval(sl$or, sl$lo, sl$hi), check.names = FALSE),
  widths = c(2.6, 1.6, 1.1, 0.9, 1, 2.2), multipage = FALSE)
```

```{r note-summary-level}
#| results: asis
cat(oaireport::table_note(legend$summary_level, label = lab("Table")), sep = "\n")
```

The summary-level correction above uses the published Table 2 counts and ignores knee clustering. It is the transparent, reproducible-from-the-paper check. It draws its own sensitivity and specificity values from the same priors as the record-level analysis, so its draws are not that analysis's draws, though both sets come from the same priors and differ only by Monte Carlo error. `r sl_flag_text`

# Methods and provenance

```{r priors}
pr <- run$bias_priors
oaireport::compare_table(data.frame(
  Outcome = named(pr$outcome, outcome_names), Scenario = pr$scenario, Stratum = pr$stratum,
  Sensitivity = sprintf("%s / %s", count(pr$se_k), count(pr$se_n)),
  Specificity = sprintf("%s / %s", count(pr$sp_k), count(pr$sp_n)), check.names = FALSE),
  widths = c(2.6, 1.8, 1.2, 1.4, 1.4), multipage = FALSE)
# Validated participants the complete-case model frame drops (a missing covariate) feed the priors only
outside <- run$metrics[grepl("^bias\\..*\\.prior_persons_outside_model$", names(run$metrics))]
outside <- outside[outside > 0]
outside_text <- if (length(outside)) {
  who <- named(sub("^bias\\.(.*)\\.prior_persons_outside_model$", "\\1", names(outside)), outcome_names)
  sprintf(" Validated participants that an outcome's complete-case model frame drops still feed its priors (%s).",
          paste(sprintf("%s: %s", who, count(outside)), collapse = ", "))
} else ""
```

```{r note-priors}
#| results: asis
n_prior <- pr$se_n[pr$stratum == "all"] + pr$sp_n[pr$stratum == "all"]  # persons per outcome
n_prior_text <- if (min(n_prior) == max(n_prior)) count(n_prior[1]) else sprintf("%s to %s by outcome", count(min(n_prior)), count(max(n_prior)))
priors_note <- sprintf("Lo 2022 participants who answered the walking item and have device data (the Lo 2022 subset, %s) with the outcome observed (%s). Sensitivity k / n: device walkers who are walkers by the item over all device walkers; specificity k / n: device non-walkers who are non-walkers over all device non-walkers; walkers and device walkers as defined under Known groups and Classification. Scenario: non-differential = everyone (stratum “all”); differential = by outcome group (“case”: the outcome in either knee, “noncase”: not).",
                       count(run$metrics[["sample.lo_subset.persons"]]), n_prior_text)
cat(oaireport::table_note(priors_note, label = lab("Table")), sep = "\n")
```

Priors are Beta(k + 1, n − k + 1) from the 2×2 counts above, taken in the Lo et al. 2022 participants with device data and the outcome observed.`r outside_text` Reclassification is done within outcome strata, and each iteration refits the replication's own GEE models. Each scenario ran `r count(pba$n[1])` iterations and each tipping-point cell `r count(assumption("bias.tipping_iterations"))`.

The assumptions ledger below lists every analytic choice, its status (confirmed, assumed or open) and its source. It is generated from `assumptions.toml`.

```{r ledger-legend}
#| results: asis
cat(oaireport::table_note("Each table lists the assumptions of one status: confirmed (from a source document), assumed or open. Key: the assumption's name; value: what this run used, with the other allowed choices; source: where it comes from; rationale: why it was chosen. The Variants table lists named sensitivity runs: the assumption values each one sets, and a description."), sep = "\n")
```

```{r ledger}
#| results: asis
cat(oaireport::ledger_markdown("ASSUMPTIONS.md"), sep = "\n")
```
