---
title: "Walking for exercise: validity of the OAI item against accelerometry"
subtitle: "The Lo et al. 2022 exposure, with a probabilistic bias analysis"
date: today
mainfont: Arial
fontsize: 10pt
format:
  typst:
    papersize: us-letter
    margin:
      x: 0.9in
      y: 0.9in
    include-in-header:
      text: |
        #set text(hyphenate: false)
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: cairo_pdf
    fig-width: 7
---

```{r setup}
run <- oaireport::load_results()$default
num <- function(x, digits = 2) formatC(x, format = "f", digits = digits, big.mark = ",")
count <- function(x) formatC(x, format = "d", big.mark = ",")
pct <- function(x, digits = 0) paste0(num(100 * x, digits), "%")
pct_range <- function(x) sprintf("%s–%s%%", num(100 * min(x), 0), num(100 * max(x), 0))
named <- function(key, map) ifelse(key %in% names(map), map[key], key)
# Odds ratios: "–" when every draw was discarded, "<0.01" for limits that collapse toward 0
or_text <- function(x) ifelse(is.na(x), "–", ifelse(x < 0.01, "<0.01", num(x)))
or_interval <- function(or, lo, hi) {
  ifelse(is.na(or), "–", sprintf("%s (%s–%s)", or_text(or), or_text(lo), or_text(hi)))
}
interval <- function(lo, hi, sep = "–") ifelse(is.na(lo), "–", paste0(num(lo), sep, num(hi)))
flag_mark <- function(flagged) ifelse(flagged, "✓", "")
# Device measures: counts/day without decimals, minutes/day to 2 decimals
dnum <- function(x, measure) ifelse(measure == "counts_per_day", num(x, 0), num(x))
measure_names <- c(purposeful_min = "Purposeful-bout min/day", counts_per_day = "Counts/day",
                   light_min = "Light min/day")
outcome_names <- c(new_pain = "New frequent knee pain", kl_worse = "KL grade worsening",
                   jsn_worse = "Medial JSN worsening", improved_pain = "Resolution of frequent knee pain")
model_names <- c(or_unadj = "Unadjusted", or_adj = "Adjusted")
sample_names <- c(validation = "Validation", lo_subset = "Lo 2022 subset")
wave_names <- c(`06` = "48 months", `08` = "72 months", `10` = "96 months")
wave <- function(x) named(sprintf("%02d", as.integer(x)), wave_names)
assumption <- function(key) run$assumptions$assumptions[[key]]$value
threshold <- assumption("bias.max_discard_share")
cls <- run$validity_classification
pick <- function(sample, measure, col = "estimate") cls[[col]][cls$sample == sample & cls$measure == measure]
youden <- function(sample) pick(sample, "se") + pick(sample, "sp") - 1
# Share answering yes: device walkers answering yes (se_k) plus device non-walkers answering yes (sp_n - sp_k)
yes_share <- (pick("lo_subset", "se", "k") + pick("lo_subset", "sp", "n") - pick("lo_subset", "sp", "k")) /
  (pick("lo_subset", "se", "n") + pick("lo_subset", "sp", "n"))
# The same share in each outcome group (case / noncase) of the Lo subset with device data, from the
# differential prior counts; a tipping cell is infeasible when Se is below the higher group's share
group_priors <- run$bias_priors[run$bias_priors$scenario == "differential", ]
group_share <- (group_priors$se_k + group_priors$sp_n - group_priors$sp_k) / (group_priors$se_n + group_priors$sp_n)
higher_share <- tapply(group_share, group_priors$outcome, max)
pba <- run$bias_pba
nondiff <- pba$scenario == "non-differential"
flag_text <- if (all(pba$flagged)) "every estimate" else sprintf("%d of %d estimates", sum(pba$flagged), nrow(pba))
which_ors <- if (all(pba$flagged)) "its" else "those"
tp <- run$bias_tipping
grid_np_all <- tp[tp$outcome == "new_pain", ]
grid_np <- grid_np_all[!is.na(grid_np_all$or), ]
grid_np_excludes <- grid_np[grid_np$excludes_1, ]
all_exclude <- nrow(grid_np_excludes) == nrow(grid_np)
lowest_sum <- min(grid_np$se + grid_np$sp)  # lowest feasible Se + Sp
edge_se <- min(grid_np$se)
below_infeasible <- all(is.na(grid_np_all$or[grid_np_all$se + grid_np_all$sp < lowest_sum - 1e-9]))
grid_sentence <- sprintf(
  "Under non-differential misclassification, the tipping-point grid for new frequent knee pain (adjusted) keeps an interval that excludes 1 in %s feasible cells (the lowest feasible Se + Sp is %s).",
  if (all_exclude) sprintf("all %d", nrow(grid_np)) else sprintf("%d of %d", nrow(grid_np_excludes), nrow(grid_np)),
  num(lowest_sum))
headline_diff <- pba[pba$outcome == "new_pain" & pba$model == "or_adj" & pba$scenario == "differential", ]
# Differential rows of both bias analyses (record-level PBA and summary-level correction)
sl_all <- run$bias_summary_level
sl_all_flagged <- if ("flagged" %in% names(sl_all)) sl_all$flagged else sl_all$discarded > threshold
pba_diff_flagged <- pba$flagged[!nondiff]
sl_diff_flagged <- sl_all_flagged[sl_all$scenario == "differential"]
differential_text <- if (all(pba_diff_flagged) && all(sl_diff_flagged)) {
  "The differential analyses below, the record-level probabilistic bias analysis and the summary-level correction, are both flagged"
} else {
  sprintf("The differential analyses below are the record-level probabilistic bias analysis (%d of %d rows flagged) and the summary-level correction (%d of %d rows flagged)",
          sum(pba_diff_flagged), length(pba_diff_flagged), sum(sl_diff_flagged), length(sl_diff_flagged))
}
info <- run$run_info
code_version <- paste0("oai-analytics ", info$oai_version,
                       if (is.null(info$git_commit)) "" else paste0(", commit ", substr(info$git_commit, 1, 7)),
                       if (isTRUE(info$git_dirty)) " (with uncommitted changes)" else "")
versions <- unlist(info$data_versions)
data_release <- if (length(versions)) paste(sprintf("%s %s", names(versions), versions), collapse = ", ") else "not recorded"
```

```{=typst}
#set text(hyphenate: auto)
#show table: set text(size: 8.5pt, hyphenate: false)
#show table: set par(justify: false)
```

# Summary

The exposure in Lo et al. 2022 is the 96-month survey item "walked for exercise since age 50". We compared it with OAI accelerometry from the 48- and 72-month waves in `r count(run$metrics[["sample.validation.persons"]])` participants, `r count(run$metrics[["sample.lo_subset.persons"]])` of them from the paper's cohort. The device is an imperfect reference: it measures current activity, while the item asks about walking since age 50. Sensitivity and specificity are therefore convergent-validity estimates. Data: the OAI public release; input table versions `r data_release`. Code: `r code_version`.

**Headline.** Against the device-defined habitual walker, the item's sensitivity is `r num(pick("validation", "se"))` (95% CI `r num(pick("validation", "se", "lo"))`–`r num(pick("validation", "se", "hi"))`) and its specificity is `r num(pick("validation", "sp"))` (`r num(pick("validation", "sp", "lo"))`–`r num(pick("validation", "sp", "hi"))`). Se + Sp − 1 (Youden's J) is `r num(youden("validation"))`: the item barely separates device-defined walkers from non-walkers. The probabilistic bias analysis was unstable, discarding `r pct_range(pba$discarded[nondiff])` of draws in the non-differential scenario and `r pct_range(pba$discarded[!nondiff])` in the differential one; `r flag_text` exceeded the pre-specified `r pct(threshold)` discard threshold, so `r which_ors` corrected odds ratios are not interpretable and the tipping-point grid is the primary sensitivity analysis. `r grid_sentence`

# Samples and device processing

```{r samples}
flow_labels <- c(answered_walking_item = "Answered the walking item (96 months)",
                 with_valid_device_wave = "With a valid device wave (48 or 72 months)",
                 lo2022_cohort = "In the Lo 2022 cohort",
                 lo2022_with_device = "In the Lo 2022 cohort with device data")
oaireport::compare_table(data.frame(Step = named(run$flow$step, flow_labels),
                                    Participants = count(run$flow$persons)),
                         widths = c(4, 1.5))
```

Our processing of the minute data reproduces the release's own day-level files:

```{r reproduction}
r <- run$device_reproduction
oaireport::compare_table(data.frame(
  Wave = wave(r$wave), `Release days` = count(r$release_days),
  Matched = count(r$matched_days), `Wear differs` = count(r$wear_mismatch_days),
  `MV differs` = count(r$mv_mismatch_days), `Bout differs` = count(r$bout_mismatch_days),
  `Valid persons (ours / release)` = sprintf("%s / %s", count(r$valid_persons), count(r$release_valid_persons)),
  check.names = FALSE), widths = c(1.4, 1.2, 1, 1.1, 1, 1, 2))
```

# Known groups and dose–response

```{r known-groups-figure}
#| fig-height: 3.2
kg <- run$validity_known_groups
kg <- kg[kg$sample == "validation", ]
long <- rbind(
  data.frame(measure = kg$measure, group = "Walkers", median = kg$median_walkers, q25 = kg$q25_walkers, q75 = kg$q75_walkers),
  data.frame(measure = kg$measure, group = "Non-walkers", median = kg$median_nonwalkers, q25 = kg$q25_nonwalkers, q75 = kg$q75_nonwalkers)
)
long$measure <- factor(named(long$measure, measure_names), levels = measure_names)
kg_plot <- ggplot2::ggplot(long, ggplot2::aes(x = median, y = group, colour = group)) +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = q25, xmax = q75), size = 0.3, linewidth = 0.6) +
  ggplot2::facet_wrap(~measure, scales = "free_x") +
  ggplot2::scale_x_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)) +
  ggplot2::scale_colour_manual(values = oaireport::oai_palette[1:2], guide = "none") +
  ggplot2::labs(x = "Median (interquartile range)", y = NULL) + oaireport::theme_oai()
oaireport::save_figure(kg_plot, "known_groups", "2col", height = 3.2)
kg_plot
```

The figure shows the validation sample. The table below gives both samples, with differences as walkers minus non-walkers.

```{r known-groups-table}
all_kg <- run$validity_known_groups
m <- all_kg$measure
oaireport::compare_table(data.frame(
  Sample = named(all_kg$sample, sample_names), Measure = named(m, measure_names),
  `Walkers, median` = dnum(all_kg$median_walkers, m), `Non-walkers, median` = dnum(all_kg$median_nonwalkers, m),
  `Hodges–Lehmann (95% CI)` = sprintf("%s (%s to %s)", dnum(all_kg$hl, m), dnum(all_kg$hl_lo, m), dnum(all_kg$hl_hi, m)),
  `Rank-biserial` = num(all_kg$rank_biserial),
  `Adjusted median difference` = sprintf("%s (%s to %s)", dnum(all_kg$adj_diff, m), dnum(all_kg$adj_lo, m), dnum(all_kg$adj_hi, m)),
  check.names = FALSE), widths = c(1.3, 2.6, 1, 1.1, 1.9, 0.9, 1.9))
```

```{r dose-table}
dz <- run$validity_dose
m <- dz$measure
jt_b <- assumption("validity.jt_permutations")
jt_floor <- 1 / (jt_b + 1)  # smallest attainable permutation p-value
round_up <- function(x, digits = 2) {  # round up to `digits` significant digits, so "<=" stays true
  step <- 10^(floor(log10(x)) - digits + 1)
  ceiling(x / step - 1e-9) * step
}
jt_text <- ifelse(dz$jt_p <= jt_floor * (1 + 1e-9),
                  paste0("≤ ", formatC(round_up(jt_floor), format = "g", digits = 2)),
                  formatC(dz$jt_p, format = "g", digits = 2))
oaireport::compare_table(data.frame(
  Sample = named(dz$sample, sample_names), Measure = named(m, measure_names), None = dnum(dz$median_none, m),
  Lower = dnum(dz$median_lower, m), Upper = dnum(dz$median_upper, m),
  `Trend p (JT)` = jt_text,
  `Upper vs none, adjusted` = sprintf("%s (%s to %s)", dnum(dz$adj_upper, m), dnum(dz$adj_upper_lo, m), dnum(dz$adj_upper_hi, m)),
  check.names = FALSE), widths = c(1.4, 2.4, 1, 1, 1, 1.1, 2.2))
```

Medians by lifetime walking amount: none (non-walkers) and the lower and upper halves among walkers. The trend test is a Jonckheere–Terpstra permutation test with `r count(jt_b)` permutations, so its smallest attainable p-value is 1/`r count(jt_b + 1)`, shown as "≤ `r formatC(round_up(jt_floor), format = "g", digits = 2)`".

# Convergent ranking

```{r convergent}
cv <- run$validity_convergent
oaireport::compare_table(data.frame(
  Sample = named(cv$sample, sample_names), Measure = named(cv$measure, measure_names), n = count(cv$n),
  `Spearman ρ (95% CI)` = sprintf("%s (%s to %s)", num(cv$rho), num(cv$lo), num(cv$hi)),
  `Between-wave ρ` = num(cv$between_wave_rho), `Deattenuated ρ` = num(cv$rho_deattenuated),
  check.names = FALSE), widths = c(1.4, 2, 0.8, 2, 1.2, 1.2))
```

# Classification and differential misclassification

A device-defined habitual walker averages at least `r num(assumption("reference.min_bout_days_per_week"), 0)` days a week with a purposeful bout (at least `r assumption("device.purposeful_bout_minutes")` minutes of moderate-to-vigorous activity).

```{r classification}
oaireport::compare_table(data.frame(
  Sample = named(cls$sample, sample_names), Measure = toupper(cls$measure), Estimate = num(cls$estimate),
  `95% CI` = interval(cls$lo, cls$hi),
  `k / n` = ifelse(is.na(cls$k), count(cls$n), sprintf("%s / %s", count(cls$k), count(cls$n))),
  check.names = FALSE), widths = c(1.4, 1, 1, 1.4, 1.4))
```

```{r strata-figure}
#| fig-height: 4.6
st <- run$validity_strata
capitalise <- function(x) paste0(toupper(substr(x, 1, 1)), substring(x, 2))
st$label <- ifelse(st$variable %in% names(outcome_names),
                   sprintf("%s: %s", named(st$variable, outcome_names), st$stratum),
                   ifelse(st$variable == "pain", sprintf("%s at baseline", capitalise(st$stratum)),
                          gsub("-", "–", st$stratum)))
variable_order <- c("kl", "pain", names(outcome_names))
st <- st[order(match(st$variable, variable_order)), ]
st$label <- factor(st$label, levels = rev(unique(st$label)))
st$sample <- factor(named(st$sample, sample_names), levels = sample_names)
st$measure <- ifelse(st$measure == "se", "Sensitivity", "Specificity")
st_plot <- ggplot2::ggplot(st, ggplot2::aes(x = estimate, y = label, colour = sample, shape = sample)) +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = lo, xmax = hi), size = 0.25, linewidth = 0.5,
                           position = ggplot2::position_dodge(width = 0.5, reverse = TRUE)) +
  ggplot2::facet_wrap(~measure) + ggplot2::scale_x_continuous(limits = c(0, 1)) +
  ggplot2::scale_colour_manual(values = stats::setNames(oaireport::oai_palette[1:2], sample_names)) +
  ggplot2::labs(x = "Estimate (95% CI)", y = NULL) + oaireport::theme_oai()
oaireport::save_figure(st_plot, "strata", "2col", height = 4.6)
st_plot
```

```{r kl-reference}
kl <- run$validity_strata[run$validity_strata$variable == "kl", ]
kl_first <- function(sample) gsub("-", "–", sort(unique(kl$stratum[kl$sample == sample]))[1])
```

The Lo 2022 subset's KL strata start at `r kl_first("lo_subset")`, because the cohort requires radiographic OA, so its reference stratum differs from the validation sample's `r kl_first("validation")`.

# PASE walking subscore (benchmark)

```{r pase-pairing}
combination <- assumption("device.wave_combination")
combined_text <- if (combination == "mean") "the mean over each participant's valid device waves" else
  sprintf("the %s device wave", sub(" months", "-month", wave(combination)))
pairing_text <- if (assumption("pase.device_pairing") == "adjacent_wave") {
  sprintf("PASE at 48 and 72 months is compared with the device wave of the same visit (that wave's measures and device-defined walker, in participants valid at that wave, deattenuated by the single-wave reliability); PASE at 96 months, the walking item's visit, has no device wave, so it is compared with %s, as the walking item is.", combined_text)
} else {
  sprintf("PASE at every visit is compared with %s, as the walking item is.", combined_text)
}
```

The same statistics for the PASE walking item (walking outside the home in the past 7 days) at each visit, as a benchmark for the walking item. `r pairing_text`

```{r pase}
pz <- run$validity_pase
statistic_names <- c(hl = "Hodges–Lehmann", rank_biserial = "Rank-biserial", adj_diff = "Adjusted difference",
                     rho = "Spearman ρ", rho_deattenuated = "Deattenuated ρ")
statistic <- ifelse(pz$statistic %in% names(statistic_names), statistic_names[pz$statistic], toupper(pz$statistic))
pase_estimate <- ifelse(pz$statistic %in% c("hl", "adj_diff"), dnum(pz$estimate, pz$measure), num(pz$estimate))
pase_ci <- ifelse(is.na(pz$lo), "–",
                  ifelse(pz$statistic %in% c("hl", "adj_diff"),
                         paste0(dnum(pz$lo, pz$measure), " to ", dnum(pz$hi, pz$measure)),
                         paste0(num(pz$lo), " to ", num(pz$hi))))
oaireport::compare_table(data.frame(
  Visit = wave(pz$visit), Against = named(pz$measure, c(measure_names, device_walker = "Device walker")),
  Statistic = statistic, Estimate = pase_estimate,
  `95% CI` = pase_ci, n = count(pz$n), check.names = FALSE),
  widths = c(1, 2, 1.6, 1.2, 1.9, 0.8))
```

# Bias analysis of Lo et al. 2022

```{r bias-numbers}
j_lo <- youden("lo_subset")
nondiff_or <- pba$or[nondiff]
smallest <- min(pba$lo_total[nondiff])
exponent <- floor(log10(smallest))
smallest_text <- sprintf("%s × 10^−%d^", num(smallest / 10^exponent, 1), -exponent)
```

The record-level correction divides by Se + Sp − 1, which is only `r num(j_lo)` in the Lo 2022 subset, so small changes in the drawn sensitivity and specificity move the corrected odds ratio a long way. The device is an imperfect reference: it measures current activity, while the item asks about walking since age 50, so part of the low specificity is device error rather than item error. A draw is discarded when it implies a true walker prevalence outside 0–1 in either outcome group, or when Se + Sp ≤ 1. Depending on outcome and model, that happened to `r pct_range(pba$discarded[nondiff])` of non-differential draws and `r pct_range(pba$discarded[!nondiff])` of differential draws, above the pre-specified `r pct(threshold)` threshold for `r flag_text`. The draws that survive are not a random sample of the priors: they sit just past the boundary where one outcome group's implied number of true walkers reaches zero, and near that boundary the corrected odds ratio collapses toward 0. This truncation produces the non-differential estimates of `r num(min(nondiff_or))`–`r num(max(nondiff_or))` and lower limits as small as `r smallest_text`.

## Tipping-point grid (primary)

```{r tipping-numbers}
se_ticks <- sort(unique(tp$se))
sp_ticks <- sort(unique(tp$sp))
```

The grid fixes the item's true sensitivity and specificity, non-differentially and without reference to the device, at Se `r num(min(se_ticks))`–`r num(max(se_ticks))` × Sp `r num(min(sp_ticks))`–`r num(max(sp_ticks))`. Each cell shows the median adjusted odds ratio; shaded cells keep a 95% simulation interval that excludes 1. Non-differential error biases toward the null, so the grid shows how much stronger the association could be, not whether it survives differential error. `r differential_text`; the record-level one keeps the published conclusion for new frequent knee pain (adjusted) in `r pct(headline_diff$conclusion_share)` of its kept draws. The device-based specificity of about `r num(pick("lo_subset", "sp"))` lies left of the grid.

Cells shown "–" are infeasible because every draw was discarded. The grid runs on each outcome's sample from the Lo 2022 cohort and reclassifies within its two outcome groups; a sensitivity below the share answering yes in either group implies a negative number of true non-walkers. In the Lo 2022 participants with device data, the higher group's share is `r pct_range(higher_share)` depending on the outcome (`r pct(yes_share)` overall); the grid's samples also include participants without device data, so these shares only approximate each outcome's exact edge. `r if (all_exclude && below_infeasible) sprintf("For new frequent knee pain the interval excludes 1 in every feasible cell, so the grid has no tipping point there: %s is the lowest feasible Se + Sp (Se %s at the grid's lowest specificity, %s), and every cell below it is infeasible.", num(lowest_sum), num(edge_se), num(min(sp_ticks))) else sprintf("For new frequent knee pain the interval excludes 1 in %d of %d feasible cells.", nrow(grid_np_excludes), nrow(grid_np))`

```{r tipping}
#| fig-height: 5.6
tp$outcome <- factor(named(tp$outcome, outcome_names), levels = outcome_names)
tp$label <- ifelse(is.na(tp$or), "–", num(tp$or))
tp$state <- factor(ifelse(is.na(tp$or) | is.na(tp$excludes_1), "infeasible",
                          ifelse(tp$excludes_1, "excludes", "includes")),
                   levels = c("excludes", "includes", "infeasible"))
tip <- ggplot2::ggplot(tp, ggplot2::aes(x = factor(num(sp)), y = factor(num(se)), fill = state)) +
  ggplot2::geom_tile(colour = "white", linewidth = 0.4) +
  ggplot2::geom_text(ggplot2::aes(label = label), size = 1.9, family = oaireport::oai_font(), colour = "grey10") +
  ggplot2::scale_fill_manual(values = c(excludes = "#cde2fb", includes = "grey92", infeasible = "white"),
                             labels = c(excludes = "Interval excludes 1", includes = "Interval includes 1",
                                        infeasible = "Infeasible (all draws discarded)"),
                             drop = FALSE) +
  ggplot2::guides(fill = ggplot2::guide_legend(override.aes = list(colour = "grey60"))) +
  ggplot2::facet_wrap(~outcome, ncol = 2) +
  ggplot2::labs(x = "Specificity", y = "Sensitivity") + oaireport::theme_oai() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5))
oaireport::save_figure(tip, "tipping", "2col", height = 5.6)
tip
```

## Probabilistic bias analysis

```{r bias-table}
oaireport::compare_table(data.frame(
  Outcome = named(pba$outcome, outcome_names), Model = named(pba$model, model_names), Scenario = pba$scenario,
  Discarded = pct(pba$discarded, 1), Flagged = flag_mark(pba$flagged),
  Observed = num(pba$observed_or), `Bias-adjusted (95% SI)` = or_interval(pba$or, pba$lo_total, pba$hi_total),
  `Conclusion holds` = ifelse(is.na(pba$conclusion_share), "–", pct(pba$conclusion_share)),
  check.names = FALSE), widths = c(90, 48, 73, 45, 41, 45, 90, 50))
```

✓ marks estimates that discarded more than `r pct(threshold)` of `r count(pba$n[1])` draws. SI is the 95% simulation interval (systematic plus random error). "Conclusion holds" is the share of kept draws with the published direction and, where the published result was significant, an interval excluding 1.

```{r forest-bias}
#| fig-height: 4.4
lo_or <- run$lo2022_t2[run$lo2022_t2$kind == "or", ]
parts <- do.call(rbind, strsplit(sub("^t2\\.", "", lo_or$metric), ".", fixed = TRUE))
pub <- oaireport::parse_or(lo_or$published)
ours <- oaireport::parse_or(lo_or$ours)
base <- data.frame(outcome = unname(outcome_names[parts[, 1]]), model = unname(model_names[parts[, 2]]))
est <- rbind(
  cbind(base, source = "Published", or = pub$or, lo = pub$lo, hi = pub$hi),
  cbind(base, source = "Replication", or = ours$or, lo = ours$lo, hi = ours$hi),
  data.frame(outcome = unname(outcome_names[pba$outcome]), model = unname(model_names[pba$model]),
             source = ifelse(pba$scenario == "differential", "Bias-adjusted (differential)",
                             "Bias-adjusted (non-differential)"),
             or = pba$or, lo = pba$lo_total, hi = pba$hi_total)
)
est <- est[order(match(est$model, model_names), match(est$outcome, outcome_names)), ]
sources <- c("Published", "Replication", "Bias-adjusted (non-differential)", "Bias-adjusted (differential)")
axis_limits <- c(0.01, 100)
axis_breaks <- c(0.01, 0.1, 1, 10, 100)
forest <- oaireport::forest_plot(est, facet = "model", sources = sources) +
  ggplot2::scale_x_log10(breaks = axis_breaks, labels = format(axis_breaks, scientific = FALSE, drop0trailing = TRUE, trim = TRUE)) +
  ggplot2::coord_cartesian(xlim = axis_limits)
oaireport::save_figure(forest, "forest_bias", "2col", height = 4.4)
forest
```

Each panel shows the published odds ratio, our replication, and the median bias-adjusted odds ratio under each scenario with its 95% simulation interval. Intervals are cut at the axis limits (`r num(axis_limits[1])` and `r num(axis_limits[2], 0)`). `r if (all(pba$flagged)) "Every bias-adjusted estimate is flagged" else sprintf("%d of %d bias-adjusted estimates are flagged", sum(pba$flagged), nrow(pba))`, so these points are shown for completeness, not interpretation.

## Summary-level correction

```{r summary-level}
sl <- run$bias_summary_level
sl_flagged <- if ("flagged" %in% names(sl)) sl$flagged else sl$discarded > threshold
oaireport::compare_table(data.frame(
  Outcome = named(sl$outcome, outcome_names), Scenario = sl$scenario,
  Discarded = pct(sl$discarded, 1), Flagged = flag_mark(sl_flagged), `Crude OR` = num(sl$crude_or),
  `Corrected (95% SI)` = or_interval(sl$or, sl$lo, sl$hi), check.names = FALSE),
  widths = c(2.6, 1.6, 1.1, 0.9, 1, 2.2))
```

The summary-level correction above uses the published Table 2 counts and ignores knee clustering. It is the transparent, reproducible-from-the-paper check. `r if (all(sl_flagged)) "It is flagged in every row as well." else sprintf("It is flagged in %d of %d rows.", sum(sl_flagged), nrow(sl))`

# Methods and provenance

```{r priors}
pr <- run$bias_priors
oaireport::compare_table(data.frame(
  Outcome = named(pr$outcome, outcome_names), Scenario = pr$scenario, Stratum = pr$stratum,
  Sensitivity = sprintf("%s / %s", count(pr$se_k), count(pr$se_n)),
  Specificity = sprintf("%s / %s", count(pr$sp_k), count(pr$sp_n)), check.names = FALSE),
  widths = c(2.6, 1.8, 1.2, 1.4, 1.4))
```

Priors are Beta(k + 1, n − k + 1) from the 2×2 counts above, taken in the Lo et al. 2022 participants with device data and the outcome observed. Reclassification is done within outcome strata, and each iteration refits the replication's own GEE models. Each scenario ran `r count(pba$n[1])` iterations and each tipping-point cell `r count(assumption("bias.tipping_iterations"))`.

The assumptions ledger below lists every analytic choice, its status (confirmed, assumed or open) and its source. It is generated from `assumptions.toml`.

```{r ledger}
#| results: asis
ledger <- readLines("ASSUMPTIONS.md", warn = FALSE)
ledger <- ledger[!grepl("^# |^<!--", ledger)]
ledger <- gsub("<br>", " `#linebreak()`{=typst} ", ledger, fixed = TRUE)  # raw HTML is dropped in Typst
spans <- gregexpr("`[A-Za-z][A-Za-z0-9_.{}]*`", ledger)
regmatches(ledger, spans) <- lapply(regmatches(ledger, spans), function(x) {
  gsub("([._])", "\\1\u200b", gsub("`", "", x))
})
cat(ledger, sep = "\n")
```
