---
title: "Walking for exercise: validity of the OAI item against accelerometry"
subtitle: "The Lo et al. 2022 exposure, with a probabilistic bias analysis"
date: today
mainfont: Arial
fontsize: 10pt
format:
  typst:
    papersize: us-letter
    margin:
      x: 0.9in
      y: 0.9in
    include-in-header:
      text: |
        #set text(hyphenate: false)
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: cairo_pdf
    fig-width: 7
---

```{r setup}
run <- oaireport::load_results()$default
num <- function(x, digits = 2) formatC(x, format = "f", digits = digits, big.mark = ",")
count <- function(x) formatC(x, format = "d", big.mark = ",")
pct <- function(x, digits = 0) paste0(num(100 * x, digits), "%")
pct_range <- function(x) {
  lo <- num(100 * min(x), 0)
  hi <- num(100 * max(x), 0)
  if (lo == hi) sprintf("%s%%", lo) else sprintf("%s–%s%%", lo, hi)
}
named <- function(key, map) ifelse(key %in% names(map), map[key], key)
# Odds ratios: "–" when every draw was discarded, "<0.01" for limits that collapse toward 0
or_text <- function(x) ifelse(is.na(x), "–", ifelse(x < 0.01, "<0.01", num(x)))
or_interval <- function(or, lo, hi) {
  ifelse(is.na(or), "–", sprintf("%s (%s–%s)", or_text(or), or_text(lo), or_text(hi)))
}
interval <- function(lo, hi, sep = "–") ifelse(is.na(lo), "–", paste0(num(lo), sep, num(hi)))
flag_mark <- function(flagged) ifelse(flagged, "✓", "")
# Device measures: counts/day without decimals, minutes/day to 2 decimals
dnum <- function(x, measure) ifelse(measure == "counts_per_day", num(x, 0), num(x))
measure_names <- c(purposeful_min = "Purposeful-bout min/day", counts_per_day = "Counts/day",
                   light_min = "Light min/day")
outcome_names <- c(new_pain = "New frequent knee pain", kl_worse = "KL grade worsening",
                   jsn_worse = "Medial JSN worsening", improved_pain = "Resolution of frequent knee pain")
model_names <- c(or_unadj = "Unadjusted", or_adj = "Adjusted")
sample_names <- c(validation = "Validation", lo_subset = "Lo 2022 subset")
wave_names <- c(`06` = "48 months", `08` = "72 months", `10` = "96 months")
wave <- function(x) named(sprintf("%02d", as.integer(x)), wave_names)
assumption <- function(key) run$assumptions$assumptions[[key]]$value
threshold <- assumption("bias.max_discard_share")
cls <- run$validity_classification
pick <- function(sample, measure, col = "estimate") cls[[col]][cls$sample == sample & cls$measure == measure]
youden <- function(sample) pick(sample, "se") + pick(sample, "sp") - 1
# Share answering yes: device walkers answering yes (se_k) plus device non-walkers answering yes (sp_n - sp_k)
yes_share <- (pick("lo_subset", "se", "k") + pick("lo_subset", "sp", "n") - pick("lo_subset", "sp", "k")) /
  (pick("lo_subset", "se", "n") + pick("lo_subset", "sp", "n"))
# The same share in each outcome group (case / noncase) of the Lo subset with device data, from the
# differential prior counts; a tipping cell is infeasible when Se is below the higher group's share
group_priors <- run$bias_priors[run$bias_priors$scenario == "differential", ]
group_share <- (group_priors$se_k + group_priors$sp_n - group_priors$sp_k) / (group_priors$se_n + group_priors$sp_n)
higher_share <- tapply(group_share, group_priors$outcome, max)
pba <- run$bias_pba
nondiff <- pba$scenario == "non-differential"
flag_text <- if (all(pba$flagged)) "every estimate" else sprintf("%d of %d estimates", sum(pba$flagged), nrow(pba))
which_ors <- if (all(pba$flagged)) "its" else "those"
# The instability story is told only when the run has flagged estimates
any_flagged <- any(pba$flagged)
flagged_points_text <- if (!any_flagged) {
  "No bias-adjusted estimate is flagged."
} else if (all(pba$flagged)) {
  "Every bias-adjusted estimate is flagged, so these points are shown for completeness, not interpretation."
} else {
  sprintf("%d of %d bias-adjusted estimates are flagged; the flagged ones are shown for completeness, not interpretation.",
          sum(pba$flagged), nrow(pba))
}
discarded_text <- sprintf("%s of draws in the non-differential scenario and %s in the differential one",
                          pct_range(pba$discarded[nondiff]), pct_range(pba$discarded[!nondiff]))
instability_text <- if (any_flagged) {
  sprintf("The probabilistic bias analysis was unstable, discarding %s; %s exceeded the pre-specified %s discard threshold, so %s corrected odds ratios are not interpretable and the tipping-point grid is the primary sensitivity analysis.",
          discarded_text, flag_text, pct(threshold), which_ors)
} else {
  sprintf("The probabilistic bias analysis discarded %s; no estimate exceeded the pre-specified %s discard threshold.",
          discarded_text, pct(threshold))
}
tp <- run$bias_tipping
grid_np_all <- tp[tp$outcome == "new_pain", ]
grid_np <- grid_np_all[!is.na(grid_np_all$or), ]
grid_np_excludes <- grid_np[grid_np$excludes_1, ]
all_exclude <- nrow(grid_np_excludes) == nrow(grid_np)
lowest_sum <- min(grid_np$se + grid_np$sp)  # lowest feasible Se + Sp
# The feasible cell(s) at that corner: the Se and Sp to name, not the extremes of each axis
corner <- grid_np[abs(grid_np$se + grid_np$sp - lowest_sum) < 1e-9, ]
corner_cells <- paste(sprintf("Se %s and Sp %s", num(corner$se), num(corner$sp)), collapse = "; ")
corner_text <- if (nrow(corner) > 1) {
  sprintf("the feasible cells with the lowest Se + Sp are %s (each a sum of %s)", corner_cells, num(lowest_sum))
} else {
  sprintf("the feasible cell with the lowest Se + Sp has %s (a sum of %s)", corner_cells, num(lowest_sum))
}
headline_diff <- pba[pba$outcome == "new_pain" & pba$model == "or_adj" & pba$scenario == "differential", ]
# Differential rows of both bias analyses (record-level PBA and summary-level correction)
sl_all <- run$bias_summary_level
sl_all_flagged <- if ("flagged" %in% names(sl_all)) sl_all$flagged else sl_all$discarded > threshold
pba_diff_flagged <- pba$flagged[!nondiff]
sl_diff_flagged <- sl_all_flagged[sl_all$scenario == "differential"]
differential_text <- if (all(pba_diff_flagged) && all(sl_diff_flagged)) {
  "The differential analyses below, the record-level probabilistic bias analysis and the summary-level correction, are both flagged"
} else {
  sprintf("The differential analyses below are the record-level probabilistic bias analysis (%d of %d rows flagged) and the summary-level correction (%d of %d rows flagged)",
          sum(pba_diff_flagged), length(pba_diff_flagged), sum(sl_diff_flagged), length(sl_diff_flagged))
}
# The grid is non-differential, so it cannot speak to differential error (Summary clause)
diff_flagged <- c(pba_diff_flagged, sl_diff_flagged)
differential_clause <- if (all(diff_flagged)) {
  "both differential bias analyses (record-level and summary-level) are flagged as unstable"
} else if (!any(diff_flagged)) {
  "neither differential bias analysis (record-level or summary-level) is flagged as unstable"
} else {
  sprintf("the differential bias analyses are flagged as unstable in %d of %d rows", sum(diff_flagged), length(diff_flagged))
}
grid_sentence <- sprintf(
  "Under non-differential misclassification, the tipping-point grid for new frequent knee pain (adjusted) keeps an interval that excludes 1 in %s feasible cells (the lowest feasible Se + Sp is %s), but non-differential error biases toward the null, so the grid cannot show whether the association survives differential error, and %s.",
  if (all_exclude) sprintf("all %d", nrow(grid_np)) else sprintf("%d of %d", nrow(grid_np_excludes), nrow(grid_np)),
  num(lowest_sum), differential_clause)
info <- run$run_info
code_version <- paste0("oai-analytics ", info$oai_version,
                       if (is.null(info$git_commit)) "" else paste0(", commit ", substr(info$git_commit, 1, 7)),
                       if (isTRUE(info$git_dirty)) " (with uncommitted changes)" else "")
versions <- unlist(info$data_versions)
data_release <- if (length(versions)) paste(sprintf("%s %s", names(versions), versions), collapse = ", ") else "not recorded"
```

```{r legends}
# Legends under the tables and figures, in plain words. Every number comes from the run's
# assumptions or results, so a variant run is described as it ran.
fmt <- function(x) format(x, trim = TRUE, big.mark = ",")
cap1 <- function(x) paste0(toupper(substr(x, 1, 1)), substring(x, 2))
combination <- assumption("device.wave_combination")
waves_at <- if (combination == "mean") "48 or 72 months" else wave(combination)
averaged <- if (combination == "mean") "averaged over a person's valid waves" else
  sprintf("at the %s wave", sub(" months", "-month", wave(combination)))
cutpoint <- assumption("device.mv_cutpoint")
purposeful_minutes <- fmt(assumption("device.purposeful_bout_minutes"))
valid_day <- sprintf("a valid day has at least %s wear hours and a valid wave at least %s valid days",
                     fmt(assumption("device.valid_day_hours")), fmt(assumption("device.min_valid_days")))

# The walking item
yes_without <- assumption("exposure.yes_without_amount_as")
item_def <- switch(yes_without,
  `non-walker` = "Walker: answered yes to the 96-month walking-for-exercise question (V10WLKAR4) and gave an amount; non-walker: answered no, or yes with no amount.",
  walker = "Walker: answered yes to the 96-month walking-for-exercise question (V10WLKAR4), with or without an amount; non-walker: answered no.",
  exclude = "Walker: answered yes to the 96-month walking-for-exercise question (V10WLKAR4) and gave an amount; non-walker: answered no (a yes with no amount is left out).",
  stop("unknown exposure.yes_without_amount_as: ", yes_without))
sessions_def <- "Lifetime sessions: years × months per year × times per month, at the midpoints of the answer categories (walkers only)."

# The device reference standard
walker_rule <- assumption("reference.walker_rule")
rule_short <- switch(walker_rule,
  bout_days = sprintf("on average at least %s days a week with a purposeful bout", fmt(assumption("reference.min_bout_days_per_week"))),
  any_bout = "at least one purposeful bout",
  bout_minutes = sprintf("on average at least %s minutes a week in purposeful bouts", fmt(assumption("reference.min_bout_minutes_per_week"))),
  stop("unknown reference.walker_rule: ", walker_rule))
device_def <- sprintf("Device walker (the reference standard): defined from the accelerometer alone (ActiGraph GT1M, %s), as a person who shows %s; a purposeful bout is at least %s minutes of moderate-to-vigorous activity.",
                      waves_at, rule_short, purposeful_minutes)
per_week <- switch(walker_rule,
  bout_days = sprintf("Days a week = the share of a person's first %s valid days with a bout × 7, %s. ", fmt(assumption("device.max_valid_days")), averaged),
  bout_minutes = sprintf("Minutes a week = purposeful-bout minutes per valid day × 7, %s. ", averaged),
  "")
device_detail <- sprintf("A bout starts when %s of %s minutes reach %s counts/min and ends when %s minutes in the window fall below it. %s%s.",
                         fmt(assumption("device.bout_need")), fmt(assumption("device.bout_window")), fmt(cutpoint),
                         fmt(assumption("device.bout_stop_below")), per_week, cap1(valid_day))
measures_def <- sprintf("Device measures are averages per valid day: purposeful-bout min/day (minutes inside purposeful bouts), counts/day (total activity counts) and light min/day (minutes at %s–%s counts/min).",
                        fmt(assumption("device.light_floor")), fmt(cutpoint - 1))

# PASE
pase_threshold <- assumption("pase.walker_threshold")
pase_def <- if (pase_threshold == 0) {
  "PASE walker: a PASE walking subscore above 0, i.e. any walking outside the home in the past 7 days (PASE item 2), at that visit."
} else {
  sprintf("PASE walker: a PASE walking subscore above %s (PASE item 2, walking outside the home in the past 7 days) at that visit.", fmt(pase_threshold))
}

# Models and tests
reps_text <- if (assumption("validity.median_regression_se") == "boot") {
  sprintf("bootstrap 95%% CI, %s resamples", fmt(assumption("validity.bootstrap_reps")))
} else "95% CI from sandwich standard errors"
adjusted_def <- sprintf("median regression adjusting for age, sex and BMI (%s)", reps_text)
jt_permutations <- assumption("validity.jt_permutations")

legend <- list()
legend$flow <- sprintf("Validation sample (first two rows): participants who answered the walking item, and of those the ones with a valid device wave at %s (%s). Lo 2022 cohort (last three rows): the cohort of the replication, then its members who answered the item, then those who also have a valid device wave, the Lo 2022 subset.",
                       waves_at, valid_day)
legend$reproduction <- sprintf("Release days: person-days in the release's by-day accelerometry file; matched: those also among our first %s valid days. Wear, MV and bout differs: matched days on which our wear minutes (by half a minute or more), moderate-to-vigorous (MV) minutes or MV-bout minutes differ from the release's. Valid persons: participants with at least %s valid days, ours / the release's own count.",
                               fmt(assumption("device.max_valid_days")), fmt(assumption("device.min_valid_days")))
legend$known_groups_figure <- sprintf("Point: median; bar: interquartile range, by walking-item status, in the validation sample. %s %s", item_def, measures_def)
legend$known_groups <- sprintf("Validation sample and Lo 2022 subset; walkers and non-walkers by the walking item, and the device measures, as defined above. Hodges–Lehmann: median of all walker − non-walker differences (95%% CI); rank-biserial: effect size from −1 to 1; adjusted median difference: walkers minus non-walkers from %s.",
                               adjusted_def)
legend$convergent <- sprintf("Walkers with a known amount, in the validation sample and the Lo 2022 subset (n). Spearman ρ: rank correlation (95%% CI) between lifetime sessions and the device measure; between-wave ρ: correlation of that measure between 48 and 72 months in participants valid at both waves; deattenuated ρ: ρ corrected for that unreliability. %s %s",
                             sessions_def, measures_def)
legend$classification <- sprintf("Validation sample and Lo 2022 subset. Se: share of device walkers who are walkers by the item; Sp: share of device non-walkers who are non-walkers by the item; PPV and NPV: share of item walkers and of item non-walkers who are device walkers and non-walkers; AUC: chance that a device walker has more lifetime sessions than a device non-walker (non-walkers score 0, walkers with no amount are left out); k / n: count over its denominator. %s %s %s",
                                 item_def, device_def, device_detail)
legend$strata <- sprintf("Se and Sp (95%% CI) against the device walker, within subgroups of the validation sample and the Lo 2022 subset: KL 0–1, KL 2 and KL 3–4 are the baseline Kellgren–Lawrence grade of the worse knee; frequent pain and no frequent pain refer to either knee at baseline; event and no event (Lo 2022 subset only) mean the outcome occurred in either knee during follow-up. %s",
                         device_def)
legend$tipping <- sprintf("Rows: the walking item's assumed sensitivity (Se); columns: its assumed specificity (Sp), taken to be the same in participants with and without the outcome (non-differential). Each cell is the median adjusted odds ratio over %s draws; shaded cells have a 95%% simulation interval that excludes 1; “–” marks a combination that is infeasible (every draw discarded). One panel per outcome.",
                          fmt(assumption("bias.tipping_iterations")))
legend$summary_level <- "Crude OR: odds ratio from the published Table 2 counts (events and participants among walkers and non-walkers; knees, clustering ignored). Corrected (95% SI): median corrected odds ratio over the kept draws with its 95% simulation interval, “–” when every draw was discarded. Scenario, discarded and flagged are as in the table above."
legend$priors <- sprintf("Lo 2022 participants with device data and the outcome observed. Sensitivity k / n: device walkers who are walkers by the item over all device walkers; specificity k / n: device non-walkers who are non-walkers over all device non-walkers. Scenario: non-differential = everyone (stratum “all”); differential = by outcome group (“case”: the outcome in either knee, “noncase”: not). %s %s",
                        device_def, item_def)
```

```{=typst}
#set text(hyphenate: auto)
#show figure: set block(sticky: true)
#show table: set text(size: 8.5pt, hyphenate: false)
#show table: set par(justify: false)
```

# Summary

The exposure in Lo et al. 2022 is the 96-month survey item "walked for exercise since age 50". We compared it with OAI accelerometry from the 48- and 72-month waves in `r count(run$metrics[["sample.validation.persons"]])` participants, `r count(run$metrics[["sample.lo_subset.persons"]])` of them from the paper's cohort. The device is an imperfect reference: it measures current activity, while the item asks about walking since age 50. Sensitivity and specificity are therefore convergent-validity estimates. Data: the OAI public release; input table versions `r data_release`. Code: `r code_version`.

**Headline.** Against the device-defined habitual walker, the item's sensitivity is `r num(pick("validation", "se"))` (95% CI `r num(pick("validation", "se", "lo"))`–`r num(pick("validation", "se", "hi"))`) and its specificity is `r num(pick("validation", "sp"))` (`r num(pick("validation", "sp", "lo"))`–`r num(pick("validation", "sp", "hi"))`). Se + Sp − 1 (Youden's J) is `r num(youden("validation"))`: the item barely separates device-defined walkers from non-walkers. `r instability_text` `r grid_sentence`

# Samples and device processing

```{r samples}
flow_labels <- c(answered_walking_item = "Answered the walking item (96 months)",
                 with_valid_device_wave = "With a valid device wave (48 or 72 months)",
                 lo2022_cohort = "In the Lo 2022 cohort",
                 lo2022_answered_item = "In the Lo 2022 cohort and answered the walking item",
                 lo2022_answered_with_device = "In the Lo 2022 cohort, answered the item, with device data")
oaireport::compare_table(data.frame(Step = named(run$flow$step, flow_labels),
                                    Participants = count(run$flow$persons)),
                         widths = c(4, 1.5), multipage = FALSE)
```

```{r note-flow}
#| results: asis
cat(oaireport::table_note(legend$flow), sep = "\n")
```

```{r flow-note}
flow <- stats::setNames(run$flow$persons, run$flow$step)
lo_unanswered <- unname(flow["lo2022_cohort"] - flow["lo2022_answered_item"])
lo_unanswered_device <- unname(run$metrics["flow.lo2022_unanswered_with_device"])
flow_note <- if (anyNA(c(lo_unanswered, lo_unanswered_device))) "" else
  sprintf("%s participants in the Lo 2022 cohort did not answer the walking item; %s of them have a valid device wave.",
          count(lo_unanswered), count(lo_unanswered_device))
```

`r flow_note`

```{=typst}
#block(sticky: true)[Our processing of the minute data reproduces the release’s own day-level files:]
```

```{r reproduction}
r <- run$device_reproduction
oaireport::compare_table(data.frame(
  Wave = wave(r$wave), `Release days` = count(r$release_days),
  Matched = count(r$matched_days), `Wear differs` = count(r$wear_mismatch_days),
  `MV differs` = count(r$mv_mismatch_days), `Bout differs` = count(r$bout_mismatch_days),
  `Valid persons (ours / release)` = sprintf("%s / %s", count(r$valid_persons), count(r$release_valid_persons)),
  check.names = FALSE), widths = c(1.4, 1.2, 1, 1.1, 1, 1, 2), multipage = FALSE)
```

```{r note-reproduction}
#| results: asis
cat(oaireport::table_note(legend$reproduction), sep = "\n")
```

# Known groups and dose–response

```{r known-groups-figure}
#| results: asis
#| fig-height: 2.4
cat(oaireport::figure_group(), sep = "\n")
kg <- run$validity_known_groups
kg <- kg[kg$sample == "validation", ]
long <- rbind(
  data.frame(measure = kg$measure, group = "Walkers", median = kg$median_walkers, q25 = kg$q25_walkers, q75 = kg$q75_walkers),
  data.frame(measure = kg$measure, group = "Non-walkers", median = kg$median_nonwalkers, q25 = kg$q25_nonwalkers, q75 = kg$q75_nonwalkers)
)
long$measure <- factor(named(long$measure, measure_names), levels = measure_names)
kg_plot <- ggplot2::ggplot(long, ggplot2::aes(x = median, y = group, colour = group)) +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = q25, xmax = q75), size = 0.3, linewidth = 0.6) +
  ggplot2::facet_wrap(~measure, scales = "free_x") +
  ggplot2::scale_x_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)) +
  ggplot2::scale_colour_manual(values = oaireport::oai_palette[1:2], guide = "none") +
  ggplot2::labs(x = "Median (interquartile range)", y = NULL) + oaireport::theme_oai()
oaireport::save_figure(kg_plot, "known_groups", "2col", height = 2.4)
kg_plot
cat(oaireport::table_note(legend$known_groups_figure, close_group = TRUE), sep = "\n")
```

The figure shows the validation sample. The table below gives both samples, with differences as walkers minus non-walkers.

```{r known-groups-table}
all_kg <- run$validity_known_groups
m <- all_kg$measure
oaireport::compare_table(data.frame(
  Sample = named(all_kg$sample, sample_names), Measure = named(m, measure_names),
  `Walkers, median` = dnum(all_kg$median_walkers, m), `Non-walkers, median` = dnum(all_kg$median_nonwalkers, m),
  `Hodges–Lehmann (95% CI)` = sprintf("%s (%s to %s)", dnum(all_kg$hl, m), dnum(all_kg$hl_lo, m), dnum(all_kg$hl_hi, m)),
  `Rank-biserial` = num(all_kg$rank_biserial),
  `Adjusted median difference` = sprintf("%s (%s to %s)", dnum(all_kg$adj_diff, m), dnum(all_kg$adj_lo, m), dnum(all_kg$adj_hi, m)),
  check.names = FALSE), widths = c(1.3, 2.6, 1, 1.1, 1.9, 0.9, 1.9), multipage = FALSE)
```

```{r note-known-groups}
#| results: asis
cat(oaireport::table_note(legend$known_groups), sep = "\n")
```

```{r dose-table}
dz <- run$validity_dose
m <- dz$measure
jt_b <- assumption("validity.jt_permutations")
jt_floor <- 1 / (jt_b + 1)  # smallest attainable permutation p-value
round_up <- function(x, digits = 2) {  # round up to `digits` significant digits, so "<=" stays true
  if (x <= 0) return(0)
  step <- 10^(floor(log10(x)) - digits + 1)
  ceiling(x / step - 1e-9) * step
}
jt_text <- ifelse(dz$jt_p <= jt_floor * (1 + 1e-9),
                  paste0("≤ ", formatC(round_up(jt_floor), format = "g", digits = 2)),
                  formatC(dz$jt_p, format = "g", digits = 2))
level_cell <- function(median, n) sprintf("%s (%s)", dnum(median, m), count(n))
oaireport::compare_table(data.frame(
  Sample = named(dz$sample, sample_names), Measure = named(m, measure_names),
  `None, median (n)` = level_cell(dz$median_none, dz$n_none),
  `Lower, median (n)` = level_cell(dz$median_lower, dz$n_lower),
  `Upper, median (n)` = level_cell(dz$median_upper, dz$n_upper),
  `Trend p (JT)` = jt_text,
  `Upper vs none, adjusted` = sprintf("%s (%s to %s)", dnum(dz$adj_upper, m), dnum(dz$adj_upper_lo, m), dnum(dz$adj_upper_hi, m)),
  check.names = FALSE), widths = c(54, 112, 64, 64, 64, 47, 77), multipage = FALSE)
no_level <- tapply(dz$n_walkers_no_level, dz$sample, max)  # the same in every measure
no_level <- no_level[intersect(names(sample_names), names(no_level))]
sample_phrase <- c(validation = "the validation sample", lo_subset = "the Lo 2022 subset")
# A "yes" with no amount answers is a walker with no level only under walker coding
no_level_who <- if (assumption("exposure.yes_without_amount_as") == "walker") {
  "Walkers who gave no amount, or only some of the amount items, have no level"
} else {
  "Walkers who answered some but not all of the amount items have no level"
}
no_level_text <- if (all(no_level == 0)) {
  "Every walker has an amount level, so none is left out of the lower and upper groups."
} else {
  sprintf("%s and are left out of the lower and upper groups: %s.", no_level_who,
          paste(sprintf("%s in %s", count(no_level), named(names(no_level), sample_phrase)), collapse = " and "))
}
```

```{r note-dose}
#| results: asis
dose_note <- sprintf("Validation sample and Lo 2022 subset; walkers and non-walkers by the walking item, and the device measures, as defined above. Median (n) of the device measure: none = non-walkers; lower and upper = walkers at or below and above the %s of lifetime sessions (years × months per year × times per month, at category midpoints) among walkers; n = participants with the measure. %s Trend p: Jonckheere–Terpstra permutation test (%s permutations, so the smallest p-value is 1/%s, shown as “≤ %s”); upper vs none, adjusted: %s.",
                     assumption("exposure.amount_split"), no_level_text, count(jt_b), count(jt_b + 1),
                     formatC(round_up(jt_floor), format = "g", digits = 2), adjusted_def)
cat(oaireport::table_note(dose_note), sep = "\n")
```

# Convergent ranking

```{r convergent}
cv <- run$validity_convergent
oaireport::compare_table(data.frame(
  Sample = named(cv$sample, sample_names), Measure = named(cv$measure, measure_names), n = count(cv$n),
  `Spearman ρ (95% CI)` = sprintf("%s (%s to %s)", num(cv$rho), num(cv$lo), num(cv$hi)),
  `Between-wave ρ` = num(cv$between_wave_rho), `Deattenuated ρ` = num(cv$rho_deattenuated),
  check.names = FALSE), widths = c(1.4, 2, 0.8, 2, 1.2, 1.2), multipage = FALSE)
```

```{r note-convergent}
#| results: asis
cat(oaireport::table_note(legend$convergent), sep = "\n")
```

# Classification and differential misclassification

```{r classification}
oaireport::compare_table(data.frame(
  Sample = named(cls$sample, sample_names), Measure = toupper(cls$measure), Estimate = num(cls$estimate),
  `95% CI` = interval(cls$lo, cls$hi),
  `k / n` = ifelse(is.na(cls$k), count(cls$n), sprintf("%s / %s", count(cls$k), count(cls$n))),
  check.names = FALSE), widths = c(1.4, 1, 1, 1.4, 1.4), multipage = FALSE)
```

```{r note-classification}
#| results: asis
cat(oaireport::table_note(legend$classification), sep = "\n")
```

```{r strata-figure}
#| results: asis
#| fig-height: 3.8
cat(oaireport::figure_group(), sep = "\n")
st <- run$validity_strata
capitalise <- function(x) paste0(toupper(substr(x, 1, 1)), substring(x, 2))
st$label <- ifelse(st$variable %in% names(outcome_names),
                   sprintf("%s: %s", named(st$variable, outcome_names), st$stratum),
                   ifelse(st$variable == "pain", sprintf("%s at baseline", capitalise(st$stratum)),
                          gsub("-", "–", st$stratum)))
variable_order <- c("kl", "pain", names(outcome_names))
st <- st[order(match(st$variable, variable_order)), ]
st$label <- factor(st$label, levels = rev(unique(st$label)))
st$sample <- factor(named(st$sample, sample_names), levels = sample_names)
st$measure <- ifelse(st$measure == "se", "Sensitivity", "Specificity")
st_plot <- ggplot2::ggplot(st, ggplot2::aes(x = estimate, y = label, colour = sample, shape = sample)) +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = lo, xmax = hi), size = 0.25, linewidth = 0.5,
                           position = ggplot2::position_dodge(width = 0.5, reverse = TRUE)) +
  ggplot2::facet_wrap(~measure) + ggplot2::scale_x_continuous(limits = c(0, 1)) +
  ggplot2::scale_colour_manual(values = stats::setNames(oaireport::oai_palette[1:2], sample_names)) +
  ggplot2::labs(x = "Estimate (95% CI)", y = NULL) + oaireport::theme_oai()
oaireport::save_figure(st_plot, "strata", "2col", height = 3.8)
st_plot
cat(oaireport::table_note(legend$strata, close_group = TRUE), sep = "\n")
```

```{r kl-reference}
kl <- run$validity_strata[run$validity_strata$variable == "kl", ]
kl_first <- function(sample) gsub("-", "–", sort(unique(kl$stratum[kl$sample == sample]))[1])
kl_lo <- kl_first("lo_subset")
kl_val <- kl_first("validation")
# Only the strata of this run are described; the radiographic-OA reason is given when they show it
kl_text <- if (is.na(kl_lo) || is.na(kl_val)) {
  ""
} else if (kl_lo == kl_val) {
  sprintf("Both samples' KL strata start at %s.", kl_val)
} else {
  sprintf("The Lo 2022 subset's KL strata start at %s%s, so its reference stratum differs from the validation sample's %s.",
          kl_lo, if (kl_lo == "KL 2") ", because the cohort requires radiographic OA" else "", kl_val)
}
```

`r kl_text`

# PASE walking subscore (benchmark)

```{r pase-pairing}
combination <- assumption("device.wave_combination")
combined_text <- if (combination == "mean") "the mean over each participant's valid device waves" else
  sprintf("the %s device wave", sub(" months", "-month", wave(combination)))
pairing <- assumption("pase.device_pairing")
# A single-wave combination limits the sample to people valid at that wave, so the adjacent-wave
# comparison at the other visit needs a valid device wave at both waves
both_waves_text <- if (pairing == "adjacent_wave" && combination != "mean") {
  other_visit <- sub(" months", "-month", wave(setdiff(c("06", "08"), combination)))
  sprintf(" The sample requires a valid %s device wave, so the %s comparison covers only participants valid at both device waves.",
          sub(" months", "-month", wave(combination)), other_visit)
} else ""
pairing_text <- if (pairing == "adjacent_wave") {
  sprintf("PASE at 48 and 72 months is compared with the device wave of the same visit (that wave's measures and device-defined walker, in participants valid at that wave, deattenuated by the single-wave reliability); PASE at 96 months, the walking item's visit, has no device wave, so it is compared with %s, as the walking item is.%s", combined_text, both_waves_text)
} else {
  sprintf("PASE at every visit is compared with %s, as the walking item is.", combined_text)
}
```

The same statistics for the PASE walking item (walking outside the home in the past 7 days) at each visit, as a benchmark for the walking item. `r pairing_text`

```{=typst}
#show figure: set block(sticky: false)
```

```{r pase}
pz <- run$validity_pase
statistic_names <- c(hl = "Hodges–Lehmann", rank_biserial = "Rank-biserial", adj_diff = "Adjusted difference",
                     rho = "Spearman ρ", rho_deattenuated = "Deattenuated ρ")
statistic <- ifelse(pz$statistic %in% names(statistic_names), statistic_names[pz$statistic], toupper(pz$statistic))
pase_estimate <- ifelse(pz$statistic %in% c("hl", "adj_diff"), dnum(pz$estimate, pz$measure), num(pz$estimate))
pase_ci <- ifelse(is.na(pz$lo), "–",
                  ifelse(pz$statistic %in% c("hl", "adj_diff"),
                         paste0(dnum(pz$lo, pz$measure), " to ", dnum(pz$hi, pz$measure)),
                         paste0(num(pz$lo), " to ", num(pz$hi))))
# One row per measure and statistic, one estimate (95% CI) and n per visit
pz$cell <- ifelse(is.na(pz$lo), pase_estimate, paste0(pase_estimate, " (", pase_ci, ")"))
pz$statistic_name <- statistic
pz$key <- paste(pz$measure, pz$statistic)
pase_rows <- unique(pz[c("measure", "statistic_name", "key")])
pase_wide <- data.frame(Statistic = pase_rows$statistic_name)
pase_labels <- "Statistic"
for (v in unique(pz$visit)) {
  at_v <- pz[pz$visit == v, ][match(pase_rows$key, pz$key[pz$visit == v]), ]
  pase_wide[[paste0("estimate_", v)]] <- at_v$cell
  pase_wide[[paste0("n_", v)]] <- count(at_v$n)
  pase_labels <- c(pase_labels, wave(v), "n")
}
# Each measure (and the device walker) heads its block of rows
block_start <- which(!duplicated(pase_rows$measure))
block_names <- named(pase_rows$measure[block_start], c(measure_names, device_walker = "Against the device walker"))
tinytable::group_tt(
  oaireport::compare_table(pase_wide, labels = pase_labels, widths = c(2.15, 2.2, 0.68, 2.2, 0.68, 2.2, 0.68)),
  i = as.list(stats::setNames(block_start, block_names)))
```

```{r note-pase}
#| results: asis
pairing_short <- if (pairing == "adjacent_wave") {
  sprintf("48 and 72 months are paired with the device wave of the same visit and 96 months with %s", combined_text)
} else sprintf("every visit is paired with %s", combined_text)
pase_note <- sprintf("Each block is a device measure (an average per valid day) or the device walker. Each visit shows the estimate (95%% CI) and n, the participants with a PASE walking subscore and valid device data at that visit (%s). Statistics are defined as in the known-groups and classification legends, with PASE walkers in place of item walkers and the subscore (0 for no walking) in place of lifetime sessions; Spearman ρ and deattenuated ρ are among PASE walkers. %s",
                     pairing_short, pase_def)
cat(oaireport::table_note(pase_note), sep = "\n")
```

```{=typst}
#show figure: set block(sticky: true)
```

# Bias analysis of Lo et al. 2022

```{r bias-numbers}
j_lo <- youden("lo_subset")
# The truncation story (surviving draws pile up at a boundary, so lower limits collapse toward 0)
# is told only when the run shows it: some non-differential estimates are flagged and the lower
# limits of the flagged ones are all below `collapse_limit`. The text quotes the largest of them,
# so a reader can judge "effectively 0".
collapse_limit <- 0.05
nondiff_flagged <- nondiff & pba$flagged
nondiff_or <- pba$or[nondiff_flagged]
nondiff_lo <- pba$lo_total[nondiff_flagged & !is.na(pba$lo_total)]
collapse <- any(nondiff_flagged) && length(nondiff_lo) > 0 && max(nondiff_lo) < collapse_limit
which_nondiff <- if (all(pba$flagged[nondiff])) "the non-differential estimates" else "the flagged non-differential estimates"
# Upper bound on those lower limits, rounded up so "at most" stays true; below 0.05, so "g" gives
# plain decimals (0.015) or scientific notation (1.9e-19), never a long string of zeros
nondiff_lo_max <- formatC(round_up(max(nondiff_lo, 0)), format = "g", digits = 2)
discard_vs_threshold <- if (any_flagged) {
  sprintf("above the pre-specified %s threshold for %s", pct(threshold), flag_text)
} else {
  sprintf("never above the pre-specified %s threshold", pct(threshold))
}
collapse_text <- if (collapse) {
  sprintf("The draws that survive are not a random sample of the priors: they sit just past the boundary where one outcome group's implied number of true walkers reaches zero, and near that boundary the corrected odds ratio collapses toward 0. This truncation produces %s of %s–%s, with lower limits that are effectively 0 (all at most %s): the draws behind them leave an outcome group with almost no true walkers, where the model separates, so these limits are numerical artefacts, not quantities.",
          which_nondiff, num(min(nondiff_or, na.rm = TRUE)), num(max(nondiff_or, na.rm = TRUE)), nondiff_lo_max)
} else ""
```

The record-level correction divides by Se + Sp − 1, which is only `r num(j_lo)` in the Lo 2022 subset, so small changes in the drawn sensitivity and specificity move the corrected odds ratio a long way. The device is an imperfect reference: it measures current activity, while the item asks about walking since age 50, so part of the low specificity is device error rather than item error. A draw is discarded when it implies a true walker prevalence outside 0–1 in either outcome group, when Se + Sp ≤ 1, or when the reclassified exposure has a single level. Depending on outcome and model, that happened to `r pct_range(pba$discarded[nondiff])` of non-differential draws and `r pct_range(pba$discarded[!nondiff])` of differential draws, `r discard_vs_threshold`. `r collapse_text`

```{r tipping-heading}
#| results: asis
cat("\n", if (any_flagged) "## Tipping-point grid (primary)" else "## Tipping-point grid", "\n\n", sep = "")
```

```{r tipping-numbers}
se_ticks <- sort(unique(tp$se))
sp_ticks <- sort(unique(tp$sp))
device_sp <- pick("lo_subset", "sp")
device_sp_text <- sprintf("The device-based specificity of about %s %s.", num(device_sp),
                          if (device_sp < min(sp_ticks)) "lies left of the grid" else "lies within the grid's range")
# The explanation of infeasible cells is given only when this run has some
infeasible_text <- if (any(is.na(tp$or))) {
  sprintf("Cells shown \"–\" are infeasible because every draw was discarded. The grid runs on each outcome's sample from the Lo 2022 cohort and reclassifies within its two outcome groups; a sensitivity below the share answering yes in either group implies a negative number of true non-walkers. In the Lo 2022 participants with device data, the higher group's share is %s depending on the outcome (%s overall); the grid's samples also include participants without device data, so these shares only approximate each outcome's exact edge.",
          pct_range(higher_share), pct(yes_share))
} else ""
new_pain_grid_text <- if (all_exclude) {
  sprintf("For new frequent knee pain the interval excludes 1 in every feasible cell, so the grid has no tipping point there: %s.",
          corner_text)
} else {
  sprintf("For new frequent knee pain the interval excludes 1 in %d of %d feasible cells.", nrow(grid_np_excludes), nrow(grid_np))
}
```

The grid fixes the item's true sensitivity and specificity, non-differentially and without reference to the device, at Se `r num(min(se_ticks))`–`r num(max(se_ticks))` × Sp `r num(min(sp_ticks))`–`r num(max(sp_ticks))`. Each cell shows the median adjusted odds ratio; shaded cells keep a 95% simulation interval that excludes 1. Non-differential error biases toward the null, so the grid shows how much stronger the association could be, not whether it survives differential error. `r differential_text`; the record-level one keeps the published conclusion for new frequent knee pain (adjusted) in `r pct(headline_diff$conclusion_share)` of its kept draws. `r device_sp_text`

`r infeasible_text` `r new_pain_grid_text`

```{r tipping}
#| results: asis
#| fig-height: 5.6
cat(oaireport::figure_group(), sep = "\n")
tp$outcome <- factor(named(tp$outcome, outcome_names), levels = outcome_names)
tp$label <- ifelse(is.na(tp$or), "–", num(tp$or))
tp$state <- factor(ifelse(is.na(tp$or) | is.na(tp$excludes_1), "infeasible",
                          ifelse(tp$excludes_1, "excludes", "includes")),
                   levels = c("excludes", "includes", "infeasible"))
tip <- ggplot2::ggplot(tp, ggplot2::aes(x = factor(num(sp)), y = factor(num(se)), fill = state)) +
  ggplot2::geom_tile(colour = "white", linewidth = 0.4) +
  ggplot2::geom_text(ggplot2::aes(label = label), size = 1.9, family = oaireport::oai_font(), colour = "grey10") +
  ggplot2::scale_fill_manual(values = c(excludes = "#cde2fb", includes = "grey92", infeasible = "white"),
                             labels = c(excludes = "Interval excludes 1", includes = "Interval includes 1",
                                        infeasible = "Infeasible (all draws discarded)"),
                             drop = FALSE) +
  ggplot2::guides(fill = ggplot2::guide_legend(override.aes = list(colour = "grey60"))) +
  ggplot2::facet_wrap(~outcome, ncol = 2) +
  ggplot2::labs(x = "Specificity", y = "Sensitivity") + oaireport::theme_oai() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5))
oaireport::save_figure(tip, "tipping", "2col", height = 5.6)
tip
cat(oaireport::table_note(legend$tipping, close_group = TRUE), sep = "\n")
```

## Probabilistic bias analysis

```{r bias-table}
oaireport::compare_table(data.frame(
  Outcome = named(pba$outcome, outcome_names), Model = named(pba$model, model_names), Scenario = pba$scenario,
  Discarded = pct(pba$discarded, 1), Flagged = flag_mark(pba$flagged),
  Observed = num(pba$observed_or), `Bias-adjusted (95% SI)` = or_interval(pba$or, pba$lo_total, pba$hi_total),
  `Conclusion holds` = ifelse(is.na(pba$conclusion_share), "–", pct(pba$conclusion_share)),
  check.names = FALSE), widths = c(86, 50, 76, 42, 40, 42, 88, 58))
```

```{r note-pba}
#| results: asis
pba_note <- sprintf("Model: the paper's unadjusted and adjusted models; scenario: non-differential = one sensitivity and specificity for everyone, differential = separate values for participants with and without the outcome; observed: odds ratio of our replication. Bias-adjusted (95%% SI): median over the kept draws with its 95%% simulation interval (systematic plus random error); conclusion holds: share of kept draws with the published direction and, where the published result was significant, an interval excluding 1. Discarded: share of the %s draws implying an impossible prevalence (outside 0–1 in either outcome group), Se + Sp ≤ 1 or a single exposure level; ✓ (flagged) = more than %s discarded.",
                    count(pba$n[1]), pct(threshold))
cat(oaireport::table_note(pba_note), sep = "\n")
```

```{r forest-bias}
#| results: asis
#| fig-height: 4.4
cat(oaireport::figure_group(), sep = "\n")
lo_or <- run$lo2022_t2[run$lo2022_t2$kind == "or", ]
parts <- do.call(rbind, strsplit(sub("^t2\\.", "", lo_or$metric), ".", fixed = TRUE))
pub <- oaireport::parse_or(lo_or$published)
ours <- oaireport::parse_or(lo_or$ours)
base <- data.frame(outcome = unname(outcome_names[parts[, 1]]), model = unname(model_names[parts[, 2]]))
est <- rbind(
  cbind(base, source = "Published", or = pub$or, lo = pub$lo, hi = pub$hi),
  cbind(base, source = "Replication", or = ours$or, lo = ours$lo, hi = ours$hi),
  data.frame(outcome = unname(outcome_names[pba$outcome]), model = unname(model_names[pba$model]),
             source = ifelse(pba$scenario == "differential", "Bias-adjusted (differential)",
                             "Bias-adjusted (non-differential)"),
             or = pba$or, lo = pba$lo_total, hi = pba$hi_total)
)
est <- est[order(match(est$model, model_names), match(est$outcome, outcome_names)), ]
sources <- c("Published", "Replication", "Bias-adjusted (non-differential)", "Bias-adjusted (differential)")
axis_limits <- c(0.01, 100)
axis_breaks <- c(0.01, 0.1, 1, 10, 100)
forest <- oaireport::forest_plot(est, facet = "model", sources = sources) +
  ggplot2::scale_x_log10(breaks = axis_breaks, labels = format(axis_breaks, scientific = FALSE, drop0trailing = TRUE, trim = TRUE)) +
  ggplot2::coord_cartesian(xlim = axis_limits)
oaireport::save_figure(forest, "forest_bias", "2col", height = 4.4)
forest
forest_note <- sprintf("Panels: the unadjusted and adjusted models; rows: outcomes. Published: odds ratio and 95%% CI from Lo et al. 2022 Table 2; replication: ours; bias-adjusted: median and 95%% simulation interval over the kept draws, under non-differential and differential misclassification. Intervals are cut at the axis limits (%s and %s). %s",
                       num(axis_limits[1]), num(axis_limits[2], 0), flagged_points_text)
cat(oaireport::table_note(forest_note, close_group = TRUE), sep = "\n")
```

## Summary-level correction

```{r summary-level}
sl <- run$bias_summary_level
sl_flagged <- if ("flagged" %in% names(sl)) sl$flagged else sl$discarded > threshold
sl_flag_text <- if (all(sl_flagged)) {
  if (any_flagged) "It is flagged in every row as well." else "It is flagged in every row."
} else if (!any(sl_flagged)) {
  "It is flagged in no row."
} else {
  sprintf("It is flagged in %d of %d rows.", sum(sl_flagged), nrow(sl))
}
oaireport::compare_table(data.frame(
  Outcome = named(sl$outcome, outcome_names), Scenario = sl$scenario,
  Discarded = pct(sl$discarded, 1), Flagged = flag_mark(sl_flagged), `Crude OR` = num(sl$crude_or),
  `Corrected (95% SI)` = or_interval(sl$or, sl$lo, sl$hi), check.names = FALSE),
  widths = c(2.6, 1.6, 1.1, 0.9, 1, 2.2), multipage = FALSE)
```

```{r note-summary-level}
#| results: asis
cat(oaireport::table_note(legend$summary_level), sep = "\n")
```

The summary-level correction above uses the published Table 2 counts and ignores knee clustering. It is the transparent, reproducible-from-the-paper check. It draws its own sensitivity and specificity values from the same priors as the record-level analysis, so its draws are not that analysis's draws, though both sets come from the same priors and differ only by Monte Carlo error. `r sl_flag_text`

# Methods and provenance

```{r priors}
pr <- run$bias_priors
oaireport::compare_table(data.frame(
  Outcome = named(pr$outcome, outcome_names), Scenario = pr$scenario, Stratum = pr$stratum,
  Sensitivity = sprintf("%s / %s", count(pr$se_k), count(pr$se_n)),
  Specificity = sprintf("%s / %s", count(pr$sp_k), count(pr$sp_n)), check.names = FALSE),
  widths = c(2.6, 1.8, 1.2, 1.4, 1.4), multipage = FALSE)
# Validated participants the complete-case model frame drops (a missing covariate) feed the priors only
outside <- run$metrics[grepl("^bias\\..*\\.prior_persons_outside_model$", names(run$metrics))]
outside <- outside[outside > 0]
outside_text <- if (length(outside)) {
  who <- named(sub("^bias\\.(.*)\\.prior_persons_outside_model$", "\\1", names(outside)), outcome_names)
  sprintf(" Validated participants that an outcome's complete-case model frame drops still feed its priors (%s).",
          paste(sprintf("%s: %s", who, count(outside)), collapse = ", "))
} else ""
```

```{r note-priors}
#| results: asis
cat(oaireport::table_note(legend$priors), sep = "\n")
```

Priors are Beta(k + 1, n − k + 1) from the 2×2 counts above, taken in the Lo et al. 2022 participants with device data and the outcome observed.`r outside_text` Reclassification is done within outcome strata, and each iteration refits the replication's own GEE models. Each scenario ran `r count(pba$n[1])` iterations and each tipping-point cell `r count(assumption("bias.tipping_iterations"))`.

The assumptions ledger below lists every analytic choice, its status (confirmed, assumed or open) and its source. It is generated from `assumptions.toml`.

```{r ledger}
#| results: asis
cat(oaireport::ledger_markdown("ASSUMPTIONS.md"), sep = "\n")
```
